### 2022-10-27 Draft_1.Rmd
Inital set up of workspace

---
title: "Untitled"
output: html_document
date: "2022-10-27"
---
```{r}
#Inital enviroment setup
setwd("~/Desktop/GLM/Coursework/GLM-CW")
expend.df <- read.table("expenditure.txt")
library(tidyverse)
library(kableExtra)
```

```{r}

```

### 2022-10-27 Draft_1.Rmd
Setting up df

@@
 ```{r}
+#Data cleaning
+
+#coding variables as factors
+expend.df$house.ten <- factor(expend.df$house.ten,levels=c(1:3),labels=c("Public Rented","Private rented","Owned"))
+expend.df$sex.hh <- factor(expend.df$sex.hh,levels=c(1,2),labels=c("Male","Female"))
+expend.df$lab.force <- factor(expend.df$lab.force,levels=c(1:4),labels=c("Full time", "Part time","Unemployed","Inactive"))
+expend.df$hh.size <- factor(expend.df$hh.size,levels=c(1:5),labels = c("1 person","2 people","3 people","4 people","5+ persons"))
+expend.df$hh.adults <- factor(expend.df$hh.adults,levels = c(1:4), labels = c("1 adult","2 adults","3 adults","4+ adults"))
 
 ```
+
+```{r}
+hist(expend.df$expenditure,breaks = 20,
+     xlab = "Total weekly household expenditure (GBP)")
+
+ggplot(data=expend.df,aes(x=income,y=expenditure))+
+  geom_point()
+```
+Looking at the above histogram it appears to have a skewed distribution

### 2022-10-27 Draft_1.Rmd
test

@@
 title: "Untitled"
-output: html_document
+output:
+  pdf_document: default
+  html_document: default
 date: "2022-10-27"
 ---
-```{r}
-#Inital enviroment setup
+```{r,echo=F}
+#Initial environment setup
 setwd("~/Desktop/GLM/Coursework/GLM-CW")
@@
 library(kableExtra)
+library(gridExtra)
+library(broom)
 ```
 
-```{r}
+```{r,echo=F}
 #Data cleaning
@@
 ```
+## Task 1:
 
-```{r}
+```{r,echo=F}
+#Histogram to examine the expenditure
+hist(expend.df$expenditure,breaks = 20,
+     xlab = "Total weekly household expenditure (GBP)",
+     main = "Histogram of expenditure")
+
+#Creating the 
+#scatter plot to look at relationship between income and expenditure
+p1<-ggplot(data=expend.df,aes(x=income,y=expenditure))+
+  geom_point(size=0.2)+
+  xlab("Gross weekly average household income (GBP)")+
+  ylab("Total expenditure")+
+  theme(axis.text=element_text(size=6),
+        axis.title=element_text(size=6,face="bold"))
+
+#boxplots
+p2<-ggplot(data=expend.df,aes(y=expenditure,x=house.ten,fill=house.ten))+
+  geom_boxplot()+
+  xlab("Household tenure")+
+  guides(fill=F)
+
+#scale_fill_discrete(name="Household\ntenure")
+
+p3<-ggplot(data=expend.df,aes(y=expenditure,x=sex.hh,fill=sex.hh))+
+  geom_boxplot()+
+  xlab("Sex of the household head")+
+  guides(fill=F)
+
+p4<-ggplot(data=expend.df,aes(y=expenditure,x=lab.force,fill=lab.force))+
+  geom_boxplot()+
+  xlab("Employment status")+
+  guides(fill=F)
+
+p5<-ggplot(data=expend.df,aes(y=expenditure,x=hh.size,fill=hh.size))+
+  geom_boxplot()+
+  xlab("Household size")+
+  guides(fill=F)
+
+p6<-ggplot(data=expend.df,aes(y=expenditure,x=hh.adults,fill=hh.adults))+
+  geom_boxplot()+
+  xlab("Number of adults in the household")+
+  guides(fill=F)
+
+grid.arrange(p1,p2,p3,p4,p5,p6,ncol=2,nrow=3)
+```
+
+Looking at the above histogram of expenditure it shows that expenditure has a positively skewed distribution.
+Noticing this after applying a log transform the histogram of log expenditure has a more normal looking distribution.
+
+Investigating covarits against expenditure we can see that there is a positive association between household income and expenditure.
+Looking at the various factor variables using box plots...
+
+```{r,echo=F,eval=F}
+pairs(expend.df)
+xtabs(~expend.df$house.ten+expend.df$sex.hh)
+```
+
+```{r,,echo=F,eval=F}
+par(mfrow=c(1,2))
 hist(expend.df$expenditure,breaks = 20,
      xlab = "Total weekly household expenditure (GBP)")
+hist(log(expend.df$expenditure),breaks = 20,
+     xlab = " Log of total weekly \n household expenditure (GBP)")
+#Applying log transform gives a more normal looking distribution
+```
 
-ggplot(data=expend.df,aes(x=income,y=expenditure))+
-  geom_point()
+#### Part 2:
+
+```{r,echo=F}
+#Question 2
+lm.income<-lm(expenditure~income,data = expend.df)
+#summary(lm.income)
+as.data.frame(tidy(lm.income)[,1:3])%>%
+  kable(escape = F)%>%
+  kable_styling(font_size = 15)
+par(mfrow=c(2,2))
+plot(lm.income)
+# #as expected the normal QQ plot shows 'bad fit'
+
+# #try gamma as good for postively skewed continous data
+# gamma.income <-glm(expenditure~income,data=expend.df,family="Gamma")
+# summary(gamma.income)
+# plot(gamma.income)
+# #bad fit (explain)
+# 
+# #out of all models log linear is the most appropriate
 ```
-Looking at the above histogram it appears to have a skewed distribution
+Fitting a linear model $y_{i}=\beta_{0}+\beta_{1}x_{i}+\epsilon$ with the response $y$ the expenditure,the explanatory variable $x$ the income and $i \in \{1,...,1200\}$.
+The above table shows the estimated values of the intercept ($\beta_0$) and the estimated 'effect' of a 1 pound increase in income on expenditure $\beta_1$ and there standard errors.
+
+Viewing the diagnostic plots for this model.
+The fitted vs residual plot shows the residuals are not equally spread around 0 with a greater spread of residuals above zero, meaning the assumption of linearity may not hold.
+Furthermore, the QQ normal plot has a u shape with the tails diverging heavily away from the straight line.
+This implies the normality assumption is invalid.
+
+#### Part 3:
+```{r,echo=F}
+#Question 3
+lm.income.sqd <- lm(expenditure~income+I(income^2),data = expend.df)
+tidy(lm.income.sqd)[,1:3]%>%
+  kable(escape = F)%>%
+  kable_styling(font_size = 15)
+par(mfrow=c(2,2))
+plot(lm.income.sqd)
+
+# lm.log.income.sqd <- lm(log(expenditure)~income+I(income^2),data = expend.df)
+# summary(lm.log.income.sqd)
+# plot(lm.log.income.sqd)
+# 
+# gamma.income.sqd <-glm(expenditure~income+I(income^2),data=expend.df,family="Gamma")
+# summary(gamma.income.sqd)
+# plot(gamma.income.sqd)
+```
+Fitting a new linear model $y_{i}=\beta_{0}+\beta_{1}x_{i}+\beta_2x^2_{i}+\epsilon$ gives the above estimated coefficients.
+Looking at the diagnostic plots, in the fitted vs residual plot the residual values increase as the fitted values increase thus the homoscedasticity assumption may not be valid.
+The tails in the normal QQ plot also still deviate heavily away from the straight line meaning the normality assumption is invalid.
+
+#### Part 4:
+```{r,echo=F}
+#Question 4
+lm.log.income<-lm(log(expenditure)~income,data = expend.df)
+tidy(lm.log.income)[,1:3]%>%
+  kable(escape = F)%>%
+  kable_styling(font_size = 15)
+par(mfrow=c(2,2))
+plot(lm.log.income)
+# #QQ plot much better but not sure about the residual vs fitted if the variance decreases as fitted value increases
+```
+Fitting the model $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\epsilon$.
+The data appears to be hetroscedastic as the fitted vs residual plot points spread decreases as fitted values increases/ not.
+The QQ plot shows the residuals stay close to the straight line thus normality assumption is valid.
+
+#### Part 5:
+```{r,echo=F}
+lm.log.income.sqd <- lm(log(expenditure)~income+I(income^2),data = expend.df)
+summary(lm.log.income.sqd)
+par(mfrow=c(2,2))
+plot(lm.log.income.sqd)
+```
+Fitting the model $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\beta_{2}x^2_{i}+\epsilon$

### 2022-10-29 Draft_1.Rmd
Updates to model fitting (work on train)

@@
 library(broom)
+library(MASS)
 ```
@@
 ```
+```{r}
+p6
+```
 
@@
 
-Investigating covarits against expenditure we can see that there is a positive association between household income and expenditure.
-Looking at the various factor variables using box plots...
+Investigating covariates against expenditure we can see that there is a positive association between household income and expenditure.
+For the factor variables I use box plots to examine any differences present between the different levels expenditure.
+Household tenure displays differences that public rented households have noticeably lower expenditure compared to the other 2 household levels
+Female household leads expend less on average
+Employment status appears to be associated with expenditure with unemployed having a dramicatlly lower expenditure compared to full and part time employment categories
+Household size appears to be positively correlated with expenditure, with larger households having higher expenditure except for 5+ person households which goes against this trend (further investigation?)
+Number of adults in the household also appears positively correlated with households containing more adults having a higher expenditure
 
 ```{r,echo=F,eval=F}
+#extra code that may not use
 pairs(expend.df)
 xtabs(~expend.df$house.ten+expend.df$sex.hh)
-```
 
-```{r,,echo=F,eval=F}
 par(mfrow=c(1,2))
@@
 lm.log.income.sqd <- lm(log(expenditure)~income+I(income^2),data = expend.df)
-summary(lm.log.income.sqd)
+tidy(lm.log.income.sqd)[,1:3]%>%
+  kable(escape = F)%>%
+  kable_styling(font_size = 15)
 par(mfrow=c(2,2))
@@
 ```
-Fitting the model $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\beta_{2}x^2_{i}+\epsilon$
+Fitting the model $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\beta_{2}x^2_{i}+\epsilon$ gives the estimated coefficients above.
+Viewing the diagnostic plots the fitted vs residuals plot shows a fairly even scattering of points around 0 thus the linearity assumption appears valid.
+The QQ plot shows the points stay close to the straight line meaning the normality assumption is valid.
+
+#### Task 6
+```{r}
+#Comparing AIC
+AIC(lm.income)
+AIC(lm.income.sqd)
+AIC(lm.log.income)
+AIC(lm.log.income.sqd)
+#Comparing R squared
+glance(lm.income)$r.squared
+glance(lm.income.sqd)$r.squared
+glance(lm.log.income)$r.squared
+glance(lm.log.income.sqd)$r.squared
+#Comparing adjusted R squared
+glance(lm.income)$adj.r.squared
+glance(lm.income.sqd)$adj.r.squared
+glance(lm.log.income)$adj.r.squared
+glance(lm.log.income.sqd)$adj.r.squared
+```
+
+Using AIC as a measure of fit the best is model 3.
+However model 4 has the highest $R^2$ and adjusted $R^2$ suggestiong that model 4 is the best fit to the data.
+Which is further backed up by the diagnostic plots giving the indication that this is the only model in which the linearity and normality assumptions are met.
+Due to the diagnostic plots and $R^2$ values I choose model 4 as my preferred model.
+
+```{r}
+tidy(lm.log.income)
+tidy(lm.log.income.sqd)
+```
+
+Using this model a person who has no income is expected to spend £4.71 per week (add in CI) and a £100 increase in income per week will increase expenditure on average by $exp(3.12x-0.00145x^2)$
+
+MODEL IS LOG SO NEED TO EXP IT!!!!
+
+#### Task 7
+```{r}
+#Investigating possible interactions
+ggplot(data=expend.df,aes(x=sex.hh,fill=hh.adults,y=expenditure))+geom_boxplot()
+```
+
+```{r}
+#Creating model with all possible 2 way interactions (not a good idea but just for interest :P)
+lm.log.full<-lm(log(expenditure)~income+I(income^2)+income*sex.hh+income*hh.adults+income*house.ten+income*lab.force+income*hh.size+sex.hh*hh.adults+sex.hh*house.ten+sex.hh*lab.force+sex.hh*hh.size+hh.adults*house.ten+hh.adults*lab.force+hh.adults*hh.size+house.ten*lab.force+house.ten*hh.size+lab.force*hh.size,data=expend.df)
+summary(lm.log.full)
+
+step.method<- step(lm.log.full,trace=F)
+summary(step.method)
+
+```
+
+```{r}
+#Using AIC as the model fitting criteria (step function)
+#Note that we are allowed to use interactions!
+lm.log.full<-lm(log(expenditure)~income+I(income^2)+income*sex.hh+income*hh.adults+income*house.ten+lab.force+hh.size,data=expend.df)
+summary(lm.log.full)
+step.method<- step(lm.log.full,trace=F)
+summary(step.method)
+AIC(step.method)
+plot(step.method)
+```
+
+I started by only considering 2 way interactions that could be plausible for example income may have a different effect on expenditure depending on sex of household head.
+Furthermore the number of adults or household tenure type may impact the effect of income on expenditure as....
+above I investigated if there were differences when sub categorising the categories to see if there where plausible interactions between factor variables, however none showed any significant differences between the different cohorts (initial factors)
+
+#### Model selection
+For the model selection process I first wanted to decide which transformation is best to apply to the response as it is clearly skewed. using the box
+
+For this model selection process. I first decided to only consider 2 way interactions as higher order interactions make interpretation of the model much more complicated.
+When considering which interactions to include I first created plots to examine if there are any clear interactions between income and the other covariates.
+Considering the factor covarites thinking about the possible interactions none made much logical sense e.g.
+a household female head would have a different effect on expenditure to male depending on the number of people in the household.
+Thus didn't include any factor:factor ionteractions in the model.
+
+```{r}
+#interaction between sex and income
+interaction.continous <-function(var){
+  ggplot(data=expend.df,aes(x=income,y=expenditure,colour=.data[[var]]))+
+    geom_point()
+}
+
+int1<-interaction.continous("house.ten")
+int2<-interaction.continous("sex.hh")
+int3<-interaction.continous("lab.force")
+int4<-interaction.continous("hh.size")
+int5<-interaction.continous("hh.adults")
+
+grid.arrange(int1,int2,int3,int4,int5,ncol=2,nrow=3)
+int1
+int2
+int3
+int4
+int5
+```
+Looking at the scatter plots each factor for the plots does appear to have a different gradient thus an interaction could be present.
+Therefore for the model selection process I will start with a full model with all covariates and the pairwise interactions between income and each of the factor variables.
+Then using AIC I remove the coefficient that gives the biggest increase in AIC until no increase in AIC can be attained.
+Using this step method I will come to my first potential final model.
+
+```{r}
+lm.log.full<-lm(log(expenditure)~income+I(income^2)+income*sex.hh+income*hh.adults+income*house.ten+lab.force*income+hh.size*income,data=expend.df)
+summary(lm.log.full)
+step.method<- step(lm.log.full,trace=F)
+summary(step.method)
+AIC(step.method)
+plot(step.method)
+
+par(mfrow=c(1,2))
+hist(sqrt(expend.df$expenditure),breaks = 20)
+hist(log(expend.df$expenditure),breaks = 20)
+
+lm.sqrt<-lm(sqrt(expenditure)~income+I(income^2),data = expend.df)
+plot(lm.sqrt)
+```
+
+```{r}
+lm.log.full.no.int<-lm(log(expenditure)~income+I(income^2)+sex.hh+hh.adults+house.ten+lab.force+hh.size,data=expend.df)
+summary(lm.log.full.no.int)
+step.method<- step(lm.log.full.no.int,trace=F)
+
+plot(lm.log.full.no.int)
+```
+
+Using AIC as the selection critera the best model without interactions is the full model.
+
+Now I try adding interactions that couold make logical sense.
+
+As a sex of the household head could impact the effect of income on expenditure (perhaps males are more likely to spend then save then females as they get more money or the reverse)
+
+The number of adults my affect the affect of income as more adults from a higher income household may be more likely to spend more
+
+The type of household tenure may affect the
+
+```{r}
+#adding these into the model 1 by 1
+new.lm.add.sex<-update(lm.log.full.no.int, .~.+income*sex.hh)
+anova(lm.log.full.no.int,new.lm.add.sex)
+summary(new.lm.add.sex)
+
+AIC(lm.log.full.no.int)
+AIC(new.lm.add.sex)
+
+glance(lm.log.full.no.int)$adj.r.squared
+glance(new.lm.add.sex)$adj.r.squared
+#Not really much difference and note that interaction is not significant
+
+new.lm.add.adults<-update(lm.log.full.no.int, .~.+income*hh.adults)
+anova(lm.log.full.no.int,new.lm.add.adults)
+summary(new.lm.add.adults)
+
+AIC(lm.log.full.no.int)
+AIC(new.lm.add.adults)
+
+glance(lm.log.full.no.int)$adj.r.squared
+glance(new.lm.add.adults)$adj.r.squared
+
+new.lm.add.ten<-update(lm.log.full.no.int, .~.+income*house.ten)
+anova(lm.log.full.no.int,new.lm.add.ten)
+summary(new.lm.add.ten)
+
+AIC(lm.log.full.no.int)
+AIC(new.lm.add.ten)
+
+glance(lm.log.full.no.int)$adj.r.squared
+glance(new.lm.add.ten)$adj.r.squared
+
+AIC(lm.log.full)
+AIC(lm.log.full.no.int)
+glance(lm.log.full.no.int)$adj.r.squared
+glance(lm.log.full)$adj.r.squared
+```
+
+```{r}
+# Attempting Box-Cox transformation to check whether distribution is better
+estim_BoxCox <- boxcox(expenditure ~ income+sex.hh+hh.adults+house.ten+lab.force+hh.size+income:hh.adults+income:house.ten+income:lab.force,data =expend.df)
+# Find optimal lambda
+lambda <- estim_BoxCox$x[ which.max( estim_BoxCox$y ) ]
+```

### 2022-10-29 Draft_1.Rmd
Removing useless code chunks and adding in interpretation of chosen basic model

@@
 
+#### Part 1:
 ```{r,echo=F}
@@
 ```
-```{r}
-p6
-```
 
-Looking at the above histogram of expenditure it shows that expenditure has a positively skewed distribution.
-Noticing this after applying a log transform the histogram of log expenditure has a more normal looking distribution.
+Looking at the above histogram of expenditure above it shows that expenditure has a positively skewed distribution.
 
-Investigating covariates against expenditure we can see that there is a positive association between household income and expenditure.
-For the factor variables I use box plots to examine any differences present between the different levels expenditure.
-Household tenure displays differences that public rented households have noticeably lower expenditure compared to the other 2 household levels
-Female household leads expend less on average
-Employment status appears to be associated with expenditure with unemployed having a dramicatlly lower expenditure compared to full and part time employment categories
+I investigated the relationships between the other variable and expenditure dataset using the plots above.
+We can see that there is a positive association between household income and expenditure in the scatter plot.
+For the factor variables I used box plots to examine any differences present between the different levels within the factor variable and expenditure:
+Different household tenure categories have different expenditure amounts with public rented households having a noticeably lower expenditure compared to the other 2 household levels
+Households with a female head appear to have a lower expenditure than that of male household heads
+Employment status appears to be associated with expenditure with unemployed having a dramatically lower expenditure compared to full and part time employment categories.
 Household size appears to be positively correlated with expenditure, with larger households having higher expenditure except for 5+ person households which goes against this trend (further investigation?)
-Number of adults in the household also appears positively correlated with households containing more adults having a higher expenditure
-
-```{r,echo=F,eval=F}
-#extra code that may not use
-pairs(expend.df)
-xtabs(~expend.df$house.ten+expend.df$sex.hh)
-
-par(mfrow=c(1,2))
-hist(expend.df$expenditure,breaks = 20,
-     xlab = "Total weekly household expenditure (GBP)")
-hist(log(expend.df$expenditure),breaks = 20,
-     xlab = " Log of total weekly \n household expenditure (GBP)")
-#Applying log transform gives a more normal looking distribution
-```
+Number of adults in the household also appears positively correlated with households containing more adults having a higher expenditure.
 
@@
 plot(lm.income)
-# #as expected the normal QQ plot shows 'bad fit'
-
-# #try gamma as good for postively skewed continous data
-# gamma.income <-glm(expenditure~income,data=expend.df,family="Gamma")
-# summary(gamma.income)
-# plot(gamma.income)
-# #bad fit (explain)
-# 
-# #out of all models log linear is the most appropriate
 ```
-Fitting a linear model $y_{i}=\beta_{0}+\beta_{1}x_{i}+\epsilon$ with the response $y$ the expenditure,the explanatory variable $x$ the income and $i \in \{1,...,1200\}$.
+Fitting a linear model $y_{i}=\beta_{0}+\beta_{1}x_{i}+\epsilon$ with the response $y$ the expenditure,the explanatory variable $x$ income and $i \in \{1,...,1200\}$ the ith household.
 The above table shows the estimated values of the intercept ($\beta_0$) and the estimated 'effect' of a 1 pound increase in income on expenditure $\beta_1$ and there standard errors.
@@
 The QQ plot shows the points stay close to the straight line meaning the normality assumption is valid.
+(scale location is slightly dogy- sign constant variance assumption has been violated)
 
@@
 
-Using AIC as a measure of fit the best is model 3.
-However model 4 has the highest $R^2$ and adjusted $R^2$ suggestiong that model 4 is the best fit to the data.
-Which is further backed up by the diagnostic plots giving the indication that this is the only model in which the linearity and normality assumptions are met.
-Due to the diagnostic plots and $R^2$ values I choose model 4 as my preferred model.
+Using AIC as a measure of fit the best is model 4 as it has the lowest AIC value.
+Furthermore, model 4 has the highest $R^2$ and adjusted $R^2$ suggestion that model 4 'fits' the data best.
+This is further backed up by the diagnostic plots which indicate that this is the only model in which the linearity and normality assumptions are met.
+Due to all these reasons model 4 as my preferred model.
 
 ```{r}
-tidy(lm.log.income)
 tidy(lm.log.income.sqd)
+
+mean(expend.df$income)
 ```
+Using the model a 1 unit increase in household income increase the expenditure by $exp(\beta_1+2x\beta_2+\beta_2)-1$x100% of the old value.
+Subbing in the values of $\beta_1$ and $\beta_2$ this gives an increase by $exp(0.0032-0.0000029x-0.0000015)-1$x100%.
+Therefore for a household earning the mean income of £512 a £1 increase in income increases expenditure by approximately 0.17%.
+Using a more likely realistic increase in weekly income of £20 for a household earning £512 gives an increase of $exp(0.016-0.000015*20-0.000036)-1$x100% = 1.6%.
 
-Using this model a person who has no income is expected to spend £4.71 per week (add in CI) and a £100 increase in income per week will increase expenditure on average by $exp(3.12x-0.00145x^2)$
-
-MODEL IS LOG SO NEED TO EXP IT!!!!
+Include CI?
 
 #### Task 7
-```{r}
-#Investigating possible interactions
-ggplot(data=expend.df,aes(x=sex.hh,fill=hh.adults,y=expenditure))+geom_boxplot()
-```
-
-```{r}
-#Creating model with all possible 2 way interactions (not a good idea but just for interest :P)
-lm.log.full<-lm(log(expenditure)~income+I(income^2)+income*sex.hh+income*hh.adults+income*house.ten+income*lab.force+income*hh.size+sex.hh*hh.adults+sex.hh*house.ten+sex.hh*lab.force+sex.hh*hh.size+hh.adults*house.ten+hh.adults*lab.force+hh.adults*hh.size+house.ten*lab.force+house.ten*hh.size+lab.force*hh.size,data=expend.df)
-summary(lm.log.full)
-
-step.method<- step(lm.log.full,trace=F)
-summary(step.method)
-
-```
-
-```{r}
-#Using AIC as the model fitting criteria (step function)
-#Note that we are allowed to use interactions!
-lm.log.full<-lm(log(expenditure)~income+I(income^2)+income*sex.hh+income*hh.adults+income*house.ten+lab.force+hh.size,data=expend.df)
-summary(lm.log.full)
-step.method<- step(lm.log.full,trace=F)
-summary(step.method)
-AIC(step.method)
-plot(step.method)
-```
 
@@
 #### Model selection
-For the model selection process I first wanted to decide which transformation is best to apply to the response as it is clearly skewed. using the box
-
-For this model selection process. I first decided to only consider 2 way interactions as higher order interactions make interpretation of the model much more complicated.
+For this model selection process I decided to only consider 2 way interactions as higher order interactions make interpretation of the model much more complicated.
 When considering which interactions to include I first created plots to examine if there are any clear interactions between income and the other covariates.
 Considering the factor covarites thinking about the possible interactions none made much logical sense e.g.
-a household female head would have a different effect on expenditure to male depending on the number of people in the household.
-Thus didn't include any factor:factor ionteractions in the model.
+a household female head would have a different effect on expenditure to a male head depending on the number of people in the household.
+Thus didn't include any factor:factor interactions in the model.
 
@@
   ggplot(data=expend.df,aes(x=income,y=expenditure,colour=.data[[var]]))+
-    geom_point()
+    geom_point(size=0.1)
 }
@@
 lm.log.full<-lm(log(expenditure)~income+I(income^2)+income*sex.hh+income*hh.adults+income*house.ten+lab.force*income+hh.size*income,data=expend.df)
-summary(lm.log.full)
 step.method<- step(lm.log.full,trace=F)
-summary(step.method)
+tidy(step.method)
 AIC(step.method)
+glance(step.method)$adj.r.squared
 plot(step.method)
-
-par(mfrow=c(1,2))
-hist(sqrt(expend.df$expenditure),breaks = 20)
-hist(log(expend.df$expenditure),breaks = 20)
-
-lm.sqrt<-lm(sqrt(expenditure)~income+I(income^2),data = expend.df)
-plot(lm.sqrt)
+#got rid of 2 interactions: income:sex and income:hhsize
 ```

### 2022-10-30 Draft_1.Rmd
Rewording early tasks

@@
 Looking at the above histogram of expenditure above it shows that expenditure has a positively skewed distribution.
+Investigating the relationship between household income and expenditure in the scatter plot above, it displays a clear positive association.
+Then to investigate the other variables which are categorical I used boxplots to display differences in the expenditure across the different levels(categories).
+The differences shown in the plots are listed below:
 
-I investigated the relationships between the other variable and expenditure dataset using the plots above.
-We can see that there is a positive association between household income and expenditure in the scatter plot.
-For the factor variables I used box plots to examine any differences present between the different levels within the factor variable and expenditure:
-Different household tenure categories have different expenditure amounts with public rented households having a noticeably lower expenditure compared to the other 2 household levels
-Households with a female head appear to have a lower expenditure than that of male household heads
-Employment status appears to be associated with expenditure with unemployed having a dramatically lower expenditure compared to full and part time employment categories.
-Household size appears to be positively correlated with expenditure, with larger households having higher expenditure except for 5+ person households which goes against this trend (further investigation?)
-Number of adults in the household also appears positively correlated with households containing more adults having a higher expenditure.
+-Different household tenures categories have different weekly expenditure amounts with public rented households having a noticeably lower expenditure compared to the other 2 household levels.
+-Households with a female head have a lower weekly expenditure compared to male household heads
+-Employment status appears to be associated with expenditure with unemployed households having a dramatically lower expenditure compared to households with full or part time employment.
+-Household size generally appears to be positively correlated with expenditure, with larger households having a higher weekly expenditure (however 5+ households break this trend with the expenditure decreasing, this could be due to other variables).
+Number of adults in the household also appears positively correlated with expenditure with households containing more adults having a higher expenditure.
+
+I note that these differences may not be due to the variable mentioned but due to a third variable (a confounder) that affects both the categorical variable and expenditure (the response) which could be other variables in the dataset or ones that were not even included.
 
 #### Part 2:
+In all linear models mentioned for part 2 to 6, the response $y$ is the weekly household expenditure and the explanatory variable $x$ is weekly income and $i \in \{1,...,1200\}$ is the ith household in the dataset.
 
@@
 ```
-Fitting a linear model $y_{i}=\beta_{0}+\beta_{1}x_{i}+\epsilon$ with the response $y$ the expenditure,the explanatory variable $x$ income and $i \in \{1,...,1200\}$ the ith household.
-The above table shows the estimated values of the intercept ($\beta_0$) and the estimated 'effect' of a 1 pound increase in income on expenditure $\beta_1$ and there standard errors.
+The estimated coefficients for the fitted linear model $y_{i}=\beta_{0}+\beta_{1}x_{i}+\epsilon$  are displayed in the above table.
+It shows the estimated values of the intercept ($\beta_0$) and the estimated 'effect' of a 1 pound increase in income on expenditure $\beta_1$ and the associated standard errors.
 
@@
 ```
-Fitting a new linear model $y_{i}=\beta_{0}+\beta_{1}x_{i}+\beta_2x^2_{i}+\epsilon$ gives the above estimated coefficients.
+Fitting a new linear model $y_{i}=\beta_{0}+\beta_{1}x_{i}+\beta_2x^2_{i}+\epsilon$ with an added squared term of income gives the above estimated coefficients.
 Looking at the diagnostic plots, in the fitted vs residual plot the residual values increase as the fitted values increase thus the homoscedasticity assumption may not be valid.
@@
 ```
-Fitting the model $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\epsilon$.
-The data appears to be hetroscedastic as the fitted vs residual plot points spread decreases as fitted values increases/ not.
+Fitting the same model as in task 2 but with the response of expenditure logged $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\epsilon$ gives the above coefficients.
+The data appears to be heteroscedastic as the fitted vs residual plot points spread decreases as fitted values increases, furthermore the linear assumption also appears violated as the residuals appear to exhibit a slight negative quadratic shape.
 The QQ plot shows the residuals stay close to the straight line thus normality assumption is valid.
@@
 Fitting the model $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\beta_{2}x^2_{i}+\epsilon$ gives the estimated coefficients above.
-Viewing the diagnostic plots the fitted vs residuals plot shows a fairly even scattering of points around 0 thus the linearity assumption appears valid.
+Viewing the diagnostic plots the fitted vs residuals plot shows a fairly even scattering of points around 0 and a linear trend, however the residuals are bunched to the very right of the plot and the variance of the residuals appears to decrease as the fitted values increase, this could be to important variables being left out of the model (MAY NEED TO CHANGE THIS).
+The non-constant variance assumption being violated is further backed up by the none straight line in the scale location plot.
 The QQ plot shows the points stay close to the straight line meaning the normality assumption is valid.
-(scale location is slightly dogy- sign constant variance assumption has been violated)
 
@@
 
-Using AIC as a measure of fit the best is model 4 as it has the lowest AIC value.
-Furthermore, model 4 has the highest $R^2$ and adjusted $R^2$ suggestion that model 4 'fits' the data best.
-This is further backed up by the diagnostic plots which indicate that this is the only model in which the linearity and normality assumptions are met.
-Due to all these reasons model 4 as my preferred model.
+The 4th model $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\beta_{2}x^2_{i}+\epsilon$ has the lowest AIC value of *1528.062* and the highest adjusted R squared value of *0.4635055*.
+Furthermore, looking at the diagnostic plots the normality assumption is valid and the fitted vs residual plot although not 'perfect' displays a more even scattering of points than that of the same plot for the other models.
+Therefore for all these reasons the 4th model is my preferred model to describe the relationship between expenditure and income.
 
@@
 ```
-Using the model a 1 unit increase in household income increase the expenditure by $exp(\beta_1+2x\beta_2+\beta_2)-1$x100% of the old value.
-Subbing in the values of $\beta_1$ and $\beta_2$ this gives an increase by $exp(0.0032-0.0000029x-0.0000015)-1$x100%.
+Income has a positive association with expenditure with a 1 pound increase in household income increasing household expenditure by $exp(0.0032-0.0000029x-0.0000015)-1$x100%.
 Therefore for a household earning the mean income of £512 a £1 increase in income increases expenditure by approximately 0.17%.
 Using a more likely realistic increase in weekly income of £20 for a household earning £512 gives an increase of $exp(0.016-0.000015*20-0.000036)-1$x100% = 1.6%.
-
-Include CI?
-
 #### Task 7
-
-I started by only considering 2 way interactions that could be plausible for example income may have a different effect on expenditure depending on sex of household head.
+For interactions I decided to only consider 2 way interactions that could be plausible for example income may have a different effect on expenditure depending on sex of household head.
 Furthermore the number of adults or household tenure type may impact the effect of income on expenditure as....
@@
 For this model selection process I decided to only consider 2 way interactions as higher order interactions make interpretation of the model much more complicated.
-When considering which interactions to include I first created plots to examine if there are any clear interactions between income and the other covariates.
-Considering the factor covarites thinking about the possible interactions none made much logical sense e.g.
+Furthermore, when considering which interactions to include I first created plots to examine if there are any clear interactions between income and the other covariates.
+Considering plausible interactions between 2 factor covarites an interaction between household size and number of adults made sense?
+and labour force with either household size or number of adults in the household.
+
+household size and number of adults (number of adults may change effect of household size)
+lab.force and hh.size or hh.adult as families who are unemployed may have different spending habits be more frugal then a singluar household that is unemployed 
+ INVESTIGATE HOW SIMILAR EACH VARIABLE IS!
+none made much logical sense e.g.
 a household female head would have a different effect on expenditure to a male head depending on the number of people in the household.
@@
 glance(step.method)$adj.r.squared
+
+#got rid of 2 interactions: income:sex and income:hhsize
+AIC(lm.log.income.sqd)
+AIC(step.method)
+glance(lm.log.income.sqd)$adj.r.squared
+glance(step.method)$adj.r.squared
+
+summary(step.method)
+
+anova(step.method)
+```
+
+Using AIC as the model selection criteria I removed 2 interactions income:sex and income:household size.
+Comparing this new model to the original the AIC is significantly lower and the adjusted R squarded is significantly higher.
+Also note that each variable is significant except for
+
+```{r}
 plot(step.method)
-#got rid of 2 interactions: income:sex and income:hhsize
 ```

### 2022-11-02 Task_2_draft.Rmd
Created 1st version of function

---
title: "Untitled"
output: html_document
date: "2022-11-02"
---
```{r}
#Setting up enviroment
setwd("~/Desktop/GLM/Coursework/GLM-CW")
airline.df <- read.table("airline.txt",header=T)

#Setting up inputs of mle function
n <- dim(airline.df)[[1]]
con <- rep(1,n)
X=matrix(c(con,airline.df$x),nrow=n,ncol=2)
y <- airline.df$y
```

```{r}
#Creating function for algorithm
Fisher.scoring.alg <- function(X,y){
  #Setting initial guess of beta coefficients
  beta.new <- matrix(c(1,1),nrow=2,ncol=1)
  #Setting objects for conditions of when stop iterations (while loop)
  flag <- 1
  eps <- 0.00001
  c <- 0
  max.iter <- 100000000

  while (flag==1){
    #Updating current value of beta to use for current iteration to find next beta value
    beta <- beta.new

    #Updating Y where YX is the score function
    Y <- as.vector((y*exp(-(X%*%beta)))-1)

    #Applying Fisher Scoring algorithm to find updated value of beta 
    beta.new <- beta + solve(t(X)%*%X)%*%t(Y%*%X)

    #Checking for convergence
    if (abs(sum(beta.new)-sum(beta))<eps){
      flag <- 0
    }
    c <- c+1
  }
  #Variance covariance matrix
  var.cov.mat <- solve(t(X)%*%X)

  return("estimates"=beta.new)
}

Fisher.scoring.alg(X,y)
```

```{r}
#Check with glm model
fit <- glm(y~X[,2], family  = Gamma(link="log"))
summary(fit,dispersion=1)  
```

### 2022-11-02 Tutorial_code.Rmd
Tutorial code

---
title: "Untitled"
output: html_document
date: "2022-11-02"
---

```{r}
#setting up variables
n=1000
x=rnorm(n,0,1)
p=exp(0.1+0.5*x)/(1+exp(0.1+0.5*x))
y=rbinom(n,1,p)
con=rep(1,n)
X=matrix(c(con,x),nrow=n,ncol=2)

# #Fiddling around to work out the matrix algebra
# T <- matrix(c(1,2,3,4),ncol=2,nrow=2,byrow=T)
# beta <- matrix(c(1,1),ncol=1,nrow=2)
# beta
# diag(as.vector(exp(T%*%beta)/(1+exp(T%*%beta))^2),2)
#   
# vec <- c(1,0)-as.vector(exp(T%*%beta)/(1+exp(T%*%beta)))
# t(vec%*%T)
```

```{r}
# First make with iteration to be specified to check function is working
MLE.iter <- function(y,X,iter){

  beta.previous <- matrix(c(1,1),nrow=2,ncol=1)

  for (i in 1:iter){
    beta <- beta.previous
    W <- diag(as.vector(exp(X%*%beta)/(1+exp(X%*%beta))^2),dim(X)[1])
    Y <- y-as.vector(exp(X%*%beta)/(1+exp(X%*%beta)))

    beta.previous<- beta+solve(t(X)%*%W%*%X)%*%t(Y%*%X)
  }
  return(beta.previous)
}

#Using convergence this time to stop iterations (or if reaches max iterations)
MLE.conv <- function(y,X){
  #Setting initial guess of beta
  beta.new <- matrix(c(1,1),nrow=2,ncol=1)
  #Setting objects for conditions of when stop iterations (while loop)
  flag <- 1
  eps <- 0.00001
  c <- 0
  max.iter <- 100

  while (flag==1){
    #Setting beta to use to find new value of beta
    beta <- beta.new

    #Finding W, where I have defined W as t(X)WX = Fisher information matrix
    W <- diag(as.vector(exp(X%*%beta)/(1+exp(X%*%beta))^2),dim(X)[1])
    #Finding Y where t(YX) is defined as the score 
    Y <- y-as.vector(exp(X%*%beta)/(1+exp(X%*%beta)))

    #Applying Fisher scoring
    beta.new<- beta+solve(t(X)%*%W%*%X)%*%t(Y%*%X)

    #Checking for convergence
    if (abs(sum(beta.new)-sum(beta))<eps){
      flag <- 0
    }
    c <- c+1
  }

  var.cov.mat <- solve(t(X)%*%W%*%X)

  return(list("estimates"=beta.new,"vcov"=var.cov.mat))
}

MLE.output <- MLE.conv(y,X)
MLE.output$estimates

#Testing against glm function
glm.model <- glm(y~x,family = "binomial")
summary(glm.model)

#Checking variance-covariance matrix against that found in glm function
vcov(glm.model)
MLE.conv(y,X)$vcov
```

### 2022-11-04 Task_2_draft.Rmd
adding comments in code

@@
   #Setting initial guess of beta coefficients
-  beta.new <- matrix(c(1,1),nrow=2,ncol=1)
+
+  beta.new <- matrix(c(0,1),nrow=2,ncol=1)
   #Setting objects for conditions of when stop iterations (while loop)
   flag <- 1
-  eps <- 0.00001
+  #setting size of absolute difference between beta iterations when convergence is reached
+  eps <- 0.0000001
+  #Setting loop counter
   c <- 0
-  max.iter <- 100000000
+  #Max iterations before 
+  max.iter <- 1000000
 
@@
     }
+
+    if (c>max.iter){
+      stop("Max number of iterations reached: Non-convergence of beta")
+    }
     c <- c+1
@@
 
-  return("estimates"=beta.new)
+  return(beta.new)
 }
@@
 ```
+
+```{r}
+#Finding var-cov matrix
+
+t(X)%*%(X)
+```

### 2022-11-05 Draft_1.Rmd
Working on model selection process

@@
 #### Model selection
-For this model selection process I decided to only consider 2 way interactions as higher order interactions make interpretation of the model much more complicated.
-Furthermore, when considering which interactions to include I first created plots to examine if there are any clear interactions between income and the other covariates.
-Considering plausible interactions between 2 factor covarites an interaction between household size and number of adults made sense?
-and labour force with either household size or number of adults in the household.
-
-household size and number of adults (number of adults may change effect of household size)
-lab.force and hh.size or hh.adult as families who are unemployed may have different spending habits be more frugal then a singluar household that is unemployed 
- INVESTIGATE HOW SIMILAR EACH VARIABLE IS!
-none made much logical sense e.g.
-a household female head would have a different effect on expenditure to a male head depending on the number of people in the household.
-Thus didn't include any factor:factor interactions in the model.
+Considering possible covariates, there are several that may be correlated with each other, number of household members and number of adult household members, house tenure and labour force.
+To evaluate if these variables are correlated I will use Pearsons Chi squared test.
 
 ```{r}
-#interaction between sex and income
+#Creating tables for chi-squared test
+prop.table(table(expend.df$hh.size,expend.df$hh.adults),2)
+chisq.test(table(expend.df$hh.size,expend.df$hh.adults))
+
+prop.table(table(expend.df$house.ten,expend.df$lab.force),2)
+chisq.test(table(expend.df$house.ten,expend.df$lab.force))
+```
+
+As evidence supports that both of these pairs are correlated (and makes logical sense) I will the model with each included and select the variable out of the pair that has the best fit as defined by AIC and R^2.
+
+```{r}
+#Checking which of the correlated variables to include
+lm.hh.size <- lm(log(expenditure)~income+I(income^2)+hh.size, data = expend.df)
+lm.hh.adults <- lm(log(expenditure)~income+I(income^2)+hh.adults, data = expend.df)
+lm.tenure <- lm(log(expenditure)~income+I(income^2)+house.ten, data = expend.df)
+lm.labour <- lm(log(expenditure)~income+I(income^2)+lab.force, data = expend.df)
+
+AIC(lm.hh.size)
+AIC(lm.hh.adults)
+#size has lower AIC
+
+AIC(lm.tenure)
+AIC(lm.labour)
+#tenure has lower AIC
+
+glance(lm.hh.size)$r.squared
+glance(lm.hh.adults)$r.squared
+#Size has higher R^2
+
+glance(lm.tenure)$r.squared
+glance(lm.labour)$r.squared
+#Tenure has higher R^2
+```
+Based on AIC and R^2 (can use as all models have same number of variables) I select household size and tenure as the 2 variables to use in the model and won't use labour or household adults variables so will no longer consider these 2 variables.
+
+For this model selection process I decided to only consider 2 way interactions as higher order interactions make interpretation of the model much more complicated.
+Furthermore, when considering which interactions to include I first created plots to examine if there are any clear interactions between income and the other covariates (those still left).
+
+sex, tenure and household size
+
+```{r}
+#interaction between 3 factor variables and income investigation
 interaction.continous <-function(var){
   ggplot(data=expend.df,aes(x=income,y=expenditure,colour=.data[[var]]))+
-    geom_point(size=0.1)
+    geom_point()+
+  geom_smooth(method='lm', formula= y~x,se=F)
 }
@@
 int2<-interaction.continous("sex.hh")
-int3<-interaction.continous("lab.force")
-int4<-interaction.continous("hh.size")
-int5<-interaction.continous("hh.adults")
+int3<-interaction.continous("hh.size")
 
-grid.arrange(int1,int2,int3,int4,int5,ncol=2,nrow=3)
+grid.arrange(int1,int2,int3,ncol=2,nrow=2)
 int1
@@
 int3
-int4
-int5
 ```
-Looking at the scatter plots each factor for the plots does appear to have a different gradient thus an interaction could be present.
+
+Looking at the scatter plots each factor for the plots does appear to have a different gradients.
+Plotting a linear model of y~x for each level of the covariate for each plot the lines cross which gives evidence that there is an interaction between income and the other covariate.
+As the level of the covariate changes the effect of income on expenditure as the slopes are of different gradients.
 Therefore for the model selection process I will start with a full model with all covariates and the pairwise interactions between income and each of the factor variables.
-Then using AIC I remove the coefficient that gives the biggest increase in AIC until no increase in AIC can be attained.
-Using this step method I will come to my first potential final model.
+Then using AIC and adjusted R^2 I will remove the coefficients.
+(lower AIC or higher adjusted R^2).
+Once I am happy with the model I end up with I will check the diagnostic plots to make sure the model fit is good/ assumptions met.
 
 ```{r}
-lm.log.full<-lm(log(expenditure)~income+I(income^2)+income*sex.hh+income*hh.adults+income*house.ten+lab.force*income+hh.size*income,data=expend.df)
+#Test with larger model with all possible 2 way interactions
+lm.log.full<-lm(log(expenditure)~income+I(income^2)+income*sex.hh+income*house.ten+hh.size*income+sex.hh:house.ten+sex.hh:hh.size+house.ten:hh.size,data=expend.df)
+step.method<- step(lm.log.full,trace=F)
+
+summary(lm.log.full)
+#Only sex:hh.size is significant (maybe keep it and remove all the others?)
+
+tidy(step.method)
+```
+
+```{r}
+lm.log.full<-lm(log(expenditure)~income+I(income^2)+income*sex.hh+income*house.ten+hh.size*income,data=expend.df)
 step.method<- step(lm.log.full,trace=F)
 tidy(step.method)
+#sex income interaction removed
+
 AIC(step.method)
@@
 ```
-
-```{r}
-# Attempting Box-Cox transformation to check whether distribution is better
-estim_BoxCox <- boxcox(expenditure ~ income+sex.hh+hh.adults+house.ten+lab.force+hh.size+income:hh.adults+income:house.ten+income:lab.force,data =expend.df)
-# Find optimal lambda
-lambda <- estim_BoxCox$x[ which.max( estim_BoxCox$y ) ]
-```

### 2022-11-06 Draft_1.Rmd
Large changes to model selection process (now fairly happy with final model chosen)

--- 2022-10-30 Draft_1.Rmd
@@
 #Comparing AIC
+paste("AIC")
 AIC(lm.income)
@@
 AIC(lm.log.income.sqd)
-#Comparing R squared
-glance(lm.income)$r.squared
-glance(lm.income.sqd)$r.squared
-glance(lm.log.income)$r.squared
-glance(lm.log.income.sqd)$r.squared
+
+# #Comparing R squared
+# glance(lm.income)$r.squared
+# glance(lm.income.sqd)$r.squared
+# glance(lm.log.income)$r.squared
+# glance(lm.log.income.sqd)$r.squared
+
 #Comparing adjusted R squared
+paste("adj R squared")
 glance(lm.income)$adj.r.squared
@@
 #### Task 7
-For interactions I decided to only consider 2 way interactions that could be plausible for example income may have a different effect on expenditure depending on sex of household head.
-Furthermore the number of adults or household tenure type may impact the effect of income on expenditure as....
-above I investigated if there were differences when sub categorising the categories to see if there where plausible interactions between factor variables, however none showed any significant differences between the different cohorts (initial factors)
-
-#### Model selection
-For this model selection process I decided to only consider 2 way interactions as higher order interactions make interpretation of the model much more complicated.
-Furthermore, when considering which interactions to include I first created plots to examine if there are any clear interactions between income and the other covariates.
-Considering plausible interactions between 2 factor covarites an interaction between household size and number of adults made sense?
-and labour force with either household size or number of adults in the household.
-
-household size and number of adults (number of adults may change effect of household size)
-lab.force and hh.size or hh.adult as families who are unemployed may have different spending habits be more frugal then a singluar household that is unemployed 
- INVESTIGATE HOW SIMILAR EACH VARIABLE IS!
-none made much logical sense e.g.
-a household female head would have a different effect on expenditure to a male head depending on the number of people in the household.
-Thus didn't include any factor:factor interactions in the model.
+To start I investigated the multicolinearity of the variables to make sure that certain variables were not linearly dependent as this would cause issues with the model if linearly dependent variables were included.
 
 ```{r}
-#interaction between sex and income
+#Checking variance inflation factors (generalized as variables include factors)
+car::vif(lm(log(expenditure)~income+I(income^2)+hh.size+hh.adults+sex.hh+lab.force+house.ten,data=expend.df))
+```
+None of the variables had a GVIF greater than 5, thus multicolinearity is not an issue with the explanatory variables.
+Thus multicolinearity will not be an issue.
+However, considering the variables it seems likely that household size is related to number of household adults.
+
+```{r}
+#Creating tables for chi-squared test
+prop.table(table(expend.df$hh.size,expend.df$hh.adults),2)
+```
+It is clear there is a relationship between the two. I now create linear models includiong each seperately and both together.
+
+```{r}
+lm.adult <- lm(log(expenditure)~income+I(income^2)+hh.adults,data = expend.df)
+summary(lm.adult)
+#adult significant
+
+lm.size <- lm(log(expenditure)~income+I(income^2)+hh.size,data = expend.df)
+summary(lm.size)
+#size significant
+
+lm.adult.size <- lm(log(expenditure)~income+I(income^2)+hh.size+hh.adults,data = expend.df)
+summary(lm.adult.size)
+#adult no longer significant
+```
+
+```{r}
+# #Checking which of the correlated variables to include
+# lm.hh.size <- lm(log(expenditure)~income+I(income^2)+hh.size, data = expend.df)
+# lm.hh.adults <- lm(log(expenditure)~income+I(income^2)+hh.adults, data = expend.df)
+# 
+# AIC(lm.hh.size)
+# AIC(lm.hh.adults)
+# #size has lower AIC
+# 
+# glance(lm.hh.size)$r.squared
+# glance(lm.hh.adults)$r.squared
+# #Size has higher R^2
+
+#Based on AIC and R^2 (can use as all models have same number of variables) I select household size and tenure as the 2 variables to use in the model and won't use labour or household adults variables so will no longer consider these 2 variables.
+```
+Looking at the summary outputs of all 3 variables the number of adults in the household is a significant predictor when on its own for predicting expenditure however when we add size of household it is no longer significant.
+Therefore the effect of number of adults can be explained by the size fo the household thus there is no point including it in the model.
+Thus, I will not consider the number of adults in the household as an explanatory variable for the model.
+
+Considering the possible interactions, I decided to only include 2 way interactions as higher order interactions make interpretation of the model much more complicated and there is no clear 3 way interactions that make sense.
+
+First I investigated interactions between income and the factor variables.
+To do this I created scatter plots of income against expenditure with income split into each factor of the chosen explanatory factor variable with a plotted linear model for each of the level of the factor to see if the effect of income varied with each level of the factor variable on expenditure.
+This would be seen by a different gradient of the linear model.
+
+```{r}
+#interaction between 3 factor variables and income investigation
 interaction.continous <-function(var){
   ggplot(data=expend.df,aes(x=income,y=expenditure,colour=.data[[var]]))+
-    geom_point(size=0.1)
+    geom_point()+
+  geom_smooth(method='lm', formula= y~x,se=F)
 }
@@
 int2<-interaction.continous("sex.hh")
-int3<-interaction.continous("lab.force")
-int4<-interaction.continous("hh.size")
-int5<-interaction.continous("hh.adults")
+int3<-interaction.continous("hh.size")
+int4<-interaction.continous("lab.force")
 
-grid.arrange(int1,int2,int3,int4,int5,ncol=2,nrow=3)
+grid.arrange(int1,int2,int3,int4,ncol=2,nrow=2)
 int1
@@
 int4
-int5
 ```
-Looking at the scatter plots each factor for the plots does appear to have a different gradient thus an interaction could be present.
-Therefore for the model selection process I will start with a full model with all covariates and the pairwise interactions between income and each of the factor variables.
-Then using AIC I remove the coefficient that gives the biggest increase in AIC until no increase in AIC can be attained.
-Using this step method I will come to my first potential final model.
+Looking at the plots all show that factor variables have an affect on the effect of income on expenditure with different gradients between the levels, however the magnitude of the difference in these gradients varies.
+With sex the difference in gradient is very small indicating that sex has little affect on the effect of income on expenditure.
+Labour force on the hand has a large difference in gradients between its levels giving evidence of an interaction between labour force and income.
+I decided to not include a interaction between sex and income as there is little evidence to support this (and makes logical sense).
+
+Looking at the plots.
+The largest differences in gradients between levels are in income:labour force, income:household adults and income: household size plots, thus these are the most likely to have interactions with income.
+Furthermore, it makes logical sense that these variables would affect to effect of income on expenditure.
+Meanwhile sex has little effect on income with the gradients although crossing remaining very similar and house tenure type having some differences.
 
 ```{r}
-lm.log.full<-lm(log(expenditure)~income+I(income^2)+income*sex.hh+income*hh.adults+income*house.ten+lab.force*income+hh.size*income,data=expend.df)
+#Checking each of the interactions to see if significant with all variables included
+# lm.int.income.sex <- lm(log(expenditure)~income+I(income^2)+hh.size+sex.hh*income+lab.force+house.ten,data=expend.df)
+# summary(lm.int.income.sex)
+# #not significant as expected
+
+lm.int.income.size <- lm(log(expenditure)~income+I(income^2)+hh.size*income+sex.hh+lab.force+house.ten,data=expend.df)
+summary(lm.int.income.size)
+#significant
+
+#significant
+lm.int.income.labour <- lm(log(expenditure)~income+I(income^2)+hh.size+sex.hh+lab.force*income+house.ten,data=expend.df)
+summary(lm.int.income.labour)
+#significant
+
+lm.int.income.tenure <- lm(log(expenditure)~income+I(income^2)+hh.size+sex.hh+lab.force+house.ten*income,data=expend.df)
+summary(lm.int.income.tenure)
+#significant
+```
+Considering factor variable interactions. There are 6 possible factor interactions.
+
+```{r}
+# attach(expend.df)
+# prop.table(table(sex.hh,hh.size),1)
+# prop.table(table(sex.hh,hh.size),2)
+# #differences
+# 
+# prop.table(table(sex.hh,house.ten),1)
+# prop.table(table(sex.hh,house.ten),2)
+# 
+# prop.table(table(sex.hh,lab.force),1)
+# prop.table(table(sex.hh,lab.force),2)
+# #differences
+# 
+# prop.table(table(house.ten,hh.size),1)
+# prop.table(table(house.ten,hh.size),2)
+# 
+# prop.table(table(lab.force,hh.size),1)
+# prop.table(table(lab.force,hh.size),2)
+# 
+# prop.table(table(lab.force,house.ten),1)
+# prop.table(table(lab.force,house.ten),2)
+```
+
+```{r}
+lm.int.sex.size <- lm(log(expenditure)~income+I(income^2)+hh.size*sex.hh+lab.force+house.ten,data=expend.df)
+summary(lm.int.sex.size)
+#significant
+
+lm.int.sex.labour <- lm(log(expenditure)~income+I(income^2)+hh.size+sex.hh*lab.force+house.ten,data=expend.df)
+summary(lm.int.sex.labour)
+#not significant
+
+lm.int.sex.tenure <- lm(log(expenditure)~income+I(income^2)+hh.size+sex.hh*house.ten+lab.force,data=expend.df)
+summary(lm.int.sex.tenure)
+#not significant
+
+lm.int.size.labour <- lm(log(expenditure)~income+I(income^2)+hh.size*lab.force+sex.hh+house.ten,data=expend.df)
+summary(lm.int.size.labour)
+#not significant
+
+lm.int.size.tenure <- lm(log(expenditure)~income+I(income^2)+hh.size*house.ten+sex.hh+lab.force,data=expend.df)
+summary(lm.int.size.tenure)
+#not significant
+
+lm.int.labour.tenure <- lm(log(expenditure)~income+I(income^2)+hh.size+sex.hh+lab.force*house.ten,data=expend.df)
+summary(lm.int.labour.tenure)
+#significant
+```
+Adding each to the basic model separately only sex:household size and labour:tenure were significant therefore these are the only 2 I will consider for the final model.
+
+#### From full model
+To choose my final model I will start with a full model that includes all possible variables that I want to consider.
+Then using AIC and adjusted R^2 I will remove the coefficients.
+(lower AIC or higher adjusted R^2).
+Once I am happy with the model I end up with I will check the diagnostic plots to make sure the model fit is good/ assumptions met.
+
+```{r}
+lm.log.full<-lm(log(expenditure)~income*lab.force+I(income^2)+income*house.ten+income*hh.size+sex.hh:hh.size+lab.force:house.ten,data=expend.df)
 step.method<- step(lm.log.full,trace=F)
 tidy(step.method)
-AIC(step.method)
-glance(step.method)$adj.r.squared
+#labour force and household tenure interaction removed
+summary(step.method)
 
-#got rid of 2 interactions: income:sex and income:hhsize
+#comparing with original model
 AIC(lm.log.income.sqd)
 AIC(step.method)
+
 glance(lm.log.income.sqd)$adj.r.squared
@@
 
-summary(step.method)
+anova(step.method)
+#All variables left statistically significant
+```
+Using AIC as the model selection criteria I removed 1 interactions labour force:house tenure type.
+Comparing this new model to the original the AIC is significantly lower and the adjusted R squared is significantly higher.
+All variables left in the model are also statistically significant.
 
-anova(step.method)
+Creating diagnostic plots
+```{r}
+par(mfrow=c(2,2))
+plot(step.method)
+```
+All look good... EXPLAIN
+
+```{r}
+# #Comparing to no interaction model
+# lm.log.full.no.int<-lm(log(expenditure)~income+I(income^2)+sex.hh+house.ten+lab.force+hh.size,data=expend.df)
+# summary(lm.log.full.no.int)
+# step.method<- step(lm.log.full.no.int,trace=F)
+# 
+# plot(lm.log.full.no.int)
+# AIC(lm.log.full.no.int)
+# #higher AIC
+
+# glance(lm.log.full.no.int)$adj.r.squared
+# #lower adjusted R^2
 ```
 
-Using AIC as the model selection criteria I removed 2 interactions income:sex and income:household size.
-Comparing this new model to the original the AIC is significantly lower and the adjusted R squarded is significantly higher.
-Also note that each variable is significant except for
-
-```{r}
-plot(step.method)
-```
-
-```{r}
-lm.log.full.no.int<-lm(log(expenditure)~income+I(income^2)+sex.hh+hh.adults+house.ten+lab.force+hh.size,data=expend.df)
-summary(lm.log.full.no.int)
-step.method<- step(lm.log.full.no.int,trace=F)
-
-plot(lm.log.full.no.int)
-```
-
-Using AIC as the selection critera the best model without interactions is the full model.
-
-Now I try adding interactions that couold make logical sense.
-
-As a sex of the household head could impact the effect of income on expenditure (perhaps males are more likely to spend then save then females as they get more money or the reverse)
-
-The number of adults my affect the affect of income as more adults from a higher income household may be more likely to spend more
-
-The type of household tenure may affect the
-
-```{r}
-#adding these into the model 1 by 1
-new.lm.add.sex<-update(lm.log.full.no.int, .~.+income*sex.hh)
-anova(lm.log.full.no.int,new.lm.add.sex)
-summary(new.lm.add.sex)
-
-AIC(lm.log.full.no.int)
-AIC(new.lm.add.sex)
-
-glance(lm.log.full.no.int)$adj.r.squared
-glance(new.lm.add.sex)$adj.r.squared
-#Not really much difference and note that interaction is not significant
-
-new.lm.add.adults<-update(lm.log.full.no.int, .~.+income*hh.adults)
-anova(lm.log.full.no.int,new.lm.add.adults)
-summary(new.lm.add.adults)
-
-AIC(lm.log.full.no.int)
-AIC(new.lm.add.adults)
-
-glance(lm.log.full.no.int)$adj.r.squared
-glance(new.lm.add.adults)$adj.r.squared
-
-new.lm.add.ten<-update(lm.log.full.no.int, .~.+income*house.ten)
-anova(lm.log.full.no.int,new.lm.add.ten)
-summary(new.lm.add.ten)
-
-AIC(lm.log.full.no.int)
-AIC(new.lm.add.ten)
-
-glance(lm.log.full.no.int)$adj.r.squared
-glance(new.lm.add.ten)$adj.r.squared
-
-AIC(lm.log.full)
-AIC(lm.log.full.no.int)
-glance(lm.log.full.no.int)$adj.r.squared
-glance(lm.log.full)$adj.r.squared
-```
-
-```{r}
-# Attempting Box-Cox transformation to check whether distribution is better
-estim_BoxCox <- boxcox(expenditure ~ income+sex.hh+hh.adults+house.ten+lab.force+hh.size+income:hh.adults+income:house.ten+income:lab.force,data =expend.df)
-# Find optimal lambda
-lambda <- estim_BoxCox$x[ which.max( estim_BoxCox$y ) ]
-```
+## Task 8

### 2022-11-06 Draft_1.Rmd
Staring part 8 of inference of final model and slightly rewording earlier inference parts

@@
 Fitting the same model as in task 2 but with the response of expenditure logged $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\epsilon$ gives the above coefficients.
-The data appears to be heteroscedastic as the fitted vs residual plot points spread decreases as fitted values increases, furthermore the linear assumption also appears violated as the residuals appear to exhibit a slight negative quadratic shape.
+The data appears to be heteroscedastic as in the fitted vs residual plot the variance in the residuals decreases as the fitted values increases, furthermore the linear assumption also appears violated as the residuals appear to exhibit a slight negative quadratic shape.
 The QQ plot shows the residuals stay close to the straight line thus normality assumption is valid.
@@
   kable_styling(font_size = 15)
+
 par(mfrow=c(2,2))
@@
 
-# #Comparing R squared
-# glance(lm.income)$r.squared
-# glance(lm.income.sqd)$r.squared
-# glance(lm.log.income)$r.squared
-# glance(lm.log.income.sqd)$r.squared
-
 #Comparing adjusted R squared
@@
 The 4th model $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\beta_{2}x^2_{i}+\epsilon$ has the lowest AIC value of *1528.062* and the highest adjusted R squared value of *0.4635055*.
-Furthermore, looking at the diagnostic plots the normality assumption is valid and the fitted vs residual plot although not 'perfect' displays a more even scattering of points than that of the same plot for the other models.
-Therefore for all these reasons the 4th model is my preferred model to describe the relationship between expenditure and income.
+Furthermore, looking at the diagnostic plots the normality assumption is valid and the fitted vs residual plot displays even scattering of points around 0.
+Therefore the 4th model is my preferred model to describe the relationship between expenditure and income.
 
@@
 Using a more likely realistic increase in weekly income of £20 for a household earning £512 gives an increase of $exp(0.016-0.000015*20-0.000036)-1$x100% = 1.6%.
+
 #### Task 7
@@
 ```{r}
-lm.log.full<-lm(log(expenditure)~income*lab.force+I(income^2)+income*house.ten+income*hh.size+sex.hh:hh.size+lab.force:house.ten,data=expend.df)
+lm.log.full<-lm(log(expenditure)~income*lab.force+I(income^2)+income*house.ten+income*hh.size+sex.hh*hh.size+lab.force:house.ten,data=expend.df)
 step.method<- step(lm.log.full,trace=F)
@@
 plot(step.method)
+
+test.final <- update(step.method,.~.-income:lab.force-income:hh.size-income:lab.force)
+
+AIC(step.method)
+AIC(test.final)
+
+summary(test.final)
+par(mfrow=c(2,2))
+plot(test.final)
+
+par(mfrow=c(2,2))
+plot(step.method)
 ```
-All look good... EXPLAIN
+Diagnostic plots show no issues with the model fit.
+The residuals display a random scattering of points indicating the linearity assumption has been met and the residuals variance does not appear to increase/decrease expect for a few values but I don't think this is enough evidence to say that the homoscedasticity assumption is invalid.
+The points follow a straight line in the QQ plot with the tails deviating slightly but not extreme enough to indicate the normality assumption is invalid.
+The cooks distance plot shows no influential points are present in the data.
+(SHOULD I SAY SOMTHING ABOUT THE SCALE-LOCATION PLOT??)
 
@@
 ## Task 8
+
+The final model is $log(y_{ijkwz})=\beta_{0}+\beta_{1}x_{ijkwz}+\beta_{2}x_{ijkwz}^{2}+\alpha_{j}^{G}+\alpha_k^{S}+\alpha_{w}^{T}+\alpha_{z}^{L}+\beta^{L}x_{ijkwz}+\beta^{S}x_{ijkwz}+\beta^{T}x_{ijkwz}+\alpha^{SG}_{jk}+\epsilon_{ijkwz}$.
+
+```{r}
+data.frame("coefficent"=tidy(step.method)[,1],exp(tidy(step.method)[,2]))
+
+summary(step.method)
+```
+Due to the model being a log-linear model the explanatory variables have a multiplicative effect on the response.
+Exponentiating the estimated coefficients to find their effect on expenditure, first I note that the income interactions although statistically significant are not meaningful in reality as there effect on expenditure is negligible.
+All their coefficients have a value of approximately 1 thus have no effect on expenditure.
+Therefore, I ignore these interaction terms when making inference of the model as they won't change the results by any meaningful amount and it allows the effects of the exploratory variables to be generalised instead of having to consider its effect with each possible value of the variable sit has interactions with.
+
+I consider the explantory variables in approximate importance (affects expenditure the most).
+
+income (with everything else constant) - interactions with  size, labour and tenure (all are insignificant) and squared term
+gender - interaction with size
+tenure - interaction with income 
+size - interaction with gender and income
+labour - interaction with income
+
+Most important variables (can ignore the interactions with income as approximately 1).
+labour force (big decrease one nearly halves),size (large increase up to 80%), tenure (large increase one is 50%), sex (increase 18%- but this is then taken out by interaction with larger households)
+
+Mention confounders, thus association not effect? As can't prove causailty

### 2022-11-07 Draft_1.Rmd
Finishing off model selection process and rewoding interpretation of chosen basic model, Next need to work on inference of final model

@@
 ```
-Income has a positive association with expenditure with a 1 pound increase in household income increasing household expenditure by $exp(0.0032-0.0000029x-0.0000015)-1$x100%.
-Therefore for a household earning the mean income of £512 a £1 increase in income increases expenditure by approximately 0.17%.
-Using a more likely realistic increase in weekly income of £20 for a household earning £512 gives an increase of $exp(0.016-0.000015*20-0.000036)-1$x100% = 1.6%.
+The log-linear model means that the covarites (income) have a multiplicative effect on the response (expenditure).
+The squared term of income included in the model means the change in the mean of expenditure now depends on the value of income.
+Income has a positive association with expenditure with an increase in income causing a percent increase in expenditure.
+With the initial value of income affecting the magnitude of the percent increase on expenditure.
+Smaller initial values of income cause greater percent increase on expenditure (due to the squared income term being negative).
+
+SAYS SUMMARISE SO MAY NOT INCLUDE THE INFOMATION BELOW
+
+In particular, a 1 pound increase in household income increases household expenditure by $exp(0.0032-0.0000029x-0.0000015)-1$x100%.
+Therefore for a household earning the mean income of £512, a £1 increase in income increases expenditure by approximately **0.17%**.
+Using a more realistic increase in weekly income of £20 for a household earning the mean income of £512 gives an increase of $exp(0.016-0.000015*20-0.000036)-1$x100% = **1.6%**.
 
 #### Task 7
-To start I investigated the multicolinearity of the variables to make sure that certain variables were not linearly dependent as this would cause issues with the model if linearly dependent variables were included.
+To start I investigated the multicolinearity.
 
@@
 None of the variables had a GVIF greater than 5, thus multicolinearity is not an issue with the explanatory variables.
-Thus multicolinearity will not be an issue.
 However, considering the variables it seems likely that household size is related to number of household adults.
@@
 ```
-It is clear there is a relationship between the two. I now create linear models includiong each seperately and both together.
+It is clear there is a relationship between the two from the above proportional table. I then created linear models including each separately and both together.
 
@@
 ```
-Looking at the summary outputs of all 3 variables the number of adults in the household is a significant predictor when on its own for predicting expenditure however when we add size of household it is no longer significant.
-Therefore the effect of number of adults can be explained by the size fo the household thus there is no point including it in the model.
-Thus, I will not consider the number of adults in the household as an explanatory variable for the model.
+Looking at the summary outputs the number of adults in a household is a significant predictor when on it is the only variable added to chosen model.
+However, when we add the size of a household variable aswell it is no longer significant.
+Likely the household size was is confounding the effect of number of adults.
+Thus, I will include the size of a household in for the final model but not the number of adults.
 
@@
 First I investigated interactions between income and the factor variables.
-To do this I created scatter plots of income against expenditure with income split into each factor of the chosen explanatory factor variable with a plotted linear model for each of the level of the factor to see if the effect of income varied with each level of the factor variable on expenditure.
-This would be seen by a different gradient of the linear model.
+To do this I created scatter plots of income against expenditure with income split into each factor of the chosen explanatory factor variable with a fitted linear model line plotted for each of the levels of the factor.
+Thus it will be possible to see if the effect of income varies within each level of the factor variable on expenditure.
+This would be seen by a different gradient of the fitted linear model lines.
 
@@
 ```
-Looking at the plots all show that factor variables have an affect on the effect of income on expenditure with different gradients between the levels, however the magnitude of the difference in these gradients varies.
-With sex the difference in gradient is very small indicating that sex has little affect on the effect of income on expenditure.
-Labour force on the hand has a large difference in gradients between its levels giving evidence of an interaction between labour force and income.
+Looking at the plots all show that each of the factor variables have an affect on the effect of income on expenditure with different gradients between the levels, however the magnitude of the difference in these gradients varies.
+For the variable of sex of the household head the difference in gradient is very small indicating that sex has little affect on the effect of income on expenditure.
+Labour force however has a large difference in gradients between its levels giving evidence of an interaction between labour force and income.
 I decided to not include a interaction between sex and income as there is little evidence to support this (and makes logical sense).
+
+To check if the remaining 3 interactions with income are significant I added each separately to the chosen model (aswell as the factor variable which is interacting with income).
 
@@
 ```{r}
-#Checking each of the interactions to see if significant with all variables included
-# lm.int.income.sex <- lm(log(expenditure)~income+I(income^2)+hh.size+sex.hh*income+lab.force+house.ten,data=expend.df)
-# summary(lm.int.income.sex)
-# #not significant as expected
-
 lm.int.income.size <- lm(log(expenditure)~income+I(income^2)+hh.size*income+sex.hh+lab.force+house.ten,data=expend.df)
@@
 
-#significant
 lm.int.income.labour <- lm(log(expenditure)~income+I(income^2)+hh.size+sex.hh+lab.force*income+house.ten,data=expend.df)
@@
 ```
+All 3 interactions were significant so I considered all 3 for my final model.
+
 Considering factor variable interactions. There are 6 possible factor interactions.
-
-```{r}
-# attach(expend.df)
-# prop.table(table(sex.hh,hh.size),1)
-# prop.table(table(sex.hh,hh.size),2)
-# #differences
-# 
-# prop.table(table(sex.hh,house.ten),1)
-# prop.table(table(sex.hh,house.ten),2)
-# 
-# prop.table(table(sex.hh,lab.force),1)
-# prop.table(table(sex.hh,lab.force),2)
-# #differences
-# 
-# prop.table(table(house.ten,hh.size),1)
-# prop.table(table(house.ten,hh.size),2)
-# 
-# prop.table(table(lab.force,hh.size),1)
-# prop.table(table(lab.force,hh.size),2)
-# 
-# prop.table(table(lab.force,house.ten),1)
-# prop.table(table(lab.force,house.ten),2)
-```
 
@@
 ```
-Adding each to the basic model separately only sex:household size and labour:tenure were significant therefore these are the only 2 I will consider for the final model.
+Adding each to the chosen model separately (plus the variables which were interacting) only sex:household size and labour:tenure were significant, therefore these are the only 2 I will consider for the final model.
 
 #### From full model
-To choose my final model I will start with a full model that includes all possible variables that I want to consider.
-Then using AIC and adjusted R^2 I will remove the coefficients.
-(lower AIC or higher adjusted R^2).
-Once I am happy with the model I end up with I will check the diagnostic plots to make sure the model fit is good/ assumptions met.
+To choose my final model I will start with a 'full' model that includes all variables that I want to consider.
+Then using AIC and adjusted R^2 as my model selection criteria I will remove explanatory variables (those that cause the AIC to become lower or increase the adjusted R^2 when removed).
+After this as the model is for exploratory analysis and not prediction a simpler model that is easier to interpret/ understand is preferable (Occam's razor), therefore if the model can simplified to only include the important variables and this does not significantly affect the AIC and R^2 this further reduced model will be my final model.
 
 ```{r}
+#Full model of all variables I want to consider
 lm.log.full<-lm(log(expenditure)~income*lab.force+I(income^2)+income*house.ten+income*hh.size+sex.hh*hh.size+lab.force:house.ten,data=expend.df)
-step.method<- step(lm.log.full,trace=F)
-tidy(step.method)
-#labour force and household tenure interaction removed
-summary(step.method)
+
+AIC(lm.log.full)
+
+#Iteratively performing model selection process
+drop1(lm.log.full)
+#removing lab.force:house.ten will decrease AIC
+lm.full.1.deletion <- update(lm.log.full,.~.-lab.force:house.ten)
+AIC(lm.full.1.deletion)
+
+drop1(lm.full.1.deletion)
+#removing any variables increases AIC thus stop iterative process
 
 #comparing with original model
-AIC(lm.log.income.sqd)
-AIC(step.method)
+glance(lm.log.income.sqd)$adj.r.squared
+glance(lm.full.1.deletion)$adj.r.squared
 
-glance(lm.log.income.sqd)$adj.r.squared
-glance(step.method)$adj.r.squared
+#AIC has been reduced and adjusted R^2 increased
 
-anova(step.method)
-#All variables left statistically significant
+anova(lm.full.1.deletion)
+#All variables left statistically significant except income:house.ten
 ```
+
 Using AIC as the model selection criteria I removed 1 interactions labour force:house tenure type.
 Comparing this new model to the original the AIC is significantly lower and the adjusted R squared is significantly higher.
-All variables left in the model are also statistically significant.
+All variables left in the model are also statistically significant except for income:house tenure.
 
-Creating diagnostic plots
+I then checked how important (magnitude of impact on the response) the variables left in the model are.
+To find their impact on expenditure I exponentiated the estimated coefficients and note that due to log-linear model the effect is multiplicative.
+```{r}
+data.frame("coefficent"=tidy(step.method)[,1],exp(tidy(step.method)[,2]))
+```
+I note that all income interactions are not meaningful/important as their exponentiated value is approximately 1.
+Thus, their effect on expenditure is negligible.
+Thus I remove them from the model as a simpler model is preferred.
+
+```{r}
+#updating to simplier model
+lm.simple.model.1 <- update(lm.full.1.deletion, .~.-income:lab.force-income:hh.size-income:house.ten)
+
+summary(lm.simple.model.1)
+
+drop1(lm.simple.model.1)
+#lowest AIC when don't remove any variables
+
+#Therefore this is my final model
+lm.final.model <- lm.simple.model.1
+```
+
 ```{r}
 par(mfrow=c(2,2))
-plot(step.method)
-
-test.final <- update(step.method,.~.-income:lab.force-income:hh.size-income:lab.force)
-
-AIC(step.method)
-AIC(test.final)
-
-summary(test.final)
-par(mfrow=c(2,2))
-plot(test.final)
-
-par(mfrow=c(2,2))
-plot(step.method)
+plot(lm.final.model)
 ```
 Diagnostic plots show no issues with the model fit.
-The residuals display a random scattering of points indicating the linearity assumption has been met and the residuals variance does not appear to increase/decrease expect for a few values but I don't think this is enough evidence to say that the homoscedasticity assumption is invalid.
+The residuals display a random scattering of points indicating the linearity assumption has been met and the residuals variance does not appear to increase/decrease except for a few values but I don't think this is enough evidence to say that the homoscedasticity assumption is invalid.
 The points follow a straight line in the QQ plot with the tails deviating slightly but not extreme enough to indicate the normality assumption is invalid.
 The cooks distance plot shows no influential points are present in the data.
-(SHOULD I SAY SOMTHING ABOUT THE SCALE-LOCATION PLOT??)
-
-```{r}
-# #Comparing to no interaction model
-# lm.log.full.no.int<-lm(log(expenditure)~income+I(income^2)+sex.hh+house.ten+lab.force+hh.size,data=expend.df)
-# summary(lm.log.full.no.int)
-# step.method<- step(lm.log.full.no.int,trace=F)
-# 
-# plot(lm.log.full.no.int)
-# AIC(lm.log.full.no.int)
-# #higher AIC
-
-# glance(lm.log.full.no.int)$adj.r.squared
-# #lower adjusted R^2
-```
+Therefore
 
@@
 
-The final model is $log(y_{ijkwz})=\beta_{0}+\beta_{1}x_{ijkwz}+\beta_{2}x_{ijkwz}^{2}+\alpha_{j}^{G}+\alpha_k^{S}+\alpha_{w}^{T}+\alpha_{z}^{L}+\beta^{L}x_{ijkwz}+\beta^{S}x_{ijkwz}+\beta^{T}x_{ijkwz}+\alpha^{SG}_{jk}+\epsilon_{ijkwz}$.
+The final model is $log(y_{ijkwz})=\beta_{0}+\beta_{1}x_{ijkwz}+\beta_{2}x_{ijkwz}^{2}+\alpha_{j}^{G}+\alpha_k^{S}+\alpha_{w}^{T}+\alpha_{z}^{L}+\alpha^{SG}_{jk}+\epsilon_{ijkwz}$.
 
 ```{r}
-data.frame("coefficent"=tidy(step.method)[,1],exp(tidy(step.method)[,2]))
+data.frame("coefficent"=tidy(lm.final.model)[,1],exp(tidy(lm.final.model)[,2]),tidy(lm.final.model)[,5])
 
-summary(step.method)
+kable(data.frame("coefficent"=tidy(lm.final.model)[,1],exp(tidy(lm.final.model)[,2]),tidy(lm.final.model)[,5]))%>%
+  kable_classic_2()
+
+summary(lm.final.model)
+
+#Calculating confidence intervals
+
+test <- tidy(lm.final.model)
+
+lwr <- tidy(lm.final.model)$estimate+c(-1)*qnorm(0.975)*tidy(lm.final.model)$std.error
+upr <- tidy(lm.final.model)$estimate+c(1)*qnorm(0.975)*tidy(lm.final.model)$std.error
 ```
 Due to the model being a log-linear model the explanatory variables have a multiplicative effect on the response.
-Exponentiating the estimated coefficients to find their effect on expenditure, first I note that the income interactions although statistically significant are not meaningful in reality as there effect on expenditure is negligible.
-All their coefficients have a value of approximately 1 thus have no effect on expenditure.
-Therefore, I ignore these interaction terms when making inference of the model as they won't change the results by any meaningful amount and it allows the effects of the exploratory variables to be generalised instead of having to consider its effect with each possible value of the variable sit has interactions with.
+Exponentiating the estimated coefficients to find their effect on expenditure gives the above table.
 

### 2022-11-08 Draft_1.Rmd
Updating inference part (part 8)

@@
 
-kable(data.frame("coefficent"=tidy(lm.final.model)[,1],exp(tidy(lm.final.model)[,2]),tidy(lm.final.model)[,5]))%>%
-  kable_classic_2()
-
 summary(lm.final.model)
@@
 
-test <- tidy(lm.final.model)
+lwr <-exp(tidy(lm.final.model)$estimate+c(-1)*qnorm(0.975)*tidy(lm.final.model)$std.error)
+upr <- exp(tidy(lm.final.model)$estimate+c(1)*qnorm(0.975)*tidy(lm.final.model)$std.error)
 
-lwr <- tidy(lm.final.model)$estimate+c(-1)*qnorm(0.975)*tidy(lm.final.model)$std.error
-upr <- tidy(lm.final.model)$estimate+c(1)*qnorm(0.975)*tidy(lm.final.model)$std.error
+Point.estimates <- round(exp(tidy(lm.final.model)[,2]),3)
+p.values <- round(tidy(lm.final.model)[,5],3)
+
+data.frame("coefficent"=tidy(lm.final.model)[,1],
+                 Point.estimates,
+                 "Confidence Interval 95%"=c(paste("(",round(lwr,3),", ",round(upr,3),")")),
+                 p.values)%>%
+  kable(digit=3,caption = "Summary of variables affecting expeniture")%>%
+  kable_classic(full_width=F,latex_options = c("hold_position"))
 ```
@@
 Exponentiating the estimated coefficients to find their effect on expenditure gives the above table.
+I note that all values have been rounded to 3 decimal places thus the zero p-values are due to p-values being too small be displayed and similarly the effect of income squared is too small to be shown.
+
+Interpreting the above table the intercept value is for a male headed household of 1 person, publicly rented in full time employment who has an income of £0. Clearly this does not make sense
+
+I start with the terms that have no interaction.
+For labour force there is a statistically significant difference between those in full time job and those unemployed with those unemployed having an associated decrease in expenditure of (8.5%,34.1%) compared to those in full time work with all other variables held constant.
+The other factor levels of labour force (Part time and inactive) are not significantly different when compared to full time workersa
+
+My inference is drawn from the final model summary [**table x**] above.
+The difference between full time work and being unemployed was highly statistically significant.
+With those unemployed having an associated decrease in expenditure of (8.5%,34.1%) 95% CI compared to those in full time work with all other variables held constant.
+However the difference between full time work and being inactive or in part time work was not significant (CI included 1) therefore with all other variables constant there is no evidence that being in one of these 3 factors compared to another influences the expenditure of the household.
+
+The difference between publicly rented properties expenditure and either privately rented or owned properties expenditure were statistically significant.
+With the associated increase in expenditure of (22%,44.1%) and (11.5%,28.1%) 95% CI's respectively with all other variables held constant.
+
+Due to income squared term the associated effect of an increase in income on expenditure depends on the initial value of income
 

### 2022-11-08 Draft_1.Rmd
Inference update

@@
 
-I start with the terms that have no interaction.
-For labour force there is a statistically significant difference between those in full time job and those unemployed with those unemployed having an associated decrease in expenditure of (8.5%,34.1%) compared to those in full time work with all other variables held constant.
-The other factor levels of labour force (Part time and inactive) are not significantly different when compared to full time workersa
-
 My inference is drawn from the final model summary [**table x**] above.
 The difference between full time work and being unemployed was highly statistically significant.
-With those unemployed having an associated decrease in expenditure of (8.5%,34.1%) 95% CI compared to those in full time work with all other variables held constant.
+With those unemployed having an associated decrease in expenditure of (*8.5%*,*34.1%*) 95% CI compared to those in full time work with all other variables held constant.
 However the difference between full time work and being inactive or in part time work was not significant (CI included 1) therefore with all other variables constant there is no evidence that being in one of these 3 factors compared to another influences the expenditure of the household.
@@
 The difference between publicly rented properties expenditure and either privately rented or owned properties expenditure were statistically significant.
-With the associated increase in expenditure of (22%,44.1%) and (11.5%,28.1%) 95% CI's respectively with all other variables held constant.
+With the associated increase in expenditure of (*22%*,*44.1%*) and (*11.5%*,*28.1%*) 95% CI's respectively with all other variables held constant.
 
-Due to income squared term the associated effect of an increase in income on expenditure depends on the initial value of income
+Due to squared income squared the effect of an increase in income on expenditure depends on the initial value of income.
+Assuming all other variables are held constant the percentage increase o expenditure is $(\frac{exp(\beta_1(x+y)+\beta_2(x+y)^2)}{exp(\beta_1x+\beta_2x^2)}-1)*100$ where x is the initial value of income and y is the increase in income, $\beta_1=0.002285$ and $\beta_2=-0.000001036$.
+Thus, for a household earning the mean income of £512, a £20 increase in income increases expenditure by approximately *2.4%* whereas for a household earning only £300 the same £20 increase, increases expenditure by approximately *3.3%*.
+Due to the negative quadratic term the higher the initial income the less of a positive multiplicative effect an increase in expenditure has on expenditure.
 
-I consider the explantory variables in approximate importance (affects expenditure the most).
+```{r,echo=F}
+#calculating percent increase
+beta_1 <- 0.002285
+beta_2 <- -0.000001036
+x <- 300
+inc <- 20
 
-income (with everything else constant) - interactions with  size, labour and tenure (all are insignificant) and squared term
-gender - interaction with size
-tenure - interaction with income 
-size - interaction with gender and income
-labour - interaction with income
+(exp(beta_1*(x+inc)+beta_2*(x+inc)^2)/exp(beta_1*x+beta_2*x^2)-1)*100
+```
 
-Most important variables (can ignore the interactions with income as approximately 1).
-labour force (big decrease one nearly halves),size (large increase up to 80%), tenure (large increase one is 50%), sex (increase 18%- but this is then taken out by interaction with larger households)
+The final 2 variables, sex of household head and household size have an interaction between them.
+However, only the household size of 3 level has a statistically significant interaction with sex of household head.
+Therefore as the other levels
+
+The sex of the household head is highly statistically significant with being female head increasing the expenditure by (*5.6%*,*25.9%*) 95% CI compared to a male head with all other variables held constant.
+Household size is extremely statistically significant with an increase in expenditure of (*22.4%*,*45.5*),  (*31.3%*,*66.8%*), (*34.9%*,*81.6%*) 95% CI's for households of size 2,4,5+ when compared to a household of size 1 with all other variables held constant.
+For the effect on expenditure between a household of size 1 and size 3 because of the significant interaction with sex if the household head is male the CI is (*36.2%*,*71.9%*) and (*-15.7%*,*+52.3%*) if female is it is with all other variables held constant.
+NOte that the for the fmale head CI 0% is included therefore the being a
+
+FINISH THIS BIT OFF!
 

### 2022-11-11 Draft_1.Rmd
New version of task 2, using code from the tutorial solutions

@@
 For the effect on expenditure between a household of size 1 and size 3 because of the significant interaction with sex if the household head is male the CI is (*36.2%*,*71.9%*) and (*-15.7%*,*+52.3%*) if female is it is with all other variables held constant.
-NOte that the for the fmale head CI 0% is included therefore the being a
+Note that the for the female head CI 0% is included therefore there is no statistically significant difference in expenditure between a household of size 1 with male head and a household of size 3 with a female head with all other variables held constant.
 
-FINISH THIS BIT OFF!
+Overall an increase in income, higher household size, being a female head and living in a owned or privately rented increase the % expenditure when compared to a single male headed household living in a publicly rented house (unless a female head with size 3 household).
+Whereas being unemployed compared to in full time work decreases the % of expenditure.
 
-Mention confounders, thus association not effect? As can't prove causailty
+All the variables mentioned above were statistically significant in their association with expenditure.
+However, just because they are significant, it does not mean they influence expenditure.
+There could be confounding or simple chance causing these results.
+However, some areas of confidence are:
+
+* income, logically if you make more money you have more available to spend 
+* household size, larger households are more expensive thus require more spending
+
+sex, tenure, labour force

### 2022-11-11 Task_2_updated.Rmd
New version of task 2, using code from the tutorial solutions

---
title: "Untitled"
output: html_document
date: "2022-11-11"
---
```{r}
#Setting up enviroment
setwd("~/Desktop/GLM/Coursework/GLM-CW")
airline.df <- read.table("airline.txt",header=T)

#Setting up inputs of mle function
n <- dim(airline.df)[[1]]
con <- rep(1,n)
X=matrix(c(con,airline.df$x),nrow=n,ncol=2)
y <- airline.df$y
```

```{r}
#Setting conditions of when to end fisher scoring algorithm

#max number of iterations of algorithm allowed (set high incase initial guess far off)
max.iter <- 1000
#Value to define convergence
eps<-0.001;

#setting up matrix to store values of beta at each iteration
beta.ini <- matrix(NA,max.iter,dim(X)[2])

#initial guess for beta
beta.ini[1,] <- c(0,1)

flag <- 0
c<-1

#Loop to apply fisher scoring algorithm
while (flag==0 & c < max.iter){

#Finding mu for cth iteration  
mu <- exp(beta.ini[c,]%*%t(X))
#finding r where X^t%*%r is the score function
r <- c((y/mu)-1)
#applying fisher scoring algorithm to find c+1 iteration
beta <- beta.ini[c,]+(solve(t(X)%*%X))%*%t(X)%*%r
#storing value of new beta
beta.ini[(c+1),] <- beta

#checking for convergence
if (abs(beta.ini[c,1]-beta.ini[c+1,1]+beta.ini[c,2]-beta.ini[c+1,2])< eps){
flag <- 1
}

c <- c+1
}
#storing answers
beta.iteration <- beta.ini[complete.cases(beta.ini),]
beta.fin=beta.ini[c,]

beta.fin
```

```{r}
#Finding the standard errors of beta
std.error=sqrt(diag(solve(t(X)%*%X)))
std.error

#T-statistic
beta.fin[2]/std.error[2]

qt(0.975,1)
```

### 2022-11-17 Draft_final.Rmd
Adding final draft

---
title: "General linear Models Coursework"
author: "Student ID: 34273638"
geometry: margin = 1.5cm
output:
  pdf_document: default
header-includes:
  - \usepackage{wrapfig}
  - \usepackage{lipsum}
  - \usepackage{setspace}
  - \usepackage{titlesec}
  - \titlespacing{\title}{0pt}{\parskip}{-\parskip}
---
# Task 1:
\vspace{-5truemm}
## Part 1: Assessing distribution of expenditure and relationship between it and covariates
\vspace{-5truemm}
```{r,echo=F,message=F}
#Initial environment setup
setwd("~/Desktop/GLM/Coursework/GLM-CW")
expend.df <- read.table("expenditure.txt")
library(tidyverse)
library(kableExtra)
library(gridExtra)
library(broom)
library(MASS)
```

```{r,echo=F}
#Data cleaning

#coding variables as factors
expend.df$house.ten <- factor(expend.df$house.ten,levels=c(1:3),labels=c("Public Rented","Private rented","Owned"))
expend.df$sex.hh <- factor(expend.df$sex.hh,levels=c(1,2),labels=c("Male","Female"))
expend.df$lab.force <- factor(expend.df$lab.force,levels=c(1:4),labels=c("Full time", "Part time","Unemployed","Inactive"))
expend.df$hh.size <- factor(expend.df$hh.size,levels=c(1:5),labels = c("1 person","2 people","3 people","4 people","5+ persons"))
expend.df$hh.adults <- factor(expend.df$hh.adults,levels = c(1:4), labels = c("1 adult","2 adults","3 adults","4+ adults"))

```

```{r,echo=F,fig.align = 'left',fig.width=10,fig.height=4}
#Histogram to examine the expenditure
hist(expend.df$expenditure,breaks = 20,
     xlab = "Total weekly household expenditure (GBP)",
     main = "Histogram of expenditure")
```

```{r,echo=F,fig.align='center'}
#Creating the 
#scatter plot to look at relationship between income and expenditure
p1<-ggplot(data=expend.df,aes(x=income,y=expenditure))+
  geom_point(size=0.2)+
  xlab("Gross weekly average household income (GBP)")+
  ylab("Total expenditure")+
  theme(axis.text=element_text(size=6),
        axis.title=element_text(size=8))

#boxplots
p2<-ggplot(data=expend.df,aes(y=expenditure,x=house.ten,fill=house.ten))+
  geom_boxplot(outlier.size = 0.5,outlier.shape = 4)+
  xlab("Household tenure")+
  ylab("Total expenditure")+
  guides(fill= "none")+
  theme(axis.text.x = element_text(size = 6))

#scale_fill_discrete(name="Household\ntenure")

p3<-ggplot(data=expend.df,aes(y=expenditure,x=sex.hh,fill=sex.hh))+
  geom_boxplot(outlier.size = 0.5,outlier.shape = 4)+
  xlab("Sex of the household head")+
  ylab("Total expenditure")+
  guides(fill= "none")

p4<-ggplot(data=expend.df,aes(y=expenditure,x=lab.force,fill=lab.force))+
  geom_boxplot(outlier.size = 0.5,outlier.shape = 4)+
  xlab("Employment status")+
  ylab("Total expenditure")+
  guides(fill= "none")+
  theme(axis.text.x = element_text(size = 6))

p5<-ggplot(data=expend.df,aes(y=expenditure,x=hh.size,fill=hh.size))+
  geom_boxplot(outlier.size = 0.5,outlier.shape = 4)+
  xlab("Household size")+
  ylab("Total expenditure")+
  guides(fill= "none")+
  theme(axis.text.x = element_text(size = 6))

p6<-ggplot(data=expend.df,aes(y=expenditure,x=hh.adults,fill=hh.adults))+
  geom_boxplot(outlier.size = 0.5,outlier.shape = 4)+
  xlab("Number of adults in the household")+
  ylab("Total expenditure")+
  guides(fill= "none")

grid.arrange(p1,p2,p3,p4,p5,p6,ncol=2,nrow=3)
```

\newpage

To assess the distribution of expenditure I produced the above histogram, it shows that expenditure has a positively skewed distribution (and thus the normality assumption used in linear regression may be invalid). Using a scatter plot to investigate the relationship between household income and expenditure , it shows there is a clear positive association.
Then to investigate the other categorical variables I used boxplots to display differences in the expenditure across the different levels (categories) of each variable. The differences shown in the plots are listed below:

* Different household tenures categories have different weekly expenditure amounts with public rented households having a noticeably lower expenditure compared to the other 2 household levels.
* Households with a female head have a slightly lower weekly expenditure compared to male household heads
* Employment status appears to be associated with expenditure with unemployed households having a dramatically lower expenditure compared to households with full or part time employment status and inactive employment status households having a lower expenditure compared to households with full or part time employment status.
* Household size generally appears to be positively correlated with expenditure, with larger households having a higher weekly expenditure (however 5+ person households break this trend with the expenditure being slightly decreasing, however this could be due to other variables or a small sample size of such households).
* Number of adults in the household also appears positively correlated with expenditure with households containing more adults having a higher expenditure.

I note that these differences and possible associations may not be due to the variable mentioned but due to a third variable (a confounder) that affects both the categorical variable and expenditure (the response).

## Part 2: Regressing expenditure on income
Fitting the model $y_{i}=\beta_{0}+\beta_{1}x_{i}+\epsilon$ where $y$ is the weekly household expenditure and the explanatory variable $x$ is weekly income and $i \in \{1,...,1200\}$ is the ith household in the dataset produces the below table (these are the same for all linear models mentioned in parts 2 to 6).It shows the estimated values of the intercept ($\beta_0$) and the estimated 'effect' of a 1 pound increase of income on expenditure ($\beta_1$) and their associated standard errors.
```{r,echo=F}
#Question 2
#Fitting linear model
lm.income<-lm(expenditure~income,data = expend.df)
#creating table to display results
knitr::kable(as.data.frame(tidy(lm.income)[,1:3]),digits=3)%>%
  kable_styling(position = "center",latex_options = "HOLD_position")
```

```{r,echo=F,fig.height=4.2}
#Creating diagnostic plots
par(mfrow=c(1,2))
plot(lm.income,1)
plot(lm.income,2)
```
\newpage

Viewing the above diagnostic plots for this model. The fitted vs residual plot shows the residuals are not equally spread around 0 with a greater spread of residuals above zero, meaning the assumption of linearity may not hold. Furthermore, the QQ normal plot has a positive U shape with the tails diverging heavily away from the straight line. This implies the normality assumption is invalid.

## Part 3: Regressing expenditure on income and income squared
Fitting a new linear model $y_{i}=\beta_{0}+\beta_{1}x_{i}+\beta_2x^2_{i}+\epsilon$ with an added squared term of income gives the below estimated for the coefficients and their standard errors.

```{r,echo=F}
#Question 3
#Fitting model
lm.income.sqd <- lm(expenditure~income+I(income^2),data = expend.df)

#Producing table of results
knitr::kable(as.data.frame(tidy(lm.income.sqd)[,1:3]),digits=3)%>%
  kable_styling(position = "center",latex_options = "HOLD_position")
```

```{r,echo=F,fig.height=4.2}
#Creating diagnostic plots
par(mfrow=c(1,2))
plot(lm.income.sqd,1)
plot(lm.income.sqd,2)
```

Looking at the produced diagnostic plots above for this model, in the fitted vs residual plot the residual values increase as the fitted values increase thus the homoscedasticity assumption may not be valid. The tails in the normal QQ plot also still deviate heavily away from the straight line meaning the normality assumption is invalid.

\newpage

## Part 4: Regressing the natural logarithm of expenditure on income
Fitting the same model as in task 2 but with the response of expenditure logged (a log-linear model) $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\epsilon$ gives the below estimated coefficients and standard errors.

```{r,echo=F}
#Question 4
#Fitting model
lm.log.income<-lm(log(expenditure)~income,data = expend.df)

#Producing table of results
knitr::kable(as.data.frame(tidy(lm.log.income)[,1:3]),digits=3)%>%
  kable_styling(position = "center",latex_options = "HOLD_position")
```

```{r,echo=F,fig.height=4.2}
#Creating diagnostic plots
par(mfrow=c(1,2))
plot(lm.log.income,1)
plot(lm.log.income,2)
```

Looking at the produced diagnostic plots above for this model, the data appears to be heteroscedastic as in the fitted vs residual plot the variance in the residuals decreases as the fitted values increases. Furthermore, the linear assumption also appears violated as the residuals appear to exhibit a slight negative quadratic shape. The QQ plot shows the residuals stay close to the straight line thus normality assumption is valid.

## Part 5: Regressing the natural logarithm of expenditure on income and income squared
Fitting the model $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\beta_{2}x^2_{i}+\epsilon$ gives the estimated coefficients and standard errors in the table below.

```{r,echo=F}
#Question 4
#Fitting model
lm.log.income.sqd <- lm(log(expenditure)~income+I(income^2),data = expend.df)

#Producing table of results
knitr::kable(as.data.frame(tidy(lm.log.income.sqd)[,1:3]),digits=3)%>%
  kable_styling(position = "center",latex_options = "HOLD_position")
```

```{r,echo=F,fig.height=8}
#Creating diagnostic plots
par(mfrow=c(2,2))
plot(lm.log.income.sqd)
```

Producing the diagnostic plots above, the fitted vs residuals plot shows a fairly even scattering of points around 0 and a linear trend, however the variance of the residuals appears to decrease slightly as the fitted values increase. But, looking at the scale-location plot the residuals appear randomly spread around the red line and the red line through them stays fairly horizontal thus homoskedasticity assumptions appears valid. The QQ plot shows the points stay close to the straight line meaning the normality assumption is valid. The residual vs Leverage plot shows no influential points.

\newpage
## Part 6: Comparing models
Comparing the models from part 2-5 using AIC and adjusted $R^2$:

```{r,echo=F}
#Creating dataframe to store measures to compare models by
Model.comparison.data <- data.frame("Model"=c("Part 2","Part 3","Part 4", "Part 5"),
           "AIC"=c(AIC(lm.income),AIC(lm.income.sqd),AIC(lm.log.income),AIC(lm.log.income.sqd)),
           "Adj R squared"=c(glance(lm.income)$adj.r.squared,
                             glance(lm.income.sqd)$adj.r.squared,
                             glance(lm.log.income)$adj.r.squared,
                             glance(lm.log.income.sqd)$adj.r.squared))

knitr::kable(Model.comparison.data,digits=3)%>%
  kable_styling(position = "center",latex_options = "HOLD_position")
```
The model from part 5, $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\beta_{2}x^2_{i}+\epsilon$ has the lowest AIC value of **1528.062** and the highest adjusted R squared value of **0.4635055**. Furthermore, from the diagnostic plots the normality assumption is valid and the fitted vs residual plot displays an even scattering of points around 0. Therefore the 4th model is my preferred model to describe the relationship between expenditure and income.

```{r,echo=F}
#Producing table of results
knitr::kable(as.data.frame(tidy(lm.log.income.sqd)[,1:3]),digits=3)%>%
  kable_styling(position = "center",latex_options = "HOLD_position")
```

The log-linear model means that the covarites (income and income squared) have a multiplicative effect on the response (expenditure). The squared term of income included in the model, means that the change in the mean of expenditure now depends on the value of income. Income has a positive association with expenditure with an increase in income causing a percent increase in expenditure. With the initial value of income affecting the magnitude of the percent increase on expenditure. Smaller initial values of income cause a greater percent increase on expenditure (due to the squared income term being negative). In particular, using a realistic increase in weekly income of £20 for a household earning the mean income of £512 gives an increase of $(\frac{exp(532\beta_1+532^{2}\beta_{2})}{exp(512\beta_1+512^{2}\beta_{2})}-1)$x100% = **3.25%**.

```{r,include=F}
#calculating percent increase
beta_1 <- 3.117180e-03	
beta_2 <- -1.453272e-06	
x <- 512
inc <- 20
#Formual to calculate oercent increase
(exp(beta_1*(x+inc)+beta_2*(x+inc)^2)/exp(beta_1*x+beta_2*x^2)-1)*100
```

## Part 7: Finding a suitable regression model for expenditure
To start I investigated the multicolinearity of the variables in the dataset.

```{r,echo=F}
#Checking variance inflation factors (generalized as variables include factors)
car::vif(lm(log(expenditure)~income+I(income^2)+hh.size+hh.adults+sex.hh+lab.force+house.ten,data=expend.df))
```
None of the variables had a GVIF greater than 5, thus multicolinearity is not an issue with the explanatory variables. However, considering the variables it seems likely that household size is related to number of household adults.

```{r,echo=F}
#Creating tables for chi-squared test
prop.table(table(expend.df$hh.size,expend.df$hh.adults),2)
```
From the proportional table above there appears to be a relationship between the two variables. To investigate this further I created linear models including each variable separately and both together added to the model from part 5. In the model with only the number of adults variable added using an F-test the variable was significant at $\alpha=0.05$ with a p-value of **2.271e-11**. However, when adding both number of adults and household size only household size was significant with a p-value of **7.953e-15** and number of adults was no longer significant with a p-value of **0.7249** . Likely the household size was is confounding the effect of number of adults. Thus, I will include the size of a household in for the final model but not the number of adults.

```{r,include=F}
#Producing models to check relationship of number of adults and household size

lm.adult <- lm(log(expenditure)~income+I(income^2)+hh.adults,data = expend.df)
anova(lm.adult)
#adult significant

lm.size <- lm(log(expenditure)~income+I(income^2)+hh.size,data = expend.df)
anova(lm.size)
#size significant

lm.adult.size <- lm(log(expenditure)~income+I(income^2)+hh.size+hh.adults,data = expend.df)
anova(lm.adult.size)
#adult no longer significant
```

Considering the possible interactions, I decided to only include 2 way interactions as higher order interactions make interpretation of the model much more complicated and there is no clear 3 way interactions that make sense.

First I investigated interactions between income and the factor variables. I added each separately to the model from part 5 and the variable part of the interaction and used an F-test to see if the interaction was significant at the 5% level. income:labour force, income: household size and income: house tenure type were all significant, however income:sex was not. Therefore I will consider the 3 significant interactions for my final model.

```{r,include=F}
#Testing if interactions are significant
lm.int.income.sex <- lm(log(expenditure)~income*sex.hh+I(income^2),data=expend.df)
anova(lm.int.income.sex)
#Not significant

lm.int.income.lab <- lm(log(expenditure)~income*lab.force+I(income^2),data=expend.df)
anova(lm.int.income.lab)
#Significant

lm.int.income.size <- lm(log(expenditure)~income*hh.size+I(income^2),data=expend.df)
anova(lm.int.income.size)
#Significant

lm.int.income.ten <- lm(log(expenditure)~income*house.ten+I(income^2),data=expend.df)
anova(lm.int.income.ten)
#Significant
```
Next considering factor variable interactions. There are 6 possible factor interactions, of which I added each to the chosen model separately (plus the variables which were interacting) only sex:household size and labour:tenure were significant at the 5% level. Therefore, these are the only 2 I will consider for the final model.

```{r, include=F}
lm.int.sex.size <- lm(log(expenditure)~income+I(income^2)+hh.size*sex.hh+lab.force+house.ten,data=expend.df)
anova(lm.int.sex.size)
#significant

lm.int.sex.labour <- lm(log(expenditure)~income+I(income^2)+hh.size+sex.hh*lab.force+house.ten,data=expend.df)
anova(lm.int.sex.labour)
#not significant

lm.int.sex.tenure <- lm(log(expenditure)~income+I(income^2)+hh.size+sex.hh*house.ten+lab.force,data=expend.df)
anova(lm.int.sex.tenure)
#not significant

lm.int.size.labour <- lm(log(expenditure)~income+I(income^2)+hh.size*lab.force+sex.hh+house.ten,data=expend.df)
anova(lm.int.size.labour)
#not significant

lm.int.size.tenure <- lm(log(expenditure)~income+I(income^2)+hh.size*house.ten+sex.hh+lab.force,data=expend.df)
anova(lm.int.size.tenure)
#not significant

lm.int.labour.tenure <- lm(log(expenditure)~income+I(income^2)+hh.size+sex.hh+lab.force*house.ten,data=expend.df)
anova(lm.int.labour.tenure)
#significant
```

To choose my final model I will start with a 'full' model that includes all variables that I want to consider. Then using AIC and adjusted $R^2$  as my model selection criteria I will remove explanatory variables (those that cause the AIC to become lower or increase the adjusted $R^2$  when removed). After this as the model is for exploratory analysis and not prediction a simpler model that is easier to interpret/ understand is preferable (Occam's razor), therefore if the model can simplified to only include the important variables and this does not significantly affect the AIC and $R^2$ this further reduced model will be my final model.

My 'full' model regresses the natural logarithm of expenditure on income, income squared, household tenure type, household size, sex of household head, labour force and interactions between income with tenure type, labour force and household size and also interactions between sex of household head and household size and labour force and tenure type. This can be fitted in R using the belwo code:

```{r,eval=F}
lm.log.full<-lm(log(expenditure)~income*lab.force+I(income^2)+income*house.ten+
                  income*hh.size+sex.hh*hh.size+lab.force:house.ten,data=expend.df)
```

Then using the iterative method of deleting the variable that will reduce the AIC by the largest amount until no decrease in AIC is available. Implementing this algorithm I removed 1 variable, the interaction between labour force and household tenure type as this reduced the AIC from **1391.914** to **1386.878**. All variables left in the updated model are also statistically significant except for income:house tenure.

```{r, include=F}
#Full model of all variables I want to consider
lm.log.full<-lm(log(expenditure)~income*lab.force+I(income^2)+income*house.ten+
                  income*hh.size+sex.hh*hh.size+lab.force:house.ten,data=expend.df)

AIC(lm.log.full)

#Iteratively performing model selection process
drop1(lm.log.full)
#removing lab.force:house.ten will decrease AIC
lm.full.1.deletion <- update(lm.log.full,.~.-lab.force:house.ten)
AIC(lm.full.1.deletion)

drop1(lm.full.1.deletion)
#removing any variables increases AIC thus stop iterative process

#comparing with original model
glance(lm.log.income.sqd)$adj.r.squared
glance(lm.full.1.deletion)$adj.r.squared

#AIC has been reduced and adjusted R^2 increased

anova(lm.full.1.deletion)
#All variables left statistically significant except income:house.ten
```

I then checked how important (the magnitude of impact on the response) the variables left in the model are. To find their impact on expenditure I exponentiated the estimated coefficients and note that due to log-linear model the effect is multiplicative.

```{r,include=F}
#Looking at which exponentiated coefficients are approx 1
data.frame("coefficent"=tidy(lm.full.1.deletion)[,1],exp(tidy(lm.full.1.deletion)[,2]))
```

```{r,echo=F}
#Producing table of all exponentiated regession coefficients that are approx 1
knitr::kable(data.frame(tidy(lm.full.1.deletion)[14:22,1],exp(tidy(lm.full.1.deletion)[14:22,2])),digits=3)%>%
  kable_styling(position = "center",latex_options = "HOLD_position")
```

Above is a table of the exponentatied estimated coefficients of all the variables which were approximately 1. These are all the possible interaction levels (excluding the dummy level) of the income interactions. Thus all income interactions are not meaningful/important as their exponentiated value is approximately 1. Therefore, their effect on expenditure is negligible so I remove them from the model as a simpler model is preferred. After removing these 3 variables, removing any more does not further decrease the AIC thus this simplified model is my final model.
```{r,include=F}
#updating to simpler model
lm.simple.model.1 <- update(lm.full.1.deletion, .~.-income:lab.force-income:hh.size-income:house.ten)

summary(lm.simple.model.1)

drop1(lm.simple.model.1)
#lowest AIC when don't remove any variables

#Therefore this is my final model
lm.final.model <- lm.simple.model.1
```

```{r,echo=F,fig.height=6}
#Producing diagnostic plots
par(mfrow=c(2,2))
plot(lm.final.model)
```

Producing diagnostic plots for this final model shows no issues with the model fit. The residuals display a random scattering of points indicating the linearity assumption has been met and the residuals variance does not appear to increase/decrease except for a few values but I don't think this is enough evidence to say that the homoscedasticity assumption is invalid. This is further backed up by a straight nearly horizontal red line in the scale-location plot with an even random scattering of residuals around it. The points follow a straight line in the QQ plot with the tails deviating slightly but not extreme enough to indicate the normality assumption is invalid. The cooks distance plot shows no influential points are present in the data.

## Part 8: Inference of final model
The final model is $log(y_{ijkwz})=\beta_{0}+\beta_{1}x_{ijkwz}+\beta_{2}x_{ijkwz}^{2}+\alpha_{j}^{G}+\alpha_k^{S}+\alpha_{w}^{T}+\alpha_{z}^{L}+\alpha^{SG}_{jk}+\epsilon_{ijkwz}$.

### 2022-11-18 Deleted_stuff.Rmd
adding chnages

---
title: "Untitled"
output: html_document
date: "2022-10-29"
---
```{r,echo=F,eval=F}
#extra code that may not use
pairs(expend.df)
xtabs(~expend.df$house.ten+expend.df$sex.hh)

par(mfrow=c(1,2))
hist(expend.df$expenditure,breaks = 20,
     xlab = "Total weekly household expenditure (GBP)")
hist(log(expend.df$expenditure),breaks = 20,
     xlab = " Log of total weekly \n household expenditure (GBP)")
#Applying log transform gives a more normal looking distribution
```

```{r}
# #as expected the normal QQ plot shows 'bad fit'

# #try gamma as good for postively skewed continous data
# gamma.income <-glm(expenditure~income,data=expend.df,family="Gamma")
# summary(gamma.income)
# plot(gamma.income)
# #bad fit (explain)
# 
# #out of all models log linear is the most appropriate
```

```{r}
#Creating model with all possible 2 way interactions (not a good idea but just for interest :P)
lm.log.full<-lm(log(expenditure)~income+I(income^2)+income*sex.hh+income*hh.adults+income*house.ten+income*lab.force+income*hh.size+sex.hh*hh.adults+sex.hh*house.ten+sex.hh*lab.force+sex.hh*hh.size+hh.adults*house.ten+hh.adults*lab.force+hh.adults*hh.size+house.ten*lab.force+house.ten*hh.size+lab.force*hh.size,data=expend.df)
summary(lm.log.full)

step.method<- step(lm.log.full,trace=F)
summary(step.method)

```

### 2022-11-18 Draft_1.tex
adding chnages

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Untitled},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Untitled}
\author{}
\date{\vspace{-2.5em}2022-10-27}

\begin{document}
\maketitle

\begin{verbatim}
## -- Attaching packages --------------------------------------- tidyverse 1.3.2 --
## v ggplot2 3.3.5     v purrr   0.3.4
## v tibble  3.1.6     v dplyr   1.0.8
## v tidyr   1.2.0     v stringr 1.4.0
## v readr   2.1.2     v forcats 0.5.1
## -- Conflicts ------------------------------------------ tidyverse_conflicts() --
## x dplyr::filter() masks stats::filter()
## x dplyr::lag()    masks stats::lag()
## 
## Attaching package: 'kableExtra'
## 
## 
## The following object is masked from 'package:dplyr':
## 
##     group_rows
## 
## 
## 
## Attaching package: 'gridExtra'
## 
## 
## The following object is masked from 'package:dplyr':
## 
##     combine
\end{verbatim}

\hypertarget{task-1}{%
\subsection{Task 1:}\label{task-1}}

\includegraphics{Draft_1_files/figure-latex/unnamed-chunk-3-1.pdf}

\begin{verbatim}
## Warning: `guides(<scale> = FALSE)` is deprecated. Please use `guides(<scale> = "none")` instead.
## `guides(<scale> = FALSE)` is deprecated. Please use `guides(<scale> = "none")` instead.
## `guides(<scale> = FALSE)` is deprecated. Please use `guides(<scale> = "none")` instead.
## `guides(<scale> = FALSE)` is deprecated. Please use `guides(<scale> = "none")` instead.
## `guides(<scale> = FALSE)` is deprecated. Please use `guides(<scale> = "none")` instead.
\end{verbatim}

\includegraphics{Draft_1_files/figure-latex/unnamed-chunk-3-2.pdf}

Looking at the above histogram of expenditure it shows that expenditure
has a positively skewed distribution. Noticing this after applying a log
transform the histogram of log expenditure has a more normal looking
distribution.

Investigating covarits against expenditure we can see that there is a
positive association between household income and expenditure. Looking
at the various factor variables using box plots\ldots{}

\hypertarget{part-2}{%
\paragraph{Part 2:}\label{part-2}}

\begin{table}
\centering\begingroup\fontsize{15}{17}\selectfont

\begin{tabular}{l|r|r}
\hline
term & estimate & std.error\\
\hline
(Intercept) & 147.223766 & 10.3856098\\
\hline
income & 0.486028 & 0.0177632\\
\hline
\end{tabular}
\endgroup{}
\end{table}

\includegraphics{Draft_1_files/figure-latex/unnamed-chunk-6-1.pdf}
Fitting a linear model \(y_{i}=\beta_{0}+\beta_{1}x_{i}+\epsilon\) with
the response \(y\) the expenditure,the explanatory variable \(x\) the
income and \(i \in \{1,...,1200\}\). The above table shows the estimated
values of the intercept (\(\beta_0\)) and the estimated `effect' of a 1
pound increase in income on expenditure \(\beta_1\) and there standard
errors.

Viewing the diagnostic plots for this model. The fitted vs residual plot
shows the residuals are not equally spread around 0 with a greater
spread of residuals above zero, meaning the assumption of linearity may
not hold. Furthermore, the QQ normal plot has a u shape with the tails
diverging heavily away from the straight line. This implies the
normality assumption is invalid.

\hypertarget{part-3}{%
\paragraph{Part 3:}\label{part-3}}

\begin{table}
\centering\begingroup\fontsize{15}{17}\selectfont

\begin{tabular}{l|r|r}
\hline
term & estimate & std.error\\
\hline
(Intercept) & 103.1753319 & 18.0506419\\
\hline
income & 0.6897522 & 0.0706466\\
\hline
I(income^2) & -0.0001761 & 0.0000591\\
\hline
\end{tabular}
\endgroup{}
\end{table}

\includegraphics{Draft_1_files/figure-latex/unnamed-chunk-7-1.pdf}
Fitting a new linear model
\(y_{i}=\beta_{0}+\beta_{1}x_{i}+\beta_2x^2_{i}+\epsilon\) gives the
above estimated coefficients. Looking at the diagnostic plots, in the
fitted vs residual plot the residual values increase as the fitted
values increase thus the homoscedasticity assumption may not be valid.
The tails in the normal QQ plot also still deviate heavily away from the
straight line meaning the normality assumption is invalid.

\hypertarget{part-4}{%
\paragraph{Part 4:}\label{part-4}}

\begin{table}
\centering\begingroup\fontsize{15}{17}\selectfont

\begin{tabular}{l|r|r}
\hline
term & estimate & std.error\\
\hline
(Intercept) & 5.0739718 & 0.0281893\\
\hline
income & 0.0014356 & 0.0000482\\
\hline
\end{tabular}
\endgroup{}
\end{table}

\includegraphics{Draft_1_files/figure-latex/unnamed-chunk-8-1.pdf}
Fitting the model \(log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\epsilon\). The
data appears to be hetroscedastic as the fitted vs residual plot points
spread decreases as fitted values increases/ not. The QQ plot shows the
residuals stay close to the straight line thus normality assumption is
valid.

\hypertarget{part-5}{%
\paragraph{Part 5:}\label{part-5}}

\begin{verbatim}
## 
## Call:
## lm(formula = log(expenditure) ~ income + I(income^2), data = expend.df)
## 
## Residuals:
##      Min       1Q   Median       3Q      Max 
## -1.73608 -0.26761 -0.00464  0.27537  2.10503 
## 
## Coefficients:
##               Estimate Std. Error t value Pr(>|t|)    
## (Intercept)  4.710e+00  4.747e-02  99.223   <2e-16 ***
## income       3.117e-03  1.858e-04  16.777   <2e-16 ***
## I(income^2) -1.453e-06  1.554e-07  -9.349   <2e-16 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Residual standard error: 0.4564 on 1197 degrees of freedom
## Multiple R-squared:  0.4644, Adjusted R-squared:  0.4635 
## F-statistic: 518.9 on 2 and 1197 DF,  p-value: < 2.2e-16
\end{verbatim}

\includegraphics{Draft_1_files/figure-latex/unnamed-chunk-9-1.pdf}

\(log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\beta_{2}x^2_{i}+\epsilon\)

\end{document}

### 2022-11-18 Draft_final.Rmd
Add files via upload

@@
 #creating table to display results
-knitr::kable(as.data.frame(tidy(lm.income)[,1:3]),digits=3)%>%
+knitr::kable(as.data.frame(tidy(lm.income)[,1:3]),digits=3,caption = "Estimated coefficients and SE for part 2 model")%>%
   kable_styling(position = "center",latex_options = "HOLD_position")
@@
 
-```{r,echo=F,fig.height=4.2}
+```{r,echo=F,fig.height=3.8}
 #Creating diagnostic plots
@@
 #Producing table of results
-knitr::kable(as.data.frame(tidy(lm.income.sqd)[,1:3]),digits=3)%>%
+knitr::kable(as.data.frame(tidy(lm.income.sqd)[,1:3]),digits=3,caption = "Estimated coefficients and SE for part 3 model")%>%
   kable_styling(position = "center",latex_options = "HOLD_position")
@@
 #Producing table of results
-knitr::kable(as.data.frame(tidy(lm.log.income)[,1:3]),digits=3)%>%
+knitr::kable(as.data.frame(tidy(lm.log.income)[,1:3]),digits=3,caption = "Estimated coefficients and SE for part 4 model")%>%
   kable_styling(position = "center",latex_options = "HOLD_position")
@@
 #Producing table of results
-knitr::kable(as.data.frame(tidy(lm.log.income.sqd)[,1:3]),digits=3)%>%
+knitr::kable(as.data.frame(tidy(lm.log.income.sqd)[,1:3]),digits=3,caption = "Estimated coefficients and SE for part 5 model")%>%
   kable_styling(position = "center",latex_options = "HOLD_position")
@@
 
-```{r,echo=F,fig.height=8}
+```{r,echo=F,fig.height=6}
 #Creating diagnostic plots
@@
 
-\newpage
 ## Part 6: Comparing models
@@
 Therefore, their effect on expenditure is negligible so I remove them from the model as a simpler model is preferred.
-After removing these 3 variables, removing any more does not further decrease the AIC thus this simplified model is my final model.
+After removing these 3 variables, removing any more does not further decrease the AIC, furthermore the difference in AIC is small with the simple models AIC being **1406.732** compared to the more complicated model with an AIC of **1386.878**.
+Thus this simplified model is my final model.
 ```{r,include=F}
@@
 lm.final.model <- lm.simple.model.1
+
+AIC(lm.final.model)
 ```
-
+Producing diagnostic plots for this final model (seen below) shows no issues with the model fit.
+The residuals display a random scattering of points indicating the linearity assumption has been met and the residuals variance does not appear to increase/decrease except for a few values but I don't think this is enough evidence to say that the homoscedasticity assumption is invalid.
+This is further backed up by a straight nearly horizontal red line in the scale-location plot with an even random scattering of residuals around it.
+The points follow a straight line in the QQ plot with the tails deviating slightly but not extreme enough to indicate the normality assumption is invalid.
+The cooks distance plot shows no influential points are present in the data.
 ```{r,echo=F,fig.height=6}
@@
 
-Producing diagnostic plots for this final model shows no issues with the model fit.
-The residuals display a random scattering of points indicating the linearity assumption has been met and the residuals variance does not appear to increase/decrease except for a few values but I don't think this is enough evidence to say that the homoscedasticity assumption is invalid.
-This is further backed up by a straight nearly horizontal red line in the scale-location plot with an even random scattering of residuals around it.
-The points follow a straight line in the QQ plot with the tails deviating slightly but not extreme enough to indicate the normality assumption is invalid.
-The cooks distance plot shows no influential points are present in the data.
+## Part 8: Inference of final model
+Let $y$ denote the expenditure of the household, $x$ the income of the household, $\alpha^{G}$ the sex of the household head, $\alpha^{S}$ the size of the household, $\alpha^{T}$ the tenure type, $\alpha^{L}$ the labour force and $\alpha^{SG}$ the interaction between sex of household head and size of household.
+Then the final model is:
+\begin{equation*}
+log(y_{ijkwz})=\beta_{0}+\beta_{1}x_{ijkwz}+\beta_{2}x_{ijkwz}^{2}+\alpha_{j}^{G}+\alpha_k^{S}+\alpha_{w}^{T}+\alpha_{z}^{L}+\alpha^{SG}_{jk}+\epsilon_{ijkwz}
+\end{equation*}
+Where $\alpha_{iM}^{G}=\alpha_{i1}^{S}=\alpha_{iPriv}^{T}=\alpha_{iFT}^{L}=\alpha^{SG}_{iMk}=\alpha^{SG}_{ij1}=0$ where $j \in \{M,F\}, k \in \{1,2,3,4,5+\}, w \in \{Pub,Priv,Own\}, z \in \{FT,PT,U\}$ and $i \in \{1,...,n_{jkwz}\}$ with $n_{jkwz}$ number households with the $j^{th}$ sex of household head, $k^{th}$ number of persons in the household, $w^{th}$ tenure type and $z^{th}$ labour force type.
+\newline
 
-## Part 8: Inference of final model
-The final model is $log(y_{ijkwz})=\beta_{0}+\beta_{1}x_{ijkwz}+\beta_{2}x_{ijkwz}^{2}+\alpha_{j}^{G}+\alpha_k^{S}+\alpha_{w}^{T}+\alpha_{z}^{L}+\alpha^{SG}_{jk}+\epsilon_{ijkwz}$.
+Due to the model being a log-linear model the explanatory variables have a multiplicative effect on the response.
+Exponentiating the estimated coefficients to find their effect on expenditure gives the below table.
+I note that all values have been rounded to 3 decimal places thus the zero p-values are due to p-values being too small be displayed and similarly the effect of income squared is too small to be shown.
+My follwoing inference will be based on this table.
+
+\newpage
+
+```{r,echo=F}
+#Calculating confidence intervals
+
+lwr <-exp(tidy(lm.final.model)$estimate+c(-1)*qnorm(0.975)*tidy(lm.final.model)$std.error)
+upr <- exp(tidy(lm.final.model)$estimate+c(1)*qnorm(0.975)*tidy(lm.final.model)$std.error)
+
+Point.estimates <- round(exp(tidy(lm.final.model)[,2]),3)
+p.values <- round(tidy(lm.final.model)[,5],3)
+
+#Producing tables to display result
+knitr::kable(data.frame("coefficent"=tidy(lm.final.model)[,1],
+                 Point.estimates,
+                 "Confidence Interval 95%"=c(paste("(",round(lwr,3),", ",round(upr,3),")")),
+                 p.values),digits=3,caption = "Summary of exponentiated variables affecting expeniture")%>%
+  kable_classic(full_width=F,latex_options = "HOLD_position",position= "center")
+
+```
+First, interpreting intercept value, it is for a male headed household of 1 person, publicly rented in full time employment who has an income of £0.
+Clearly this does not make sense.
+The difference between full time work and being unemployed was highly statistically significant.
+With those unemployed having an associated decrease in expenditure of (*8.5%*,*34.1%*) 95% CI compared to those in full time work with all other variables held constant.
+However the difference between full time work and being inactive or in part time work was not significant (CI included 1) therefore with all other variables constant there is no evidence that being in one of these 3 factors compared to another influences the expenditure of the household.
+
+The difference between publicly rented properties expenditure and either privately rented or owned properties expenditure were statistically significant.
+With the associated increase in expenditure of (*22%*,*44.1%*) and (*11.5%*,*28.1%*) 95% CI's respectively with all other variables held constant.
+
+Due to squared income squared the effect of an increase in income on expenditure depends on the initial value of income.
+Assuming all other variables are held constant the percentage increase o expenditure is $(\frac{exp(\beta_1(x+y)+\beta_2(x+y)^2)}{exp(\beta_1x+\beta_2x^2)}-1)*100$ where x is the initial value of income and y is the increase in income, $\beta_1=0.002285$ and $\beta_2=-0.000001036$.
+Thus, for a household earning the mean income of £512, a £20 increase in income increases expenditure by approximately *2.4%* whereas for a household earning only £300 the same £20 increase, increases expenditure by approximately *3.3%*.
+Due to the negative quadratic term the higher the initial income the less of a positive multiplicative effect an increase in expenditure has on expenditure.
+
+```{r,include=F}
+#calculating percent increase
+beta_1 <- 0.002285
+beta_2 <- -0.000001036
+x <- 300
+inc <- 20
+
+(exp(beta_1*(x+inc)+beta_2*(x+inc)^2)/exp(beta_1*x+beta_2*x^2)-1)*100
+```
+
+The final 2 variables, sex of household head and household size have an interaction between them.
+However, only the household size of 3 level has a statistically significant interaction with sex of household head.
+Therefore as the other levels
+
+The sex of the household head is highly statistically significant with being female head increasing the expenditure by (*5.6%*,*25.9%*) 95% CI compared to a male head with all other variables held constant.
+Household size is extremely statistically significant with an increase in expenditure of (*22.4%*,*45.5*),  (*31.3%*,*66.8%*), (*34.9%*,*81.6%*) 95% CI's for households of size 2,4,5+ when compared to a household of size 1 with all other variables held constant.
+For the effect on expenditure between a household of size 1 and size 3 because of the significant interaction with sex if the household head is male the CI is (*36.2%*,*71.9%*) and (*-15.7%*,*+52.3%*) if female is it is with all other variables held constant.
+Note that the for the female head CI 0% is included therefore there is no statistically significant difference in expenditure between a household of size 1 with male head and a household of size 3 with a female head with all other variables held constant.
+
+Overall an increase in income, higher household size, being a female head and living in a owned or privately rented increase the % expenditure when compared to a single male headed household living in a publicly rented house (unless a female head with size 3 household).
+Whereas being unemployed compared to in full time work decreases the % of expenditure.
+
+All the variables mentioned above were statistically significant in their association with expenditure.
+However, just because they are significant, it does not mean they influence expenditure.
+There could be confounding or simple chance causing these results.
+However, some areas of confidence are:
+
+* income, logically if you make more money you have more available to spend 
+* household size, larger households are more expensive thus require more spending

### 2022-11-18 Draft_final.Rmd
Resizing and fitting into 9 pages and adding plot to explain effect of income

@@
 I note that these differences and possible associations may not be due to the variable mentioned but due to a third variable (a confounder) that affects both the categorical variable and expenditure (the response).
+\vspace{-5truemm}
 
@@
 Fitting the model $y_{i}=\beta_{0}+\beta_{1}x_{i}+\epsilon$ where $y$ is the weekly household expenditure and the explanatory variable $x$ is weekly income and $i \in \{1,...,1200\}$ is the ith household in the dataset produces the below table (these are the same for all linear models mentioned in parts 2 to 6).It shows the estimated values of the intercept ($\beta_0$) and the estimated 'effect' of a 1 pound increase of income on expenditure ($\beta_1$) and their associated standard errors.
+\vspace{-5truemm}
 ```{r,echo=F}
@@
 knitr::kable(as.data.frame(tidy(lm.income)[,1:3]),digits=3,caption = "Estimated coefficients and SE for part 2 model")%>%
+  row_spec(0,bold=T)%>%
   kable_styling(position = "center",latex_options = "HOLD_position")
 ```
-
+\vspace{-10truemm}
 ```{r,echo=F,fig.height=3.8}
@@
 ```
-\newpage
 
@@
 This implies the normality assumption is invalid.
+\newpage
 
@@
 knitr::kable(as.data.frame(tidy(lm.income.sqd)[,1:3]),digits=3,caption = "Estimated coefficients and SE for part 3 model")%>%
+  row_spec(0,bold=T)%>%
   kable_styling(position = "center",latex_options = "HOLD_position")
 ```
+Looking at the produced diagnostic plots below for this model, in the fitted vs residual plot the residual values increase as the fitted values increase thus the homoscedasticity assumption may not be valid.
+The tails in the normal QQ plot also still deviate heavily away from the straight line meaning the normality assumption is invalid.
 
-```{r,echo=F,fig.height=4.2}
+```{r,echo=F,fig.height=3.8}
 #Creating diagnostic plots
@@
 ```
-
-Looking at the produced diagnostic plots above for this model, in the fitted vs residual plot the residual values increase as the fitted values increase thus the homoscedasticity assumption may not be valid.
-The tails in the normal QQ plot also still deviate heavily away from the straight line meaning the normality assumption is invalid.
-
-\newpage
 
@@
 knitr::kable(as.data.frame(tidy(lm.log.income)[,1:3]),digits=3,caption = "Estimated coefficients and SE for part 4 model")%>%
+  row_spec(0,bold=T)%>%
   kable_styling(position = "center",latex_options = "HOLD_position")
 ```
+Looking at the produced diagnostic plots above for this model, the data appears to be heteroscedastic as in the fitted vs residual plot the variance in the residuals decreases as the fitted values increases.
+Furthermore, the linear assumption also appears violated as the residuals appear to exhibit a slight negative quadratic shape.
+The QQ plot shows the residuals stay close to the straight line thus normality assumption is valid.
 
-```{r,echo=F,fig.height=4.2}
+\vspace{-25truemm}
+```{r,echo=F,fig.height=3.8}
 #Creating diagnostic plots
@@
 ```
-
-Looking at the produced diagnostic plots above for this model, the data appears to be heteroscedastic as in the fitted vs residual plot the variance in the residuals decreases as the fitted values increases.
-Furthermore, the linear assumption also appears violated as the residuals appear to exhibit a slight negative quadratic shape.
-The QQ plot shows the residuals stay close to the straight line thus normality assumption is valid.
+\vspace{-10truemm}
 
 ## Part 5: Regressing the natural logarithm of expenditure on income and income squared
+\vspace{-2truemm}
+
 Fitting the model $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\beta_{2}x^2_{i}+\epsilon$ gives the estimated coefficients and standard errors in the table below.
-
+\vspace{-5truemm}
 ```{r,echo=F}
@@
 knitr::kable(as.data.frame(tidy(lm.log.income.sqd)[,1:3]),digits=3,caption = "Estimated coefficients and SE for part 5 model")%>%
+  row_spec(0,bold=T)%>%
   kable_styling(position = "center",latex_options = "HOLD_position")
 ```
+\vspace{-15truemm}
 
-```{r,echo=F,fig.height=6}
+```{r,echo=F,fig.height=5}
 #Creating diagnostic plots
@@
 
-Producing the diagnostic plots above, the fitted vs residuals plot shows a fairly even scattering of points around 0 and a linear trend, however the variance of the residuals appears to decrease slightly as the fitted values increase.
+Producing the diagnostic plots below, the fitted vs residuals plot shows a fairly even scattering of points around 0 and a linear trend, however the variance of the residuals appears to decrease slightly as the fitted values increase.
 But, looking at the scale-location plot the residuals appear randomly spread around the red line and the red line through them stays fairly horizontal thus homoskedasticity assumptions appears valid.
@@
 ## Part 6: Comparing models
-Comparing the models from part 2-5 using AIC and adjusted $R^2$:
-
 ```{r,echo=F}
@@
 
-knitr::kable(Model.comparison.data,digits=3)%>%
+knitr::kable(Model.comparison.data,digits=3,caption = "Comparing the AIC and adjusted R squared of models from part 2-5")%>%
+  row_spec(0,bold=T)%>%
   kable_styling(position = "center",latex_options = "HOLD_position")
@@
 knitr::kable(as.data.frame(tidy(lm.log.income.sqd)[,1:3]),digits=3,caption = "Estimated coefficients and SE for part 5 model")%>%
+  row_spec(0,bold=T)%>%
   kable_styling(position = "center",latex_options = "HOLD_position")
@@
 knitr::kable(data.frame(tidy(lm.full.1.deletion)[14:22,1],exp(tidy(lm.full.1.deletion)[14:22,2])),digits=3)%>%
+  row_spec(0,bold=T)%>%
   kable_styling(position = "center",latex_options = "HOLD_position")
@@
 Therefore, their effect on expenditure is negligible so I remove them from the model as a simpler model is preferred.
-After removing these 3 variables, removing any more does not further decrease the AIC, furthermore the difference in AIC is small with the simple models AIC being **1406.732** compared to the more complicated model with an AIC of **1386.878**.
+After removing these 3 variables, removing any more does not further decrease the AIC.
+Furthermore the difference in AIC between the simpler and more complicated model with the simple models AIC being **1406.732** compared to the more complicated model AIC of **1386.878**.
 Thus this simplified model is my final model.
+
 ```{r,include=F}
@@
 ```
+
 Producing diagnostic plots for this final model (seen below) shows no issues with the model fit.
@@
 The cooks distance plot shows no influential points are present in the data.
+
 ```{r,echo=F,fig.height=6}
@@
 \end{equation*}
-Where $\alpha_{iM}^{G}=\alpha_{i1}^{S}=\alpha_{iPriv}^{T}=\alpha_{iFT}^{L}=\alpha^{SG}_{iMk}=\alpha^{SG}_{ij1}=0$ where $j \in \{M,F\}, k \in \{1,2,3,4,5+\}, w \in \{Pub,Priv,Own\}, z \in \{FT,PT,U\}$ and $i \in \{1,...,n_{jkwz}\}$ with $n_{jkwz}$ number households with the $j^{th}$ sex of household head, $k^{th}$ number of persons in the household, $w^{th}$ tenure type and $z^{th}$ labour force type.
+Where $\alpha_{M}^{G}=\alpha_{1}^{S}=\alpha_{Priv}^{T}=\alpha_{FT}^{L}=\alpha^{SG}_{Mk}=\alpha^{SG}_{j1}=0$, $j \in \{M,F\}, k \in \{1,2,3,4,5+\}, w \in \{Pub,Priv,Own\}, z \in \{FT,PT,U,I\}$ and $i \in \{1,...,n_{jkwz}\}$ with $n_{jkwz}$ number households with the $j^{th}$ sex of household head, $k^{th}$ number of persons in the household, $w^{th}$ tenure type and $z^{th}$ labour force type.
 \newline
 
-Due to the model being a log-linear model the explanatory variables have a multiplicative effect on the response.
-Exponentiating the estimated coefficients to find their effect on expenditure gives the below table.
+Due to the model being a log-linear model the explanatory variables have a multiplicative effect on the weekly expenditure.
+Exponentiating the estimated coefficients to find their effect on expenditure gives the below [*Table 7*].
 I note that all values have been rounded to 3 decimal places thus the zero p-values are due to p-values being too small be displayed and similarly the effect of income squared is too small to be shown.
-My follwoing inference will be based on this table.
-
-\newpage
+My following inference will be based on this table and I use a significance level of 5% for all results.
 
@@
                  p.values),digits=3,caption = "Summary of exponentiated variables affecting expeniture")%>%
+  row_spec(0,bold=T)%>%
   kable_classic(full_width=F,latex_options = "HOLD_position",position= "center")
@@
 ```
-First, interpreting intercept value, it is for a male headed household of 1 person, publicly rented in full time employment who has an income of £0.
-Clearly this does not make sense.
+First, interpreting the intercept value, it is for a male headed household of 1 person in full time employment in a publicly rented property who has an income of £0.
+Clearly this does not make sense (having a job which pays no income).
+
 The difference between full time work and being unemployed was highly statistically significant.
-With those unemployed having an associated decrease in expenditure of (*8.5%*,*34.1%*) 95% CI compared to those in full time work with all other variables held constant.
-However the difference between full time work and being inactive or in part time work was not significant (CI included 1) therefore with all other variables constant there is no evidence that being in one of these 3 factors compared to another influences the expenditure of the household.
+With unemployed households compared to full time employment households having a 95% CI of [0.659, 0.915] which corresponds to a **8.5%** to **34.1%** decrease in expenditure with all other variables held constant.
+
+However the difference between full time work and being inactive or in part time work was not significant (the 95% CI included 1) therefore with all other variables constant there is no evidence that being in one of these 3 factors compared to another of influences the expenditure of the household.
 
 The difference between publicly rented properties expenditure and either privately rented or owned properties expenditure were statistically significant.
-With the associated increase in expenditure of (*22%*,*44.1%*) and (*11.5%*,*28.1%*) 95% CI's respectively with all other variables held constant.
+With the associated 95% CI of [1.22, 1.441] and [1.115 , 1.281] which corresponds to a **22%** to **44.1%** and a **11.5%** **28.1%** increase in expenditure respectively with all other variables held constant.
 
-Due to squared income squared the effect of an increase in income on expenditure depends on the initial value of income.
-Assuming all other variables are held constant the percentage increase o expenditure is $(\frac{exp(\beta_1(x+y)+\beta_2(x+y)^2)}{exp(\beta_1x+\beta_2x^2)}-1)*100$ where x is the initial value of income and y is the increase in income, $\beta_1=0.002285$ and $\beta_2=-0.000001036$.
-Thus, for a household earning the mean income of £512, a £20 increase in income increases expenditure by approximately *2.4%* whereas for a household earning only £300 the same £20 increase, increases expenditure by approximately *3.3%*.
-Due to the negative quadratic term the higher the initial income the less of a positive multiplicative effect an increase in expenditure has on expenditure.
-
+Due to income having a squared term the effect of an increase in income on expenditure depends on the initial value of income.
+To demonstrate this I produced the plot below showing how the percent increase in expenditure varies with the initial income for several different increases in income.
+For all but the most extreme weekly incomes (over £1000) income has a positive effect on expenditure.
 ```{r,include=F}
@@
 beta_2 <- -0.000001036
-x <- 300
+x <- 900
 inc <- 20
@@
 (exp(beta_1*(x+inc)+beta_2*(x+inc)^2)/exp(beta_1*x+beta_2*x^2)-1)*100
+
+#Finding max value of icome in dataset
+max(expend.df$income)
+#max income is 1184.472 (approx 1200)
+
+#Creating plot to show change in percent of 
+x <- seq(from=0,to=1200,by=10)
+inc <- 20
+expend.inc.20 <- (exp(beta_1*(x+inc)+beta_2*((x+inc)^2))/exp(beta_1*x+beta_2*x^2)-1)*100
+inc<-50
+expend.inc.50 <- (exp(beta_1*(x+inc)+beta_2*((x+inc)^2))/exp(beta_1*x+beta_2*x^2)-1)*100
+
+test <- data.frame("Init.inc"=c(x,x),
+                   "Inc.expend"=c(expend.inc.20,expend.inc.50),
+                   "Inc.inc"=c(rep("£20",121),rep("£50",121)))%>%
+  mutate(Inc.inc=factor(Inc.inc,levels = c("£20","£50"),labels=c("£20","£50")))
+
+ggplot(aes(Init.inc,Inc.expend,colour=Inc.inc),data=test)+
+  geom_line()+
+  ylab("Increase in expenditure (%)")+
+  xlab("Initial income (£)")
+```
+
+```{r,echo=F,fig.height=3}
+ggplot(aes(Init.inc,Inc.expend,colour=Inc.inc),data=test)+
+  geom_line()+
+  ylab("Increase in expenditure (%)")+
+  xlab("Initial income (£)")+
+  ggtitle("Effect of income on expenditure")
 ```

### 2022-11-18 Draft_final.Rmd
finishing off task 8 inference

@@
 
-However the difference between full time work and being inactive or in part time work was not significant (the 95% CI included 1) therefore with all other variables constant there is no evidence that being in one of these 3 factors compared to another of influences the expenditure of the household.
+However the difference between full time work and being inactive or in part time work was not significant (the 95% CI included 1) therefore with all other variables constant there is no evidence that being in one of these 3 factors compared to another influences the expenditure of the household.
 
-The difference between publicly rented properties expenditure and either privately rented or owned properties expenditure were statistically significant.
+The difference between a households expenditure for households which live in a publicly rented property and either a privately rented or owned property were statistically significant.
 With the associated 95% CI of [1.22, 1.441] and [1.115 , 1.281] which corresponds to a **22%** to **44.1%** and a **11.5%** **28.1%** increase in expenditure respectively with all other variables held constant.
@@
 Due to income having a squared term the effect of an increase in income on expenditure depends on the initial value of income.
-To demonstrate this I produced the plot below showing how the percent increase in expenditure varies with the initial income for several different increases in income.
-For all but the most extreme weekly incomes (over £1000) income has a positive effect on expenditure.
+To demonstrate this I produced the plot below showing how the percent increase in expenditure varies with the initial income of a household for several different increases in household income.
+For all but the most extreme household incomes (over £1000), income has a positive effect on expenditure with the higher the initial income the less of a positive effect on expenditure an increase in income has (seen by negative gradient in below plot and which is due to the negative income squared regression coefficient).
 ```{r,include=F}
@@
 
-#Finding max value of icome in dataset
+#Finding max value of income in dataset
 max(expend.df$income)
@@
 
-test <- data.frame("Init.inc"=c(x,x),
+income.affect.df <- data.frame("Init.inc"=c(x,x),
                    "Inc.expend"=c(expend.inc.20,expend.inc.50),
@@
 
-ggplot(aes(Init.inc,Inc.expend,colour=Inc.inc),data=test)+
-  geom_line()+
-  ylab("Increase in expenditure (%)")+
-  xlab("Initial income (£)")
 ```
 
-```{r,echo=F,fig.height=3}
-ggplot(aes(Init.inc,Inc.expend,colour=Inc.inc),data=test)+
+```{r,echo=F,fig.height=4}
+#Plotting affect of income on expenditure
+
+ggplot(aes(Init.inc,Inc.expend,colour=Inc.inc),data=income.affect.df)+
   geom_line()+
@@
   xlab("Initial income (£)")+
-  ggtitle("Effect of income on expenditure")
+  labs(colour = "Increase in\n income amount") +
+  ggtitle("Affect of income on expenditure using final model")+
+  theme(legend.position="bottom",
+        legend.title=element_text(size=8))
 ```
@@
 However, only the household size of 3 level has a statistically significant interaction with sex of household head.
-Therefore as the other levels
+Therefore as there is no evidence the other levels of the interaction are different from the reported associations without the interaction included I will not report them.
 
-The sex of the household head is highly statistically significant with being female head increasing the expenditure by (*5.6%*,*25.9%*) 95% CI compared to a male head with all other variables held constant.
-Household size is extremely statistically significant with an increase in expenditure of (*22.4%*,*45.5*),  (*31.3%*,*66.8%*), (*34.9%*,*81.6%*) 95% CI's for households of size 2,4,5+ when compared to a household of size 1 with all other variables held constant.
-For the effect on expenditure between a household of size 1 and size 3 because of the significant interaction with sex if the household head is male the CI is (*36.2%*,*71.9%*) and (*-15.7%*,*+52.3%*) if female is it is with all other variables held constant.
-Note that the for the female head CI 0% is included therefore there is no statistically significant difference in expenditure between a household of size 1 with male head and a household of size 3 with a female head with all other variables held constant.
+The sex of the household head is highly statistically significant with being female head having a 95% CI of [1.056, 1.259] which corresponds to a **5.6%** to **25.9%**  increase the expenditure compared to a male head with all other variables held constant (and for a household not of size 3).
+Household size is also extremely statistically significant with a 95% CI of [1.224, 1.455], [1.313, 1.668] and [1.349, 1.816] for households of size 2, 4 and 5+ respectively when compared to a household of size 1 with all other variables held constant which corresponds to a **22.4%** to **45.5%**, a **31.3%** to **66.8%** and a **34.9%** to **81.6%** respectively.
 
-Overall an increase in income, higher household size, being a female head and living in a owned or privately rented increase the % expenditure when compared to a single male headed household living in a publicly rented house (unless a female head with size 3 household).
-Whereas being unemployed compared to in full time work decreases the % of expenditure.
+For the effect on expenditure between a household of size 1 and size 3 because of the significant interaction with sex of household head if the household head is male the 95% CI is [1.362, 1.719] and [0.912, 1.920] if female with all other variables held constant.
+Note that for the female head 1 is included in the CI, therefore there is no evidence of a difference in expenditure between a household of size 1 with a female head and a household of size 3 with a female head with all other variables held constant.
+However for a for a male household head, a household of size 3 compared to size 1 corresponds to a **36.2%** to **71.9%** increase in expenditure with all other variables held constant.
 
-All the variables mentioned above were statistically significant in their association with expenditure.
-However, just because they are significant, it does not mean they influence expenditure.
-There could be confounding or simple chance causing these results.
-However, some areas of confidence are:
+Overall an increase in income (except for extremely high incomes), higher household size, being household with a female head and living in a owned or privately rented increase the expenditure of the household (compared to a single male headed household living in a publicly rented house, unless a female head with size 3 household).
+Whereas being unemployed compared to in full time work decreases the expenditure of the household.
 
-* income, logically if you make more money you have more available to spend 
-* household size, larger households are more expensive thus require more spending
+I note that even though the variables mentions above had a  statistically significant association with expenditure this could be spurious and due to confounding from variables not included in the model, simple chance or bias from the sample the model was based on.
+However, some areas of confidence of causation due to logical arguments are:
+
+* Income, logically if you make more money you have more available to spend thus increasing expenditure
+* Household size, larger households are more expensive thus require more spending
+* Labour force, if you are unemployed compared to having work you are likely to be more frugal and thus reduce expenditure
+
+Further investigation may be required for tenure type as it could be other variables which cause a difference in tenure are responsible.

### 2022-11-19 Draft_final.Rmd
Creating table for CI of sex and household

@@
 
+```{r,echo=F}
+#Creating table for effect of sex and household size
+sum.final.model <- summary(lm.final.model)
+
+#Vector of point estimate of household size
+house.size.effect <- c(0,
+                       sum.final.model$coefficients[9,1],
+                       sum.final.model$coefficients[10,1],
+                       sum.final.model$coefficients[11,1],
+                       sum.final.model$coefficients[12,1])
+#Vector of point estimate of effect of household size and female interaction
+house.size.int <- c(0,
+                       sum.final.model$coefficients[14,1],
+                       sum.final.model$coefficients[15,1],
+                       sum.final.model$coefficients[16,1],
+                       sum.final.model$coefficients[17,1])
+#vecotr of standard error of household size
+house.size.effect.std <- c(0,
+                       sum.final.model$coefficients[9,2],
+                       sum.final.model$coefficients[10,2],
+                       sum.final.model$coefficients[11,2],
+                       sum.final.model$coefficients[12,2])
+#vector of standard error of interation of household size and female
+house.size.int.std <- c(0,
+                       sum.final.model$coefficients[14,2],
+                       sum.final.model$coefficients[15,2],
+                       sum.final.model$coefficients[16,2],
+                       sum.final.model$coefficients[17,2])
+#lower and upper 95% CI bound for household size
+house.size.lwr <- house.size.effect-1.96*house.size.effect.std
+house.size.upr <- house.size.effect+1.96*house.size.effect.std
+#lower and upper 95% CI bound for household and female interaction
+house.size.int.lwr <- house.size.int-1.96*house.size.int.std
+house.size.int.upr <- house.size.int+1.96*house.size.int.std
+
+#lower and upper CI bound for being female and of household size [i]
+female.house.lwr <- house.size.lwr+house.size.int.lwr+(sum.final.model$coefficients[13,1]-1.96*sum.final.model$coefficients[13,2])
+female.house.upr <- house.size.upr+house.size.int.upr+(sum.final.model$coefficients[13,1]+1.96*sum.final.model$coefficients[13,2])
+
+#Creating vector of 95% CI for male and female household head with [ith] household size
+male.CI <- paste0("(",round(exp(house.size.lwr),3), ", ",round(exp(house.size.upr),3),")")
+female.CI <- paste0("(",round(exp(female.house.lwr),3), ", ",round(exp(female.house.upr),3),")")
+
+data.sex.household.CI <- data.frame(t(data.frame("Male"=male.CI, "Female"=female.CI)))%>%
+  rename("1 Person"="X1",
+         "2 Persons"="X2",
+         "3 Persons"="X3",
+         "4 Persons"="X4",
+         "5+ Persons"="X5")
+
+#creating table of 95% CI calculated
+knitr::kable(data.sex.household.CI,digits=3,caption = "95% CI compared to male headed household of 1 person")%>%
+  row_spec(0,bold=T)%>%
+  column_spec(1,bold=T)%>%
+  kable_styling(position = "center",latex_options = "HOLD_position")
+```
+
 For the effect on expenditure between a household of size 1 and size 3 because of the significant interaction with sex of household head if the household head is male the 95% CI is [1.362, 1.719] and [0.912, 1.920] if female with all other variables held constant.

### 2022-11-20 Draft_final.Rmd
Fixing error with final kable output table

@@
 
-```{r,echo=F}
-#Producing table of results
-knitr::kable(as.data.frame(tidy(lm.log.income.sqd)[,1:3]),digits=3,caption = "Estimated coefficients and SE for part 5 model")%>%
-  row_spec(0,bold=T)%>%
-  kable_styling(position = "center",latex_options = "HOLD_position")
-```
-
 The log-linear model means that the covarites (income and income squared) have a multiplicative effect on the response (expenditure).
@@
 #Producing table of all exponentiated regession coefficients that are approx 1
-knitr::kable(data.frame(tidy(lm.full.1.deletion)[14:22,1],exp(tidy(lm.full.1.deletion)[14:22,2])),digits=3)%>%
+knitr::kable(data.frame(tidy(lm.full.1.deletion)[14:22,1],exp(tidy(lm.full.1.deletion)[14:22,2])),digits=3,caption="Coefficents of explantory variables with negliable affect on expenditure",)%>%
   row_spec(0,bold=T)%>%
-  kable_styling(position = "center",latex_options = "HOLD_position")
+  kable_styling(position = "center",latex_options = "HOLD_position", font_size = 7)
 ```
@@
   row_spec(0,bold=T)%>%
-  kable_classic(full_width=F,latex_options = "HOLD_position",position= "center")
+  kable_styling(full_width=F,latex_options = "HOLD_position",position= "center")
 
@@
 
-data.sex.household.CI <- data.frame(t(data.frame("Male"=male.CI, "Female"=female.CI)))%>%
+#creating table of 95% CI calculated
+data.frame(t(data.frame("Male"=male.CI, "Female"=female.CI)))%>%
   rename("1 Person"="X1",
@@
          "4 Persons"="X4",
-         "5+ Persons"="X5")
+         "5+ Persons"="X5")%>%
+  kable(caption = "95% CI compared to male headed household of 1 person",format = 'pandoc')
+```
 
-#creating table of 95% CI calculated
-knitr::kable(data.sex.household.CI,digits=3,caption = "95% CI compared to male headed household of 1 person")%>%
-  row_spec(0,bold=T)%>%
-  column_spec(1,bold=T)%>%
-  kable_styling(position = "center",latex_options = "HOLD_position")
+```{r,include=F}
+# #creating table of 95% CI calculated
+# kable(data.frame(data.sex.household.CI), caption = "95% CI compared to male headed household of 1 person")%>%
+#   row_spec(0,bold=T)%>%
+#   column_spec(1,bold=T)%>%
+#   kable_classic(position = "center",latex_options = "HOLD_position")
+# 
+# knitr::kable(data.frame(tidy(lm.full.1.deletion)[14:22,1],exp(tidy(lm.full.1.deletion)[14:22,2])),digits=3)%>%
+#   row_spec(0,bold=T)%>%
+#   kable_styling(position = "center",latex_options = "HOLD_position")
 ```

### 2022-11-20 Draft_final.Rmd
Updating inference on sex and household size (simplifing as now have table)

@@
 ## Part 4: Regressing the natural logarithm of expenditure on income
-Fitting the same model as in task 2 but with the response of expenditure logged (a log-linear model) $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\epsilon$ gives the below estimated coefficients and standard errors.
+Fitting the same model as in task 2 but with the log of the response expenditure (a log-linear model) $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\epsilon$ gives the below estimated coefficients and standard errors.
 
@@
 
-Due to income having a squared term the effect of an increase in income on expenditure depends on the initial value of income.
-To demonstrate this I produced the plot below showing how the percent increase in expenditure varies with the initial income of a household for several different increases in household income.
-For all but the most extreme household incomes (over £1000), income has a positive effect on expenditure with the higher the initial income the less of a positive effect on expenditure an increase in income has (seen by negative gradient in below plot and which is due to the negative income squared regression coefficient).
-```{r,include=F}
-#calculating percent increase
-beta_1 <- 0.002285
-beta_2 <- -0.000001036
-x <- 900
-inc <- 20
-
-(exp(beta_1*(x+inc)+beta_2*(x+inc)^2)/exp(beta_1*x+beta_2*x^2)-1)*100
-
-#Finding max value of income in dataset
-max(expend.df$income)
-#max income is 1184.472 (approx 1200)
-
-#Creating plot to show change in percent of 
-x <- seq(from=0,to=1200,by=10)
-inc <- 20
-expend.inc.20 <- (exp(beta_1*(x+inc)+beta_2*((x+inc)^2))/exp(beta_1*x+beta_2*x^2)-1)*100
-inc<-50
-expend.inc.50 <- (exp(beta_1*(x+inc)+beta_2*((x+inc)^2))/exp(beta_1*x+beta_2*x^2)-1)*100
-
-income.affect.df <- data.frame("Init.inc"=c(x,x),
-                   "Inc.expend"=c(expend.inc.20,expend.inc.50),
-                   "Inc.inc"=c(rep("£20",121),rep("£50",121)))%>%
-  mutate(Inc.inc=factor(Inc.inc,levels = c("£20","£50"),labels=c("£20","£50")))
-
-```
-
-```{r,echo=F,fig.height=4}
-#Plotting affect of income on expenditure
-
-ggplot(aes(Init.inc,Inc.expend,colour=Inc.inc),data=income.affect.df)+
-  geom_line()+
-  ylab("Increase in expenditure (%)")+
-  xlab("Initial income (£)")+
-  labs(colour = "Increase in\n income amount") +
-  ggtitle("Affect of income on expenditure using final model")+
-  theme(legend.position="bottom",
-        legend.title=element_text(size=8))
-```
-
-The final 2 variables, sex of household head and household size have an interaction between them.
-However, only the household size of 3 level has a statistically significant interaction with sex of household head.
-Therefore as there is no evidence the other levels of the interaction are different from the reported associations without the interaction included I will not report them.
-
-The sex of the household head is highly statistically significant with being female head having a 95% CI of [1.056, 1.259] which corresponds to a **5.6%** to **25.9%**  increase the expenditure compared to a male head with all other variables held constant (and for a household not of size 3).
-Household size is also extremely statistically significant with a 95% CI of [1.224, 1.455], [1.313, 1.668] and [1.349, 1.816] for households of size 2, 4 and 5+ respectively when compared to a household of size 1 with all other variables held constant which corresponds to a **22.4%** to **45.5%**, a **31.3%** to **66.8%** and a **34.9%** to **81.6%** respectively.
+Sex of household head and household size have an interaction between them therefore cannot be reported separately.
+To easily display the effects of both on expenditure I created the below table which displays the multiplicative effect of sex and household size on expenditure.
 
@@
 #creating table of 95% CI calculated
-data.frame(t(data.frame("Male"=male.CI, "Female"=female.CI)))%>%
+data.frame(t(data.frame("Male head"=male.CI, "Female head"=female.CI)))%>%
   rename("1 Person"="X1",
@@
          "5+ Persons"="X5")%>%
-  kable(caption = "95% CI compared to male headed household of 1 person",format = 'pandoc')
+  kable(caption = "95% CI of multiplcative effect on expenditure compared to male headed household of 1 person",format = 'pandoc')
+
+table(expend.df$sex.hh,expend.df$hh.size)
 ```
 
+For households of size 1,2 and 4 a female household head corresponds to an increase in expenditure compared to a male household of size 1 however, for a household of size 3 or 5+ there is not enough evidence to support a difference in expenditure (CI includes 1).
+I note the confidence intervals are very large thus there is alot of uncertainty in these results and thus the reason for no found effect may be due to insufficient power of thr sample.
+
+Compared to a household of size 1 all larger households increase expenditure if household head is male whereas due to the uncertainty there is no evidence this is the case for female household heads due to overlapping CI.
+
+Due to income having a squared term the effect of an increase in income on expenditure depends on the initial value of income.
+To demonstrate this I produced the plot below showing how the percent increase in expenditure varies with the initial income of a household for several different increases in household income.
+For all but the most extreme household incomes (over £1000), income has a positive effect on expenditure with the higher the initial income the less of a positive effect on expenditure an increase in income has (seen by negative gradient in below plot and which is due to the negative income squared regression coefficient).
 ```{r,include=F}
-# #creating table of 95% CI calculated
-# kable(data.frame(data.sex.household.CI), caption = "95% CI compared to male headed household of 1 person")%>%
-#   row_spec(0,bold=T)%>%
-#   column_spec(1,bold=T)%>%
-#   kable_classic(position = "center",latex_options = "HOLD_position")
-# 
-# knitr::kable(data.frame(tidy(lm.full.1.deletion)[14:22,1],exp(tidy(lm.full.1.deletion)[14:22,2])),digits=3)%>%
-#   row_spec(0,bold=T)%>%
-#   kable_styling(position = "center",latex_options = "HOLD_position")
+#calculating percent increase
+beta_1 <- 0.002285
+beta_2 <- -0.000001036
+x <- 900
+inc <- 20
+
+(exp(beta_1*(x+inc)+beta_2*(x+inc)^2)/exp(beta_1*x+beta_2*x^2)-1)*100
+
+#Finding max value of income in dataset
+max(expend.df$income)
+#max income is 1184.472 (approx 1200)
+
+#Creating plot to show change in percent of 
+x <- seq(from=0,to=1200,by=10)
+inc <- 20
+expend.inc.20 <- (exp(beta_1*(x+inc)+beta_2*((x+inc)^2))/exp(beta_1*x+beta_2*x^2)-1)*100
+inc<-50
+expend.inc.50 <- (exp(beta_1*(x+inc)+beta_2*((x+inc)^2))/exp(beta_1*x+beta_2*x^2)-1)*100
+
+income.affect.df <- data.frame("Init.inc"=c(x,x),
+                   "Inc.expend"=c(expend.inc.20,expend.inc.50),
+                   "Inc.inc"=c(rep("£20",121),rep("£50",121)))%>%
+  mutate(Inc.inc=factor(Inc.inc,levels = c("£20","£50"),labels=c("£20","£50")))
+
 ```
 
-For the effect on expenditure between a household of size 1 and size 3 because of the significant interaction with sex of household head if the household head is male the 95% CI is [1.362, 1.719] and [0.912, 1.920] if female with all other variables held constant.
-Note that for the female head 1 is included in the CI, therefore there is no evidence of a difference in expenditure between a household of size 1 with a female head and a household of size 3 with a female head with all other variables held constant.
-However for a for a male household head, a household of size 3 compared to size 1 corresponds to a **36.2%** to **71.9%** increase in expenditure with all other variables held constant.
+```{r,echo=F,fig.height=4}
+#Plotting affect of income on expenditure
+
+ggplot(aes(Init.inc,Inc.expend,colour=Inc.inc),data=income.affect.df)+
+  geom_line()+
+  ylab("Increase in expenditure (%)")+
+  xlab("Initial income (£)")+
+  labs(colour = "Increase in\n income amount") +
+  ggtitle("Affect of income on expenditure using final model")+
+  theme(legend.position="bottom",
+        legend.title=element_text(size=8))
+```
 
@@
 
-Further investigation may be required for tenure type as it could be other variables which cause a difference in tenure are responsible.
+Further investigation is required for the affect of sex of household head as due to the small sub samples t
+
+may be required for tenure type as it could be other variables which cause a difference in tenure are responsible.

### 2022-11-21 Draft_final.Rmd
Updating inference of final model and conclusion of results

@@
 With unemployed households compared to full time employment households having a 95% CI of [0.659, 0.915] which corresponds to a **8.5%** to **34.1%** decrease in expenditure with all other variables held constant.
+Furthermore, as the CI of unemployed and part-time employment do not overlap there is a statistically significant increase in expenditure if a household is in part time employment compared to unemployed households with all other variables held constant.
 
@@
 
-table(expend.df$sex.hh,expend.df$hh.size)
+#table(expend.df$sex.hh,expend.df$hh.size)
 ```
 
-For households of size 1,2 and 4 a female household head corresponds to an increase in expenditure compared to a male household of size 1 however, for a household of size 3 or 5+ there is not enough evidence to support a difference in expenditure (CI includes 1).
-I note the confidence intervals are very large thus there is alot of uncertainty in these results and thus the reason for no found effect may be due to insufficient power of thr sample.
+For households of size 1,2 and 4 a female household head corresponds to an increase in expenditure compared to a male household of size 1, however for a household of size 3 or 5+ there is not enough evidence to support a difference in expenditure (CI includes 1).
+I note the confidence intervals are very large for female headed households of size 2 and above, this means there is alot of variation in the data causing the uncertainty in the results.
+This could be due to the small sample of female household heads of size 3 and above (insufficient power) or additional confounders affecting female household heads that were not included in the analysis.
 
-Compared to a household of size 1 all larger households increase expenditure if household head is male whereas due to the uncertainty there is no evidence this is the case for female household heads due to overlapping CI.
+Compared to a household of size 1, all larger sized households increase expenditure if the household head is male with an increase of expenditure of **22.4%** to **45.5%**, **36.2%** to **71.9%**, **31.3%** to **66.8%** and **34.9%** to **81.6%** for households of size 2, 3, 4 and 5+ respectively with all other variables held constant.
+There is no evidence larger households increase expenditure for female headed households as the 95% CI's overlap for all household sizes.
 
@@
 
-Overall an increase in income (except for extremely high incomes), higher household size, being household with a female head and living in a owned or privately rented increase the expenditure of the household (compared to a single male headed household living in a publicly rented house, unless a female head with size 3 household).
-Whereas being unemployed compared to in full time work decreases the expenditure of the household.
+Overall an increase in income (except for extremely high incomes), living in a owned or privately rented property compared to a publicly rented property and being in full or part time employment compared to unemployed increases the expenditure of the household.
+With household size increasing income if household head is male and being a female household head increases expenditure for certain household sizes (1, 2 and 4) and having no proven effect for others (3 and 5+).
 
@@
 
-Further investigation is required for the affect of sex of household head as due to the small sub samples t
+Further investigation is required for the affect of sex of household head as due to the small sub samples for some household sizes.
+Furthermore additional factors such as the age of the household head, ethnicity of household and location of household may improve the models accuracy.
 
-may be required for tenure type as it could be other variables which cause a difference in tenure are responsible.
+Also I note that generalization of these results may not be possible as different countries, cultures and ethnicities may have different attitudes on spending.
+As there was no information on where/when the original sample was drawn from and the process for sample selection I cannot recommend what groups/ areas these results may be representative of.

### 2022-11-21 Draft_final.Rmd
Going over write up and changing words for final version

@@
 To assess the distribution of expenditure I produced the above histogram, it shows that expenditure has a positively skewed distribution (and thus the normality assumption used in linear regression may be invalid).
-Using a scatter plot to investigate the relationship between household income and expenditure , it shows there is a clear positive association.
+Using a scatter plot to investigate the relationship between household income and expenditure, there is a clear positive association.
 Then to investigate the other categorical variables I used boxplots to display differences in the expenditure across the different levels (categories) of each variable.
@@
 
-* Different household tenures categories have different weekly expenditure amounts with public rented households having a noticeably lower expenditure compared to the other 2 household levels.
-* Households with a female head have a slightly lower weekly expenditure compared to male household heads
+* Different household tenure categories have different weekly expenditure amounts with public rented households having a noticeably lower expenditure compared to the other 2 household levels.
+* Households with a female household head have a slightly lower weekly expenditure compared to male household heads.
 * Employment status appears to be associated with expenditure with unemployed households having a dramatically lower expenditure compared to households with full or part time employment status and inactive employment status households having a lower expenditure compared to households with full or part time employment status.
-* Household size generally appears to be positively correlated with expenditure, with larger households having a higher weekly expenditure (however 5+ person households break this trend with the expenditure being slightly decreasing, however this could be due to other variables or a small sample size of such households).
+* Household size generally appears to be positively correlated with expenditure, with larger households having a higher weekly expenditure (however 5+ person households break this trend with the expenditure slightly decreasing, however this could be due to other variables or a small sample size of such households).
 * Number of adults in the household also appears positively correlated with expenditure with households containing more adults having a higher expenditure.
@@
 ## Part 2: Regressing expenditure on income
-Fitting the model $y_{i}=\beta_{0}+\beta_{1}x_{i}+\epsilon$ where $y$ is the weekly household expenditure and the explanatory variable $x$ is weekly income and $i \in \{1,...,1200\}$ is the ith household in the dataset produces the below table (these are the same for all linear models mentioned in parts 2 to 6).It shows the estimated values of the intercept ($\beta_0$) and the estimated 'effect' of a 1 pound increase of income on expenditure ($\beta_1$) and their associated standard errors.
+Fitting the model $y_{i}=\beta_{0}+\beta_{1}x_{i}+\epsilon_{i}$ where $y$ is the weekly household expenditure and the explanatory variable $x$ is weekly income and $i \in \{1,...,1200\}$ is the ith household in the dataset produces the below table (these are the same for all linear models mentioned in parts 2 to 6).
 \vspace{-5truemm}
@@
 Viewing the above diagnostic plots for this model.
-The fitted vs residual plot shows the residuals are not equally spread around 0 with a greater spread of residuals above zero, meaning the assumption of linearity may not hold.
+The fitted vs residual plot shows the residuals are not equally spread around 0 with a greater spread of residuals above zero, meaning the assumption of linearity is invalid.
 Furthermore, the QQ normal plot has a positive U shape with the tails diverging heavily away from the straight line.
@@
 ## Part 3: Regressing expenditure on income and income squared
-Fitting a new linear model $y_{i}=\beta_{0}+\beta_{1}x_{i}+\beta_2x^2_{i}+\epsilon$ with an added squared term of income gives the below estimated for the coefficients and their standard errors.
+Fitting a new linear model $y_{i}=\beta_{0}+\beta_{1}x_{i}+\beta_2x^2_{i}+\epsilon_{i}$ (adding a squared term of income) gives the below estimate for the coefficients and their standard errors.
 
@@
 #Producing table of results
-knitr::kable(as.data.frame(tidy(lm.income.sqd)[,1:3]),digits=3,caption = "Estimated coefficients and SE for part 3 model")%>%
+knitr::kable(as.data.frame(tidy(lm.income.sqd)[,1:3]),digits=5,caption = "Estimated coefficients and SE for part 3 model")%>%
   row_spec(0,bold=T)%>%
@@
 ```
-Looking at the produced diagnostic plots below for this model, in the fitted vs residual plot the residual values increase as the fitted values increase thus the homoscedasticity assumption may not be valid.
+Looking at the produced diagnostic plots below for this model, in the fitted vs residual plot the residual values increase as the fitted values increase thus the homoscedasticity assumption is invalid.
 The tails in the normal QQ plot also still deviate heavily away from the straight line meaning the normality assumption is invalid.
@@
 ## Part 4: Regressing the natural logarithm of expenditure on income
-Fitting the same model as in task 2 but with the log of the response expenditure (a log-linear model) $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\epsilon$ gives the below estimated coefficients and standard errors.
+Fitting the same model as in task 2 but with the log of the response expenditure (a log-linear model) $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\epsilon_{i}$ gives the below estimated coefficients and standard errors.
 
@@
 #Producing table of results
-knitr::kable(as.data.frame(tidy(lm.log.income)[,1:3]),digits=3,caption = "Estimated coefficients and SE for part 4 model")%>%
+knitr::kable(as.data.frame(tidy(lm.log.income)[,1:3]),digits=5,caption = "Estimated coefficients and SE for part 4 model")%>%
   row_spec(0,bold=T)%>%
@@
 ```
-Looking at the produced diagnostic plots above for this model, the data appears to be heteroscedastic as in the fitted vs residual plot the variance in the residuals decreases as the fitted values increases.
+Looking at the produced diagnostic plots below for this model, the data appears to be heteroscedastic as the variance of the residuals in the fitted vs residual plot decreases as the fitted values increases.
 Furthermore, the linear assumption also appears violated as the residuals appear to exhibit a slight negative quadratic shape.
@@
 
-Fitting the model $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\beta_{2}x^2_{i}+\epsilon$ gives the estimated coefficients and standard errors in the table below.
+Fitting the model $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\beta_{2}x^2_{i}+\epsilon_{i}$ gives the estimated coefficients and standard errors in the table below.
 \vspace{-5truemm}
@@
 #Producing table of results
-knitr::kable(as.data.frame(tidy(lm.log.income.sqd)[,1:3]),digits=3,caption = "Estimated coefficients and SE for part 5 model")%>%
+knitr::kable(as.data.frame(tidy(lm.log.income.sqd)[,1:3]),digits=5,caption = "Estimated coefficients and SE for part 5 model")%>%
   row_spec(0,bold=T)%>%
@@
 
-Producing the diagnostic plots below, the fitted vs residuals plot shows a fairly even scattering of points around 0 and a linear trend, however the variance of the residuals appears to decrease slightly as the fitted values increase.
-But, looking at the scale-location plot the residuals appear randomly spread around the red line and the red line through them stays fairly horizontal thus homoskedasticity assumptions appears valid.
+Producing the diagnostic plots above, the fitted vs residuals plot shows a fairly even scattering of points around 0 and a linear trend, however the variance of the residuals appears to decrease slightly as the fitted values increase.
+But, looking at the scale-location plot the residuals appear randomly spread around the red line which stays fairly horizontal thus the homoskedasticity assumption appears valid.
 The QQ plot shows the points stay close to the straight line meaning the normality assumption is valid.
-The residual vs Leverage plot shows no influential points.
+The residual vs leverage plot shows there no influential points.
 
@@
 The log-linear model means that the covarites (income and income squared) have a multiplicative effect on the response (expenditure).
-The squared term of income included in the model, means that the change in the mean of expenditure now depends on the value of income.
+The squared term of income included in the model, means that the change in the mean of expenditure depends on the value of income.
 Income has a positive association with expenditure with an increase in income causing a percent increase in expenditure.
@@
 Smaller initial values of income cause a greater percent increase on expenditure (due to the squared income term being negative).
-In particular, using a realistic increase in weekly income of £20 for a household earning the mean income of £512 gives an increase of $(\frac{exp(532\beta_1+532^{2}\beta_{2})}{exp(512\beta_1+512^{2}\beta_{2})}-1)$x100% = **3.25%**.
+In particular, using a realistic increase in weekly income of £20 for a household earning the mean income of £512 gives an increase of $(\frac{exp(532\beta_1+532^{2}\beta_{2})}{exp(512\beta_1+512^{2}\beta_{2})}-1)$x100% = **3.25%** in expenditure.
 
@@
 ```
-From the proportional table above there appears to be a relationship between the two variables.
-To investigate this further I created linear models including each variable separately and both together added to the model from part 5.
-In the model with only the number of adults variable added using an F-test the variable was significant at $\alpha=0.05$ with a p-value of **2.271e-11**.
-However, when adding both number of adults and household size only household size was significant with a p-value of **7.953e-15** and number of adults was no longer significant with a p-value of **0.7249** .
-Likely the household size was is confounding the effect of number of adults.
-Thus, I will include the size of a household in for the final model but not the number of adults.
+From the proportional table there is a clear relationship between the two variables.
+To investigate this further I created linear models based on the model selected in part 6 which I then added each covariate (household size and number of adults in household) separately and both together.
+In the model with only the covariate number of adults added using an F-test the variable was significant at $\alpha=0.05$ with a p-value of **2.271e-11**.
+However, when adding both the number of adults and household size only household size was significant with a p-value of **7.953e-15** and number of adults was no longer significant with a p-value of **0.7249** .
+This is due to the two covariates explaining much of the same variation in the data thus the number of adults becomes insignificant when both are included in the model.
+Therefore, I will consider the size of a household for the final model but not the number of adults.
 
@@
 First I investigated interactions between income and the factor variables.
-I added each separately to the model from part 5 and the variable part of the interaction and used an F-test to see if the interaction was significant at the 5% level.
-income:labour force, income: household size and income: house tenure type were all significant, however income:sex was not.
-Therefore I will consider the 3 significant interactions for my final model.
+I added each interaction separately to the model from part 5 and the variables included in the interaction then used an F-test to see if the interaction was significant at the 5% level.
+income:labour force, income: household size and income: house tenure type were all significant with p-values of **9.417e-05**, **0.002731** and **0.002709**, however income:sex was not with a p-value of **0.9657**.
+Therefore I will consider only the 3 significant interactions for my final model.
 
@@
 Next considering factor variable interactions.
-There are 6 possible factor interactions, of which I added each to the chosen model separately (plus the variables which were interacting) only sex:household size and labour:tenure were significant at the 5% level.
-Therefore, these are the only 2 I will consider for the final model.
+There are 6 possible factor interactions, of which I added each to the chosen model separately (plus the variables which were interacting) only sex:household size and labour:tenure were significant at the 5% level with p-values of **0.017408** and **0.033667**.
+Therefore, these are the only 2 way factor interactions I will consider for the final model.
 
@@
 To choose my final model I will start with a 'full' model that includes all variables that I want to consider.
-Then using AIC and adjusted $R^2$  as my model selection criteria I will remove explanatory variables (those that cause the AIC to become lower or increase the adjusted $R^2$  when removed).
+Then using AIC and my model selection criteria I will remove explanatory variables (those that cause the AIC to become lower when removed).
 After this as the model is for exploratory analysis and not prediction a simpler model that is easier to interpret/ understand is preferable (Occam's razor), therefore if the model can simplified to only include the important variables and this does not significantly affect the AIC and $R^2$ this further reduced model will be my final model.
@@
 My 'full' model regresses the natural logarithm of expenditure on income, income squared, household tenure type, household size, sex of household head, labour force and interactions between income with tenure type, labour force and household size and also interactions between sex of household head and household size and labour force and tenure type.
-This can be fitted in R using the belwo code:
+This can be fitted in R using the below code:
 
@@
 
-Then using the iterative method of deleting the variable that will reduce the AIC by the largest amount until no decrease in AIC is available.
-Implementing this algorithm I removed 1 variable, the interaction between labour force and household tenure type as this reduced the AIC from **1391.914** to **1386.878**.
+Then using the iterative method of deleting the variable that will reduce the AIC by the largest amount until no decrease in AIC is available I removed 1 variable.
+This was the interaction between labour force and household tenure type as it reduced the AIC from **1391.914** to **1386.878** when removed.
 All variables left in the updated model are also statistically significant except for income:house tenure.
@@
 I then checked how important (the magnitude of impact on the response) the variables left in the model are.
-To find their impact on expenditure I exponentiated the estimated coefficients and note that due to log-linear model the effect is multiplicative.
+To find their impact on expenditure I exponentiated the estimated coefficients and note that due to log-linear model the effect of the exponentiated regression coefficients is multiplicative on expenditure.
 
@@
 #Producing table of all exponentiated regession coefficients that are approx 1
-knitr::kable(data.frame(tidy(lm.full.1.deletion)[14:22,1],exp(tidy(lm.full.1.deletion)[14:22,2])),digits=3,caption="Coefficents of explantory variables with negliable affect on expenditure",)%>%
+knitr::kable(data.frame(tidy(lm.full.1.deletion)[14:22,1],exp(tidy(lm.full.1.deletion)[14:22,2])),digits=5,caption="Coefficents of explantory variables with negliable affect on expenditure",)%>%
   row_spec(0,bold=T)%>%
@@
 
-Above is a table of the exponentatied estimated coefficients of all the variables which were approximately 1.
-These are all the possible interaction levels (excluding the dummy level) of the income interactions.
+Above is a table of the exponentatied regression coefficients which were approximately 1.
+These were in fact all the possible interaction levels (excluding the dummy level) of the interactions with income.
 Thus all income interactions are not meaningful/important as their exponentiated value is approximately 1.
-Therefore, their effect on expenditure is negligible so I remove them from the model as a simpler model is preferred.
+Therefore, their effect on expenditure is negligible so I removed them from the model as a simpler model is preferred.
 After removing these 3 variables, removing any more does not further decrease the AIC.
-Furthermore the difference in AIC between the simpler and more complicated model with the simple models AIC being **1406.732** compared to the more complicated model AIC of **1386.878**.
-Thus this simplified model is my final model.
+Furthermore the difference in AIC between the simpler and more complicated model is small with the simpler models AIC being **1406.732** compared to the more complicated models AIC of **1386.878**.
+Thus the simplified model is my final model.
 
@@
 
-Producing diagnostic plots for this final model (seen below) shows no issues with the model fit.
-The residuals display a random scattering of points indicating the linearity assumption has been met and the residuals variance does not appear to increase/decrease except for a few values but I don't think this is enough evidence to say that the homoscedasticity assumption is invalid.
-This is further backed up by a straight nearly horizontal red line in the scale-location plot with an even random scattering of residuals around it.
+Producing diagnostic plots for this final model (seen below) there is no evidence of issues with the model fit.
+The residuals display a random scattering of points indicating the linearity assumption has been met and the variance of the residuals does not increase/decrease except for a few values, however, this isn't enough evidence to say that the homoscedasticity assumption is invalid.
+This is further backed up by a straight, nearly horizontal red line in the scale-location plot with an even random scattering of residuals around it.
 The points follow a straight line in the QQ plot with the tails deviating slightly but not extreme enough to indicate the normality assumption is invalid.
@@
 
-Due to the model being a log-linear model the explanatory variables have a multiplicative effect on the weekly expenditure.
-Exponentiating the estimated coefficients to find their effect on expenditure gives the below [*Table 7*].
-I note that all values have been rounded to 3 decimal places thus the zero p-values are due to p-values being too small be displayed and similarly the effect of income squared is too small to be shown.
-My following inference will be based on this table and I use a significance level of 5% for all results.
+Due to the model being a log-linear model the explanatory variables have a multiplicative effect on expenditure.
+Exponentiating the estimated coefficients to find their effect on expenditure gives the below table.
+I note that all values have been rounded to 4 decimal places thus the zero p-values are due to p-values being too small to be displayed and similarly for effect of income squared.
+My following inference will be based on this table and I use a significance level of 5%.
 
@@
 
-Point.estimates <- round(exp(tidy(lm.final.model)[,2]),3)
-p.values <- round(tidy(lm.final.model)[,5],3)
+Point.estimates <- round(exp(tidy(lm.final.model)[,2]),4)
+p.values <- round(tidy(lm.final.model)[,5],4)
 
@@
                  Point.estimates,
-                 "Confidence Interval 95%"=c(paste("(",round(lwr,3),", ",round(upr,3),")")),
-                 p.values),digits=3,caption = "Summary of exponentiated variables affecting expeniture")%>%
+                 "Confidence Interval 95%"=c(paste("(",round(lwr,4),", ",round(upr,4),")")),
+                 p.values),digits=4,caption = "Summary of exponentiated variables affecting expeniture")%>%
   row_spec(0,bold=T)%>%
@@
 ```
-First, interpreting the intercept value, it is for a male headed household of 1 person in full time employment in a publicly rented property who has an income of £0.
-Clearly this does not make sense (having a job which pays no income).
-
 The difference between full time work and being unemployed was highly statistically significant.
-With unemployed households compared to full time employment households having a 95% CI of [0.659, 0.915] which corresponds to a **8.5%** to **34.1%** decrease in expenditure with all other variables held constant.
+With unemployed households compared to full time employment households having a 95% confidence interval (CI) of [0.659, 0.915] which corresponds to a **8.5%** to **34.1%** decrease in expenditure with all other variables held constant.
 Furthermore, as the CI of unemployed and part-time employment do not overlap there is a statistically significant increase in expenditure if a household is in part time employment compared to unemployed households with all other variables held constant.
 
-However the difference between full time work and being inactive or in part time work was not significant (the 95% CI included 1) therefore with all other variables constant there is no evidence that being in one of these 3 factors compared to another influences the expenditure of the household.
+However the difference between full time work and being inactive or in part time work was not significant (the 95% CI included 1) therefore with all other variables constant there is no evidence that being in one of these three employment groups compared to another influences the expenditure of the household.
 
 The difference between a households expenditure for households which live in a publicly rented property and either a privately rented or owned property were statistically significant.
-With the associated 95% CI of [1.22, 1.441] and [1.115 , 1.281] which corresponds to a **22%** to **44.1%** and a **11.5%** **28.1%** increase in expenditure respectively with all other variables held constant.
+With the associated 95% CI of [1.22, 1.441] and [1.115 , 1.281] respectively which corresponds to a **22%** to **44.1%** and a **11.5%** to **28.1%** increase in expenditure respectively with all other variables held constant.
 
@@
 
-For households of size 1,2 and 4 a female household head corresponds to an increase in expenditure compared to a male household of size 1, however for a household of size 3 or 5+ there is not enough evidence to support a difference in expenditure (CI includes 1).
+For households of size 1,2 and 4 a female household head corresponds to an increase in expenditure compared to a male household of size 1, however for a household of size 3 or 5+ there is no evidence of a difference in expenditure (the CI includes 1).
 I note the confidence intervals are very large for female headed households of size 2 and above, this means there is alot of variation in the data causing the uncertainty in the results.
@@
 To demonstrate this I produced the plot below showing how the percent increase in expenditure varies with the initial income of a household for several different increases in household income.
-For all but the most extreme household incomes (over £1000), income has a positive effect on expenditure with the higher the initial income the less of a positive effect on expenditure an increase in income has (seen by negative gradient in below plot and which is due to the negative income squared regression coefficient).
+For all but the most extreme household incomes (slightly over £1000), income has a positive effect on expenditure.
+With the higher the initial income the less of a positive effect on expenditure an increase in income has (seen by negative gradient in below plot and which is due to the negative income squared regression coefficient).
 ```{r,include=F}
@@
 
-Overall an increase in income (except for extremely high incomes), living in a owned or privately rented property compared to a publicly rented property and being in full or part time employment compared to unemployed increases the expenditure of the household.
+Overall an increase in income (except for intially extremely high incomes), living in an owned or privately rented property compared to a publicly rented property and being in full or part time employment compared to unemployed increases the expenditure of the household.
 With household size increasing income if household head is male and being a female household head increases expenditure for certain household sizes (1, 2 and 4) and having no proven effect for others (3 and 5+).
 
-I note that even though the variables mentions above had a  statistically significant association with expenditure this could be spurious and due to confounding from variables not included in the model, simple chance or bias from the sample the model was based on.
-However, some areas of confidence of causation due to logical arguments are:
+It is important to remember that the relationships found in this analysis does not imply causation as they may be affected by confounding, chance or bias from the sample the model was based on.
+Including additional factors that could potentially be confounders such as the age of the household head, ethnicity of household and location of household would improve the confidence in the relationships found.
+Furthermore, a larger sample would enable the effect of sex of household head to be better understood for all household sizes.
 
-* Income, logically if you make more money you have more available to spend thus increasing expenditure
-* Household size, larger households are more expensive thus require more spending
-* Labour force, if you are unemployed compared to having work you are likely to be more frugal and thus reduce expenditure
-
-Further investigation is required for the affect of sex of household head as due to the small sub samples for some household sizes.
-Furthermore additional factors such as the age of the household head, ethnicity of household and location of household may improve the models accuracy.
-
-Also I note that generalization of these results may not be possible as different countries, cultures and ethnicities may have different attitudes on spending.
+I note that generalization of these results may not be possible as different countries, cultures and ethnicities may have different attitudes on spending.
 As there was no information on where/when the original sample was drawn from and the process for sample selection I cannot recommend what groups/ areas these results may be representative of.

### 2022-11-22 Draft_final.Rmd
Going over and slightly rewording inference

@@
 ```
-The model from part 5, $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\beta_{2}x^2_{i}+\epsilon$ has the lowest AIC value of **1528.062** and the highest adjusted R squared value of **0.4635055**.
+The model from part 5, $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\beta_{2}x^2_{i}+\epsilon_{i}$ has the lowest AIC value of **1528.062** and the highest adjusted $R^{2}$ value of **0.464** thus both model selection criteria select model 5 as the best fitting model of the data.
 Furthermore, from the diagnostic plots the normality assumption is valid and the fitted vs residual plot displays an even scattering of points around 0.
-Therefore the 4th model is my preferred model to describe the relationship between expenditure and income.
+Therefore the 4th model from part 5 is my preferred model to describe the relationship between expenditure and income.
 
-The log-linear model means that the covarites (income and income squared) have a multiplicative effect on the response (expenditure).
-The squared term of income included in the model, means that the change in the mean of expenditure depends on the value of income.
-Income has a positive association with expenditure with an increase in income causing a percent increase in expenditure.
-With the initial value of income affecting the magnitude of the percent increase on expenditure.
-Smaller initial values of income cause a greater percent increase on expenditure (due to the squared income term being negative).
-In particular, using a realistic increase in weekly income of £20 for a household earning the mean income of £512 gives an increase of $(\frac{exp(532\beta_1+532^{2}\beta_{2})}{exp(512\beta_1+512^{2}\beta_{2})}-1)$x100% = **3.25%** in expenditure.
+This is a log-linear model which means that the covarites (income and income squared) have a multiplicative effect on the response (expenditure).
+The squared term of income, means that the effect of income on expenditure depends on the initial value of income.
+Due to the negative value of the squared income regression coefficient the lower the initial value of income the greater the percent increase on expenditure for a given increase in income.
+Overall, income has a positive association with expenditure in the model with an increase in income causing a percent increase in expenditure (for all but the most extreme initial values of income of slightly over £1060).
+In particular, using a realistic increase in weekly income of £20 for a household earning the mean income of £512 gives an increase of $(\frac{exp(532\beta_1+532^{2}\beta_{2})}{exp(512\beta_1+512^{2}\beta_{2})}-1)$x100% = **3.25%** in household expenditure.
 
@@
 #calculating percent increase
+
+model.p5.increase <- function(initial,increase){
+  #values of regression coefficients
   beta_1 <- 3.117180e-03	
   beta_2 <- -1.453272e-06
-x <- 512
-inc <- 20
-#Formual to calculate oercent increase
-(exp(beta_1*(x+inc)+beta_2*(x+inc)^2)/exp(beta_1*x+beta_2*x^2)-1)*100
+
+ #Formula to calculate percent increase
+ perc.inc <- (exp(beta_1*(initial+increase)+beta_2*(initial+increase)^2)/exp(beta_1*initial+beta_2*initial^2)-1)*100
+
+ return(perc.inc)
+}
+
+model.p5.increase(512,20)
+
+#finding when increase in expenditure becomes negative
+model.p5.increase(1050,20)
+model.p5.increase(1060,20)
+model.p5.increase(1070,20)
 ```
@@
 I note that all values have been rounded to 4 decimal places thus the zero p-values are due to p-values being too small to be displayed and similarly for effect of income squared.
-My following inference will be based on this table and I use a significance level of 5%.
+The exponentiated estimate of the income squared regression coefficient is 0.99999896 (8 d.p.) with a 95% confidence interval (CI) of [0.99999866, 0.99999927].
+My following inference will be based on this table using a significance level of 0.05.
 
@@
 ```
-The difference between full time work and being unemployed was highly statistically significant.
-With unemployed households compared to full time employment households having a 95% confidence interval (CI) of [0.659, 0.915] which corresponds to a **8.5%** to **34.1%** decrease in expenditure with all other variables held constant.
+The difference between households in full time employment and those unemployed was highly statistically significant.
+With unemployed households compared to full time employment households having an exponentiated 95% CI of [0.659, 0.915] which corresponds to a **8.5%** to **34.1%** decrease in expenditure with all other variables held constant.
 Furthermore, as the CI of unemployed and part-time employment do not overlap there is a statistically significant increase in expenditure if a household is in part time employment compared to unemployed households with all other variables held constant.
-
-However the difference between full time work and being inactive or in part time work was not significant (the 95% CI included 1) therefore with all other variables constant there is no evidence that being in one of these three employment groups compared to another influences the expenditure of the household.
+However, there was no evidence to support a difference in expenditure between full time employment households and those that were inactive or in part time work (the exponentiated 95% CI included 1).
 
 The difference between a households expenditure for households which live in a publicly rented property and either a privately rented or owned property were statistically significant.
-With the associated 95% CI of [1.22, 1.441] and [1.115 , 1.281] respectively which corresponds to a **22%** to **44.1%** and a **11.5%** to **28.1%** increase in expenditure respectively with all other variables held constant.
+With an exponentiated 95% CI of [1.22, 1.441] and [1.115 , 1.281] respectively which corresponds to a **22%** to **44.1%** and a **11.5%** to **28.1%** increase in expenditure respectively with all other variables held constant.
 
 Sex of household head and household size have an interaction between them therefore cannot be reported separately.
-To easily display the effects of both on expenditure I created the below table which displays the multiplicative effect of sex and household size on expenditure.
+To easily display the effects of both on expenditure simultaneously I created the below 2x2 table which displays the multiplicative effect of sex and household size on expenditure.
 
@@
 
-For households of size 1,2 and 4 a female household head corresponds to an increase in expenditure compared to a male household of size 1, however for a household of size 3 or 5+ there is no evidence of a difference in expenditure (the CI includes 1).
-I note the confidence intervals are very large for female headed households of size 2 and above, this means there is alot of variation in the data causing the uncertainty in the results.
+For households of size 1,2 and 4 a female household head corresponds to an increase in expenditure compared to a male household of size 1, however for a household of size 3 or 5+ there is no evidence of a difference in expenditure (the exponentiated CI includes 1) with all other variables held constant.
+I note the confidence intervals are very large for female headed households of size 2 and above, this means there is alot of variation in the data causing this uncertainty in the results.
 This could be due to the small sample of female household heads of size 3 and above (insufficient power) or additional confounders affecting female household heads that were not included in the analysis.
@@
 Compared to a household of size 1, all larger sized households increase expenditure if the household head is male with an increase of expenditure of **22.4%** to **45.5%**, **36.2%** to **71.9%**, **31.3%** to **66.8%** and **34.9%** to **81.6%** for households of size 2, 3, 4 and 5+ respectively with all other variables held constant.
-There is no evidence larger households increase expenditure for female headed households as the 95% CI's overlap for all household sizes.
+There is no evidence larger households increase expenditure for female headed households as the 95% exponentiated CI's overlap for all household sizes.
 
@@
 For all but the most extreme household incomes (slightly over £1000), income has a positive effect on expenditure.
-With the higher the initial income the less of a positive effect on expenditure an increase in income has (seen by negative gradient in below plot and which is due to the negative income squared regression coefficient).
+With the higher the initial incomes the less the percent increase in expenditure for the same increase in income (seen by negative gradient in below plot, due to the negative income squared regression coefficient).
+For an initial income of £512 (the mean income) an increase in income of £20 and £50 increases expenditure by **2.44%** and **6.04%** respectively.
 ```{r,include=F}
-#calculating percent increase
+#calculating percent increase of expenditure from increasing income
+model.final.increase <- function(initial,increase){
+  #values of regression coefficients
   beta_1 <- 0.002285
   beta_2 <- -0.000001036
-x <- 900
-inc <- 20
 
-(exp(beta_1*(x+inc)+beta_2*(x+inc)^2)/exp(beta_1*x+beta_2*x^2)-1)*100
+ #Formula to calculate percent increase
+ perc.inc <- (exp(beta_1*(initial+increase)+beta_2*(initial+increase)^2)/exp(beta_1*initial+beta_2*initial^2)-1)*100
+
+ return(perc.inc)
+}
+
+model.final.increase(512,20)
+model.final.increase(512,50)
 
@@
 
-Overall an increase in income (except for intially extremely high incomes), living in an owned or privately rented property compared to a publicly rented property and being in full or part time employment compared to unemployed increases the expenditure of the household.
+Overall an increase in income (except for households on extremely high incomes), living in an owned or privately rented property compared to a publicly rented property and being in full or part time employment compared to unemployed increases the expenditure of a household.
 With household size increasing income if household head is male and being a female household head increases expenditure for certain household sizes (1, 2 and 4) and having no proven effect for others (3 and 5+).
 
-It is important to remember that the relationships found in this analysis does not imply causation as they may be affected by confounding, chance or bias from the sample the model was based on.
+It is important to remember that the relationships found in this analysis does not imply these variables cause an increase in expenditure as they may be affected by confounding, chance or bias from the sample the model was based on.
+However some areas of causal confidence are: (1) income, logically if you make more money you have more available to spend, (2) Employment status, if you are unemployed you are going to be more frugal with purchases.
 Including additional factors that could potentially be confounders such as the age of the household head, ethnicity of household and location of household would improve the confidence in the relationships found.

### 2022-11-23 Draft_final.Rmd
Another touch up of wording (on more brief go through then done for Task 1)

@@
 ---
-title: "General linear Models Coursework"
+title: "General linear Models (STAT6123) Coursework"
 author: "Student ID: 34273638"
@@
   geom_point(size=0.2)+
-  xlab("Gross weekly average household income (GBP)")+
-  ylab("Total expenditure")+
+  xlab("Gross weekly average household income (£)")+
+  ylab("Total expenditure (£)")+
   theme(axis.text=element_text(size=6),
-        axis.title=element_text(size=8))
+        axis.title=element_text(size=9))
 
@@
   xlab("Household tenure")+
-  ylab("Total expenditure")+
+  ylab("Total expenditure (£)")+
   guides(fill= "none")+
-  theme(axis.text.x = element_text(size = 6))
+  theme(axis.text.x = element_text(size = 6),
+        axis.title=element_text(size=9))
 
@@
   xlab("Sex of the household head")+
-  ylab("Total expenditure")+
-  guides(fill= "none")
+  ylab("Total expenditure (£)")+
+  guides(fill= "none")+
+  theme(axis.title = element_text(size = 9))
 
@@
   xlab("Employment status")+
-  ylab("Total expenditure")+
+  ylab("Total expenditure (£)")+
   guides(fill= "none")+
-  theme(axis.text.x = element_text(size = 6))
+  theme(axis.text.x = element_text(size = 6),
+        axis.title=element_text(size=9))
 
@@
   xlab("Household size")+
-  ylab("Total expenditure")+
+  ylab("Total expenditure (£)")+
   guides(fill= "none")+
-  theme(axis.text.x = element_text(size = 6))
+  theme(axis.text.x = element_text(size = 6),
+        axis.title=element_text(size=9))
 
@@
   xlab("Number of adults in the household")+
-  ylab("Total expenditure")+
-  guides(fill= "none")
+  ylab("Total expenditure (£)")+
+  guides(fill= "none")+
+  theme(axis.title = element_text(size = 9))
 
@@
 To assess the distribution of expenditure I produced the above histogram, it shows that expenditure has a positively skewed distribution (and thus the normality assumption used in linear regression may be invalid).
-Using a scatter plot to investigate the relationship between household income and expenditure, there is a clear positive association.
-Then to investigate the other categorical variables I used boxplots to display differences in the expenditure across the different levels (categories) of each variable.
-The differences shown in the plots are listed below:
+Using a scatter plot to investigate the relationship between household income and expenditure, there is a clear positive association with an increase in income associated with an increase in expenditure.
+Then to investigate the relationship between the categorical variables and expenditure I used boxplots to display differences in the expenditure across the different levels (categories) of each variable.
+The differences shown in the plots above are listed below:
 
-* Different household tenure categories have different weekly expenditure amounts with public rented households having a noticeably lower expenditure compared to the other 2 household levels.
-* Households with a female household head have a slightly lower weekly expenditure compared to male household heads.
-* Employment status appears to be associated with expenditure with unemployed households having a dramatically lower expenditure compared to households with full or part time employment status and inactive employment status households having a lower expenditure compared to households with full or part time employment status.
-* Household size generally appears to be positively correlated with expenditure, with larger households having a higher weekly expenditure (however 5+ person households break this trend with the expenditure slightly decreasing, however this could be due to other variables or a small sample size of such households).
-* Number of adults in the household also appears positively correlated with expenditure with households containing more adults having a higher expenditure.
+* Different household tenure categories have different weekly expenditure amounts with public rented households having a lower expenditure compared to the other two household tenure categories of owned and privately rented.
+* Households with a female household head have a slightly lower weekly expenditure compared to households with a male household heads.
+* Employment status is associated with expenditure, with unemployed households having a dramatically lower expenditure compared to households with full or part time employment status.
+Moreover, households with an inactive employment status have a lower expenditure compared to households with full or part time employment status.
+* Household size generally appears to have a positive association with expenditure, with larger households having a higher weekly expenditure (however 5+ person households break this trend with the expenditure slightly decreasing, however this could be due to other variables or a small sample size of such households).
+* Number of adults in the household is has a positive association with expenditure with households containing more adults having a higher expenditure.
 
-I note that these differences and possible associations may not be due to the variable mentioned but due to a third variable (a confounder) that affects both the categorical variable and expenditure (the response).
 \vspace{-5truemm}
@@
 ## Part 2: Regressing expenditure on income
-Fitting the model $y_{i}=\beta_{0}+\beta_{1}x_{i}+\epsilon_{i}$ where $y$ is the weekly household expenditure and the explanatory variable $x$ is weekly income and $i \in \{1,...,1200\}$ is the ith household in the dataset produces the below table (these are the same for all linear models mentioned in parts 2 to 6).
+Fitting the model $y_{i}=\beta_{0}+\beta_{1}x_{i}+\epsilon_{i}$ where the response $y$ is the weekly household expenditure and the explanatory variable $x$ is weekly income and $i \in \{1,...,1200\}$ is the ith household in the dataset produces the below table of the estimated coefficients and their standard errors (SE).
+I note these variables are the same for all models mentioned in parts 2 to 6.
 \vspace{-5truemm}
@@
 Viewing the above diagnostic plots for this model.
-The fitted vs residual plot shows the residuals are not equally spread around 0 with a greater spread of residuals above zero, meaning the assumption of linearity is invalid.
-Furthermore, the QQ normal plot has a positive U shape with the tails diverging heavily away from the straight line.
+The fitted vs residual plot shows the residuals are not equally spread around 0 with a higher density of residuals just below the 0 compared with above 0 (residual distribution is not symmetric about 0).
+This means the errors are not normally distributed, thus the assumption that the residuals are normally distributed is invalid.
+Furthermore, the QQ normal plot has a positive U shape with the top tail diverging heavily away from the straight line.
 This implies the normality assumption is invalid.
@@
 ```
-Looking at the produced diagnostic plots below for this model, in the fitted vs residual plot the residual values increase as the fitted values increase thus the homoscedasticity assumption is invalid.
+Looking at the produced diagnostic plots below for this model, in the fitted vs residual plot the residual variance increases as the fitted values increase thus the homoscedasticity assumption is invalid.
 The tails in the normal QQ plot also still deviate heavily away from the straight line meaning the normality assumption is invalid.
@@
 Looking at the produced diagnostic plots below for this model, the data appears to be heteroscedastic as the variance of the residuals in the fitted vs residual plot decreases as the fitted values increases.
-Furthermore, the linear assumption also appears violated as the residuals appear to exhibit a slight negative quadratic shape.
-The QQ plot shows the residuals stay close to the straight line thus normality assumption is valid.
+Furthermore, the linear assumption also appears violated as the residuals display a slight negative quadratic shape.
+The QQ plot shows the residuals stay close to the straight line thus the normality assumption is valid.
 
@@
 #Producing table of results
-knitr::kable(as.data.frame(tidy(lm.log.income.sqd)[,1:3]),digits=5,caption = "Estimated coefficients and SE for part 5 model")%>%
+knitr::kable(as.data.frame(tidy(lm.log.income.sqd)[,1:3]),digits=7,caption = "Estimated coefficients and SE for part 5 model")%>%
   row_spec(0,bold=T)%>%
@@
 
-Producing the diagnostic plots above, the fitted vs residuals plot shows a fairly even scattering of points around 0 and a linear trend, however the variance of the residuals appears to decrease slightly as the fitted values increase.
-But, looking at the scale-location plot the residuals appear randomly spread around the red line which stays fairly horizontal thus the homoskedasticity assumption appears valid.
+Producing the diagnostic plots above, the fitted vs residuals plot shows a fairly even scattering of points around 0 and a linear trend.
+The scale-location plot shows residuals that are randomly spread around the red line which stays fairly horizontal thus the homoskedasticity assumption is valid.
 The QQ plot shows the points stay close to the straight line meaning the normality assumption is valid.
@@
 The model from part 5, $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\beta_{2}x^2_{i}+\epsilon_{i}$ has the lowest AIC value of **1528.062** and the highest adjusted $R^{2}$ value of **0.464** thus both model selection criteria select model 5 as the best fitting model of the data.
-Furthermore, from the diagnostic plots the normality assumption is valid and the fitted vs residual plot displays an even scattering of points around 0.
+Furthermore, the diagnostic plots find no issues with the model assumptions.
 Therefore the 4th model from part 5 is my preferred model to describe the relationship between expenditure and income.
 
-This is a log-linear model which means that the covarites (income and income squared) have a multiplicative effect on the response (expenditure).
+The chosen model is a log-linear model which means that the covarites (income and income squared) have a multiplicative effect on the response (expenditure).
 The squared term of income, means that the effect of income on expenditure depends on the initial value of income.
-Due to the negative value of the squared income regression coefficient the lower the initial value of income the greater the percent increase on expenditure for a given increase in income.
-Overall, income has a positive association with expenditure in the model with an increase in income causing a percent increase in expenditure (for all but the most extreme initial values of income of slightly over £1060).
+Due to the negative value of the squared income regression coefficient the lower the initial value of income the greater the percent increase in expenditure for a given increase in income.
+Overall, income has a positive association with expenditure in the model with an increase in income causing a percent increase in expenditure (for all but the most extreme initial values of income slightly over £1060).
 In particular, using a realistic increase in weekly income of £20 for a household earning the mean income of £512 gives an increase of $(\frac{exp(532\beta_1+532^{2}\beta_{2})}{exp(512\beta_1+512^{2}\beta_{2})}-1)$x100% = **3.25%** in household expenditure.
@@
 ```
-From the proportional table there is a clear relationship between the two variables.
+From the proportional table of number of adults against people in household above there is a clear relationship between the two variables.
 To investigate this further I created linear models based on the model selected in part 6 which I then added each covariate (household size and number of adults in household) separately and both together.
@@
 This is due to the two covariates explaining much of the same variation in the data thus the number of adults becomes insignificant when both are included in the model.
-Therefore, I will consider the size of a household for the final model but not the number of adults.
+Therefore, I will only consider the size of a household for the final model and not the number of adults.
 
@@
 
-First I investigated interactions between income and the factor variables.
+First I investigated interactions between income and the categorical variables.
 I added each interaction separately to the model from part 5 and the variables included in the interaction then used an F-test to see if the interaction was significant at the 5% level.
-income:labour force, income: household size and income: house tenure type were all significant with p-values of **9.417e-05**, **0.002731** and **0.002709**, however income:sex was not with a p-value of **0.9657**.
-Therefore I will consider only the 3 significant interactions for my final model.
+Labour force, household size and house tenure type all had significant interactions with income with p-values of **9.417e-05**, **0.002731** and **0.002709**.
+Therefore, I will consider these 3 significant interactions for my final model.
 
@@
 ```
-Next considering factor variable interactions.
-There are 6 possible factor interactions, of which I added each to the chosen model separately (plus the variables which were interacting) only sex:household size and labour:tenure were significant at the 5% level with p-values of **0.017408** and **0.033667**.
-Therefore, these are the only 2 way factor interactions I will consider for the final model.
+Next considering the 6 possible 2 way categorical variable interactions.
+I added each to the chosen model from part 6 separately (plus the variables which were interacting) and found only interactions between sex of the household head and household size, and employment status and household tenure were significant at the 5% level with p-values of **0.017408** and **0.033667**.
+Therefore, these are the only 2 way categorical interactions I will consider for the final model.
 
@@
 Then using AIC and my model selection criteria I will remove explanatory variables (those that cause the AIC to become lower when removed).
-After this as the model is for exploratory analysis and not prediction a simpler model that is easier to interpret/ understand is preferable (Occam's razor), therefore if the model can simplified to only include the important variables and this does not significantly affect the AIC and $R^2$ this further reduced model will be my final model.
+After this as the model is for exploratory analysis and not prediction a simpler model that is easier to interpret/ understand is preferable (Occam's razor),.
+Therefore, if the model can simplified to only include the important variables and this does not significantly affect the AIC this further reduced model will be my final model.
 
-My 'full' model regresses the natural logarithm of expenditure on income, income squared, household tenure type, household size, sex of household head, labour force and interactions between income with tenure type, labour force and household size and also interactions between sex of household head and household size and labour force and tenure type.
+My 'full' model regresses the natural logarithm of expenditure on income, income squared, household tenure, household size, sex of household head, employment status and interactions between income and household tenure, employment status and household size and also interactions between sex of household head and household size and employment status and household tenure.
 This can be fitted in R using the below code:
@@
 Then using the iterative method of deleting the variable that will reduce the AIC by the largest amount until no decrease in AIC is available I removed 1 variable.
-This was the interaction between labour force and household tenure type as it reduced the AIC from **1391.914** to **1386.878** when removed.
+This was the interaction between employment status and household tenure type as it reduced the AIC from **1391.914** to **1386.878** when removed.
 All variables left in the updated model are also statistically significant except for income:house tenure.
@@
 
-Above is a table of the exponentatied regression coefficients which were approximately 1.
+Above is a table of all the exponentatied regression coefficients which were approximately 1.
 These were in fact all the possible interaction levels (excluding the dummy level) of the interactions with income.
@@
 After removing these 3 variables, removing any more does not further decrease the AIC.
-Furthermore the difference in AIC between the simpler and more complicated model is small with the simpler models AIC being **1406.732** compared to the more complicated models AIC of **1386.878**.
+Furthermore, the difference in AIC between the simpler and more complicated model is small with the simpler models AIC being **1406.732** compared to the more complicated models AIC of **1386.878**.
 Thus the simplified model is my final model.
@@
 Producing diagnostic plots for this final model (seen below) there is no evidence of issues with the model fit.
-The residuals display a random scattering of points indicating the linearity assumption has been met and the variance of the residuals does not increase/decrease except for a few values, however, this isn't enough evidence to say that the homoscedasticity assumption is invalid.
-This is further backed up by a straight, nearly horizontal red line in the scale-location plot with an even random scattering of residuals around it.
+The residuals display a random scattering of points indicating the linearity assumption has been met.
+The straight, nearly horizontal red line in the scale-location plot with an even random scattering of residuals around it indicates no issue with homoscedasticity.
 The points follow a straight line in the QQ plot with the tails deviating slightly but not extreme enough to indicate the normality assumption is invalid.
@@
 ## Part 8: Inference of final model
-Let $y$ denote the expenditure of the household, $x$ the income of the household, $\alpha^{G}$ the sex of the household head, $\alpha^{S}$ the size of the household, $\alpha^{T}$ the tenure type, $\alpha^{L}$ the labour force and $\alpha^{SG}$ the interaction between sex of household head and size of household.
+Let $y$ denote the expenditure of the household, $x$ the income of the household, $\alpha^{G}$ the sex of the household head, $\alpha^{S}$ the size of the household, $\alpha^{T}$ the household tenure type, $\alpha^{L}$ the employment status and $\alpha^{SG}$ the interaction between sex of household head and size of household.
 Then the final model is:
@@
 \end{equation*}
-Where $\alpha_{M}^{G}=\alpha_{1}^{S}=\alpha_{Priv}^{T}=\alpha_{FT}^{L}=\alpha^{SG}_{Mk}=\alpha^{SG}_{j1}=0$, $j \in \{M,F\}, k \in \{1,2,3,4,5+\}, w \in \{Pub,Priv,Own\}, z \in \{FT,PT,U,I\}$ and $i \in \{1,...,n_{jkwz}\}$ with $n_{jkwz}$ number households with the $j^{th}$ sex of household head, $k^{th}$ number of persons in the household, $w^{th}$ tenure type and $z^{th}$ labour force type.
+Where $\alpha_{M}^{G}=\alpha_{1}^{S}=\alpha_{Priv}^{T}=\alpha_{FT}^{L}=\alpha^{SG}_{Mk}=\alpha^{SG}_{j1}=0$, $j \in \{M,F\}, k \in \{1,2,3,4,5+\}, w \in \{Pub,Priv,Own\}, z \in \{FT,PT,U,I\}$ and $i \in \{1,...,n_{jkwz}\}$ with $n_{jkwz}$ number households with the $j^{th}$ sex of household head, $k^{th}$ number of persons in the household, $w^{th}$ household tenure type and $z^{th}$ employment staus type.
 \newline
@@
 
-Sex of household head and household size have an interaction between them therefore cannot be reported separately.
+Sex of household head and household size have an interaction between them and therefore cannot be reported separately.
 To easily display the effects of both on expenditure simultaneously I created the below 2x2 table which displays the multiplicative effect of sex and household size on expenditure.
@@
          "5+ Persons"="X5")%>%
-  kable(caption = "95% CI of multiplcative effect on expenditure compared to male headed household of 1 person",format = 'pandoc')
+  kable(caption = "95% CI of multiplicative effect on expenditure of sex of household head and household size compared to a male headed household of 1 person",format = 'pandoc')
 
@@
 
-For households of size 1,2 and 4 a female household head corresponds to an increase in expenditure compared to a male household of size 1, however for a household of size 3 or 5+ there is no evidence of a difference in expenditure (the exponentiated CI includes 1) with all other variables held constant.
+For households of size 1,2 and 4 a female household head corresponds to an increase in expenditure compared to a male household of size 1, however for a household of size 3 or 5+ there is no evidence of a difference in expenditure (the exponentiated CI included 1) with all other variables held constant.
 I note the confidence intervals are very large for female headed households of size 2 and above, this means there is alot of variation in the data causing this uncertainty in the results.
@@
 For all but the most extreme household incomes (slightly over £1000), income has a positive effect on expenditure.
-With the higher the initial incomes the less the percent increase in expenditure for the same increase in income (seen by negative gradient in below plot, due to the negative income squared regression coefficient).
+With the higher the initial income, the less the percent increase in expenditure for the same increase in income (seen by the negative gradient in the below plot).
 For an initial income of £512 (the mean income) an increase in income of £20 and £50 increases expenditure by **2.44%** and **6.04%** respectively.
@@
 #Creating plot to show change in percent of 
+beta_1 <- 0.002285
+beta_2 <- -0.000001036
+
 x <- seq(from=0,to=1200,by=10)
@@
   ylab("Increase in expenditure (%)")+
-  xlab("Initial income (£)")+
-  labs(colour = "Increase in\n income amount") +
+  xlab("Gross weekly average household income (£)")+
+  labs(colour = "Increase in\n weekly income amount") +
   ggtitle("Affect of income on expenditure using final model")+
@@
 Overall an increase in income (except for households on extremely high incomes), living in an owned or privately rented property compared to a publicly rented property and being in full or part time employment compared to unemployed increases the expenditure of a household.
-With household size increasing income if household head is male and being a female household head increases expenditure for certain household sizes (1, 2 and 4) and having no proven effect for others (3 and 5+).
+With household sizes larger than 1 increasing income compared to singular households if the household head is male and no proven difference if household head is female.
+A female headed household increases expenditure for certain household sizes (1, 2 and 4) and has no proven effect for others (3 and 5+) compared to a male headed household.
 
 It is important to remember that the relationships found in this analysis does not imply these variables cause an increase in expenditure as they may be affected by confounding, chance or bias from the sample the model was based on.
-However some areas of causal confidence are: (1) income, logically if you make more money you have more available to spend, (2) Employment status, if you are unemployed you are going to be more frugal with purchases.
-Including additional factors that could potentially be confounders such as the age of the household head, ethnicity of household and location of household would improve the confidence in the relationships found.
-Furthermore, a larger sample would enable the effect of sex of household head to be better understood for all household sizes.
+However, some areas of causal confidence are: (1) income, logically if you make more money you have more money available to spend, (2) Employment status, if you are unemployed you are going to be more frugal with purchases.
+Including additional factors that could potentially be confounding results such as the age of the household head, ethnicity of household and location of household would improve the confidence in the relationships found.
+Furthermore, a larger sample would enable the effect of the sex of the household head to be better understood for all household sizes.
 
 I note that generalization of these results may not be possible as different countries, cultures and ethnicities may have different attitudes on spending.
-As there was no information on where/when the original sample was drawn from and the process for sample selection I cannot recommend what groups/ areas these results may be representative of.
+As there was no information on where/when the original sample was drawn from and the process for the samples selection I cannot recommend what groups/areas these results may be representative of.

### 2022-11-23 Draft_final.Rmd
Final go through task 1 report

@@
 hist(expend.df$expenditure,breaks = 20,
-     xlab = "Total weekly household expenditure (GBP)",
+     xlab = "Total weekly household expenditure (£)",
      main = "Histogram of expenditure")
+```
+```{r,include=F}
+#calculating mean of expenditure
+mean.expend <- mean(expend.df$expenditure)
+mean.expend
+
+#finding number of households spending less than the mean
+sum(expend.df$expenditure<mean.expend)
+
+#percent of households earning less than the mean
+round(sum(expend.df$expenditure<mean.expend)/dim(expend.df)[1],3)
 ```
@@
 
-To assess the distribution of expenditure I produced the above histogram, it shows that expenditure has a positively skewed distribution (and thus the normality assumption used in linear regression may be invalid).
+To assess the distribution of expenditure I produced the above histogram, it shows that expenditure has a positively skewed distribution.
+A larger proportion of households (**56.6%**) spent less than the mean household expenditure of £396.
 Using a scatter plot to investigate the relationship between household income and expenditure, there is a clear positive association with an increase in income associated with an increase in expenditure.
-Then to investigate the relationship between the categorical variables and expenditure I used boxplots to display differences in the expenditure across the different levels (categories) of each variable.
+Then to investigate the relationship between the categorical variables and household expenditure I used boxplots to display differences in the expenditure across the different categories of each variable.
 The differences shown in the plots above are listed below:
@@
 * Different household tenure categories have different weekly expenditure amounts with public rented households having a lower expenditure compared to the other two household tenure categories of owned and privately rented.
-* Households with a female household head have a slightly lower weekly expenditure compared to households with a male household heads.
+* Households with a female household head have a slightly lower weekly expenditure compared to households with a male household head.
 * Employment status is associated with expenditure, with unemployed households having a dramatically lower expenditure compared to households with full or part time employment status.
@@
 * Household size generally appears to have a positive association with expenditure, with larger households having a higher weekly expenditure (however 5+ person households break this trend with the expenditure slightly decreasing, however this could be due to other variables or a small sample size of such households).
-* Number of adults in the household is has a positive association with expenditure with households containing more adults having a higher expenditure.
+* Number of adults in a household has a positive association with expenditure with households containing more adults having a higher expenditure.
 
@@
 Fitting the model $y_{i}=\beta_{0}+\beta_{1}x_{i}+\epsilon_{i}$ where the response $y$ is the weekly household expenditure and the explanatory variable $x$ is weekly income and $i \in \{1,...,1200\}$ is the ith household in the dataset produces the below table of the estimated coefficients and their standard errors (SE).
-I note these variables are the same for all models mentioned in parts 2 to 6.
+I note $y$ and $x$ represent the same variables for all models mentioned in parts 2 to 6.
 \vspace{-5truemm}
@@
 Viewing the above diagnostic plots for this model.
-The fitted vs residual plot shows the residuals are not equally spread around 0 with a higher density of residuals just below the 0 compared with above 0 (residual distribution is not symmetric about 0).
+The fitted vs residual plot shows the residuals are not equally spread around 0 with a higher density of residuals just below 0 compared with those above 0 (residual distribution is not symmetric about 0).
 This means the errors are not normally distributed, thus the assumption that the residuals are normally distributed is invalid.
@@
 ```
-Looking at the produced diagnostic plots below for this model, in the fitted vs residual plot the residual variance increases as the fitted values increase thus the homoscedasticity assumption is invalid.
+Looking at the produced diagnostic plots below for this model, in the fitted vs residual plot the residual variance increases as the fitted values increase, thus the homoscedasticity assumption is invalid.
 The tails in the normal QQ plot also still deviate heavily away from the straight line meaning the normality assumption is invalid.
@@
 ## Part 4: Regressing the natural logarithm of expenditure on income
-Fitting the same model as in task 2 but with the log of the response expenditure (a log-linear model) $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\epsilon_{i}$ gives the below estimated coefficients and standard errors.
+Fitting a log-linear model for expenditure, $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\epsilon_{i}$ gives the below estimated coefficients and standard errors.
 
@@
 
-Producing the diagnostic plots above, the fitted vs residuals plot shows a fairly even scattering of points around 0 and a linear trend.
+Producing the diagnostic plots above, the fitted vs residuals plot shows a fairly even scattering of points around 0 and a linear trend, thus the linearity assumption is valid.
 The scale-location plot shows residuals that are randomly spread around the red line which stays fairly horizontal thus the homoskedasticity assumption is valid.
@@
 ```
-The model from part 5, $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\beta_{2}x^2_{i}+\epsilon_{i}$ has the lowest AIC value of **1528.062** and the highest adjusted $R^{2}$ value of **0.464** thus both model selection criteria select model 5 as the best fitting model of the data.
+The model from part 5, $log(y_{i})=\beta_{0}+\beta_{1}x_{i}+\beta_{2}x^2_{i}+\epsilon_{i}$ has the lowest AIC value of **1528.062** and the highest adjusted $R^{2}$ value of **0.464**, thus both model selection criteria select model 5 as the best fitting model of the data.
 Furthermore, the diagnostic plots find no issues with the model assumptions.
@@
 The chosen model is a log-linear model which means that the covarites (income and income squared) have a multiplicative effect on the response (expenditure).
+Both variables were statistically significant at $\alpha = 0.05$.
 The squared term of income, means that the effect of income on expenditure depends on the initial value of income.
@@
 None of the variables had a GVIF greater than 5, thus multicolinearity is not an issue with the explanatory variables.
-However, considering the variables it seems likely that household size is related to number of household adults.
+However, considering the variables it seems likely that household size is related to the number of household adults.
 
@@
 ```
-From the proportional table of number of adults against people in household above there is a clear relationship between the two variables.
+From the proportional table above of the number of adults against people in a household there is a clear relationship between the two variables.
 To investigate this further I created linear models based on the model selected in part 6 which I then added each covariate (household size and number of adults in household) separately and both together.
-In the model with only the covariate number of adults added using an F-test the variable was significant at $\alpha=0.05$ with a p-value of **2.271e-11**.
+In the model with only the covariate the number of adults added, using an F-test the variable was significant at $\alpha=0.05$ with a p-value of **2.271e-11**.
 However, when adding both the number of adults and household size only household size was significant with a p-value of **7.953e-15** and number of adults was no longer significant with a p-value of **0.7249** .
@@
 I added each interaction separately to the model from part 5 and the variables included in the interaction then used an F-test to see if the interaction was significant at the 5% level.
-Labour force, household size and house tenure type all had significant interactions with income with p-values of **9.417e-05**, **0.002731** and **0.002709**.
+Employment status, household size and household tenure type all had significant interactions with income with p-values of **9.417e-05**, **0.002731** and **0.002709** respectively.
 Therefore, I will consider these 3 significant interactions for my final model.
@@
 Then using AIC and my model selection criteria I will remove explanatory variables (those that cause the AIC to become lower when removed).
-After this as the model is for exploratory analysis and not prediction a simpler model that is easier to interpret/ understand is preferable (Occam's razor),.
-Therefore, if the model can simplified to only include the important variables and this does not significantly affect the AIC this further reduced model will be my final model.
+After this as the model is for exploratory analysis and not prediction a simpler model that is easier to interpret/ understand is preferable (Occam's razor).
+Therefore, if the model can be simplified to only include the important variables and this does not significantly affect the AIC this further reduced model will be my final model.
 
@@
 Then using the iterative method of deleting the variable that will reduce the AIC by the largest amount until no decrease in AIC is available I removed 1 variable.
-This was the interaction between employment status and household tenure type as it reduced the AIC from **1391.914** to **1386.878** when removed.
-All variables left in the updated model are also statistically significant except for income:house tenure.
+This was the interaction between employment status and household tenure type, it reduced the AIC from **1391.914** to **1386.878** when removed.
+All variables left in the updated model are also statistically significant except for the interaction between income and household tenure.
 
@@
 
-I then checked how important (the magnitude of impact on the response) the variables left in the model are.
+I then checked how important (the magnitude of impact on the response) the variables left in the model were.
 To find their impact on expenditure I exponentiated the estimated coefficients and note that due to log-linear model the effect of the exponentiated regression coefficients is multiplicative on expenditure.
@@
 #Producing table of all exponentiated regession coefficients that are approx 1
-knitr::kable(data.frame(tidy(lm.full.1.deletion)[14:22,1],exp(tidy(lm.full.1.deletion)[14:22,2])),digits=5,caption="Coefficents of explantory variables with negliable affect on expenditure",)%>%
+knitr::kable(data.frame(tidy(lm.full.1.deletion)[14:22,1],exp(tidy(lm.full.1.deletion)[14:22,2])),digits=5,caption="Exponentiated coefficents of explantory variables with negliable effect on expenditure",)%>%
   row_spec(0,bold=T)%>%
@@
 \end{equation*}
-Where $\alpha_{M}^{G}=\alpha_{1}^{S}=\alpha_{Priv}^{T}=\alpha_{FT}^{L}=\alpha^{SG}_{Mk}=\alpha^{SG}_{j1}=0$, $j \in \{M,F\}, k \in \{1,2,3,4,5+\}, w \in \{Pub,Priv,Own\}, z \in \{FT,PT,U,I\}$ and $i \in \{1,...,n_{jkwz}\}$ with $n_{jkwz}$ number households with the $j^{th}$ sex of household head, $k^{th}$ number of persons in the household, $w^{th}$ household tenure type and $z^{th}$ employment staus type.
+Where $\alpha_{M}^{G}=\alpha_{1}^{S}=\alpha_{Priv}^{T}=\alpha_{FT}^{L}=\alpha^{SG}_{Mk}=\alpha^{SG}_{j1}=0$, $j \in \{M,F\}, k \in \{1,2,3,4,5+\}, w \in \{Pub,Priv,Own\}, z \in \{FT,PT,U,I\}$ (FT is full time, PT is part time, U is unemployed and I is inactive) and $i \in \{1,...,n_{jkwz}\}$ with $n_{jkwz}$ number households with the $j^{th}$ sex of household head, $k^{th}$ number of persons in the household, $w^{th}$ household tenure type and $z^{th}$ employment staus type.
 \newline
@@
 Exponentiating the estimated coefficients to find their effect on expenditure gives the below table.
-I note that all values have been rounded to 4 decimal places thus the zero p-values are due to p-values being too small to be displayed and similarly for effect of income squared.
+I note that all values have been rounded to 4 decimal places thus the zero p-values are due to the p-values being too small to be displayed and similarly for the effect of income squared.
 The exponentiated estimate of the income squared regression coefficient is 0.99999896 (8 d.p.) with a 95% confidence interval (CI) of [0.99999866, 0.99999927].
-My following inference will be based on this table using a significance level of 0.05.
+My following inference will be based on this table using a significance level of 5%.
 
@@
                  "Confidence Interval 95%"=c(paste("(",round(lwr,4),", ",round(upr,4),")")),
-                 p.values),digits=4,caption = "Summary of exponentiated variables affecting expeniture")%>%
+                 p.values),digits=4,caption = "Summary of multiplicative effect of variables on expenditure")%>%
   row_spec(0,bold=T)%>%
@@
 With unemployed households compared to full time employment households having an exponentiated 95% CI of [0.659, 0.915] which corresponds to a **8.5%** to **34.1%** decrease in expenditure with all other variables held constant.
-Furthermore, as the CI of unemployed and part-time employment do not overlap there is a statistically significant increase in expenditure if a household is in part time employment compared to unemployed households with all other variables held constant.
-However, there was no evidence to support a difference in expenditure between full time employment households and those that were inactive or in part time work (the exponentiated 95% CI included 1).
+Furthermore, as the CI's of unemployed and part-time employment do not overlap there is a statistically significant increase in expenditure if a household is in part time employment compared to unemployed households with all other variables held constant.
+However, there was no evidence to support a difference in expenditure between full time employment households and those that were inactive or in part time work (the exponentiated 95% CI's included 1).
 
-The difference between a households expenditure for households which live in a publicly rented property and either a privately rented or owned property were statistically significant.
+The difference between a households expenditure for households which live in a publicly rented property and either a privately rented or owned property was statistically significant.
 With an exponentiated 95% CI of [1.22, 1.441] and [1.115 , 1.281] respectively which corresponds to a **22%** to **44.1%** and a **11.5%** to **28.1%** increase in expenditure respectively with all other variables held constant.
@@
          "5+ Persons"="X5")%>%
-  kable(caption = "95% CI of multiplicative effect on expenditure of sex of household head and household size compared to a male headed household of 1 person",format = 'pandoc')
+  kable(caption = "95% CI of multiplicative effect on household expenditure of sex of household head and household size compared to a male headed household of 1 person",format = 'pandoc')
 
@@
 I note the confidence intervals are very large for female headed households of size 2 and above, this means there is alot of variation in the data causing this uncertainty in the results.
-This could be due to the small sample of female household heads of size 3 and above (insufficient power) or additional confounders affecting female household heads that were not included in the analysis.
+This could be due to the small sample of households with female heads of size 3 and above (insufficient power) or/and additional confounders affecting female household heads that were not included in the analysis.
 
-Compared to a household of size 1, all larger sized households increase expenditure if the household head is male with an increase of expenditure of **22.4%** to **45.5%**, **36.2%** to **71.9%**, **31.3%** to **66.8%** and **34.9%** to **81.6%** for households of size 2, 3, 4 and 5+ respectively with all other variables held constant.
+Compared to a household of size 1, all larger sized households increased expenditure if the household head was male with an increase of expenditure of **22.4%** to **45.5%**, **36.2%** to **71.9%**, **31.3%** to **66.8%** and **34.9%** to **81.6%** for households of size 2, 3, 4 and 5+ respectively with all other variables held constant.
 There is no evidence larger households increase expenditure for female headed households as the 95% exponentiated CI's overlap for all household sizes.
@@
 Overall an increase in income (except for households on extremely high incomes), living in an owned or privately rented property compared to a publicly rented property and being in full or part time employment compared to unemployed increases the expenditure of a household.
-With household sizes larger than 1 increasing income compared to singular households if the household head is male and no proven difference if household head is female.
+With household sizes larger than 1 increasing income compared to singular households if the household head is male and no proven difference if the household head is female.
 A female headed household increases expenditure for certain household sizes (1, 2 and 4) and has no proven effect for others (3 and 5+) compared to a male headed household.
 
-It is important to remember that the relationships found in this analysis does not imply these variables cause an increase in expenditure as they may be affected by confounding, chance or bias from the sample the model was based on.
+It is important to remember that the relationships found in this analysis does not imply causality between the variables and expenditure as they may be affected by confounding, chance or bias from the sample the model was based on.
 However, some areas of causal confidence are: (1) income, logically if you make more money you have more money available to spend, (2) Employment status, if you are unemployed you are going to be more frugal with purchases.

### 2022-11-23 Draft_final.Rmd
Updating script to add bits that will make sense in r script

@@
 ```
+
 ```{r,include=F}
@@
 ```{r,include=F}
+#finaing mean income
+mean(expend.df$income)
+
 #calculating percent increase
-
 model.p5.increase <- function(initial,increase){
@@
 AIC(lm.log.full)
+#AIC is 1391.914
 
@@
 AIC(lm.full.1.deletion)
+#AIC is 1386.878
 
@@
 AIC(lm.final.model)
+# AIC is 1406.732
 ```
@@
 \newline
+
+```{r, include=F}
+#Finding more precise value of income squared than table provides
+options(digits=10)
+
+#estimate of exp regression coefficient
+round(exp(tidy(lm.final.model)$estimate[6]),8)
+#0.99999896
+
+#lower and upper of 95% CI
+round(exp(tidy(lm.final.model)$estimate[6]-1.96*tidy(lm.final.model)$std.error[6]),8)
+#0.99999866
+round(exp(tidy(lm.final.model)$estimate[6]+1.96*tidy(lm.final.model)$std.error[6]),8)
+#0.99999927
+```
 

### 2022-11-23 Draft_final.R
USing purl to create R script

@@ -1,0 +1,597 @@
+#' ---
+#' title: "General linear Models (STAT6123) Coursework"
+#' author: "Student ID: 34273638"
+#' geometry: margin = 1.5cm
+#' output:
+#'   pdf_document: default
+#' header-includes:
+#'   - \usepackage{wrapfig}
+#'   - \usepackage{lipsum}
+#'   - \usepackage{setspace}
+#'   - \usepackage{titlesec}
+#'   - \titlespacing{\title}{0pt}{\parskip}{-\parskip}
+#' ---
+#' # Task 1:
+#' \vspace{-5truemm}
+#' ## Part 1: Assessing distribution of expenditure and relationship between it and covariates
+#' \vspace{-5truemm}
+## ----echo=F,message=F-----------------------------------------------------------------------------------
+#Initial environment setup
+setwd("~/Desktop/GLM/Coursework/GLM-CW")
+expend.df <- read.table("expenditure.txt")
+library(tidyverse)
+library(kableExtra)
+library(gridExtra)
+library(broom)
+library(MASS)
+
+#' 
+## ----echo=F---------------------------------------------------------------------------------------------
+#Data cleaning
+
+#coding variables as factors
+expend.df$house.ten <- factor(expend.df$house.ten,levels=c(1:3),labels=c("Public Rented","Private rented","Owned"))
+expend.df$sex.hh <- factor(expend.df$sex.hh,levels=c(1,2),labels=c("Male","Female"))
+expend.df$lab.force <- factor(expend.df$lab.force,levels=c(1:4),labels=c("Full time", "Part time","Unemployed","Inactive"))
+expend.df$hh.size <- factor(expend.df$hh.size,levels=c(1:5),labels = c("1 person","2 people","3 people","4 people","5+ persons"))
+expend.df$hh.adults <- factor(expend.df$hh.adults,levels = c(1:4), labels = c("1 adult","2 adults","3 adults","4+ adults"))
+
+
+#' 
+## ----echo=F,fig.align = 'left',fig.width=10,fig.height=4------------------------------------------------
+#Histogram to examine the expenditure
+hist(expend.df$expenditure,breaks = 20,
+     xlab = "Total weekly household expenditure (£)",
+     main = "Histogram of expenditure")
+
+#' 
+## ----include=F------------------------------------------------------------------------------------------
+#calculating mean of expenditure
+mean.expend <- mean(expend.df$expenditure)
+mean.expend
+
[…]

### 2022-11-24 Task_2_updated.R
Combining into submission format

@@ -1,9 +1,9 @@
----
-title: "Untitled"
-output: html_document
-date: "2022-11-11"
----
-```{r}
+#' ---
+#' title: "Untitled"
+#' output: html_document
+#' date: "2022-11-11"
+#' ---
+## -----------------------------------------------------------------------------------------------------------------------------------------------
 #Setting up enviroment
 setwd("~/Desktop/GLM/Coursework/GLM-CW")
 airline.df <- read.table("airline.txt",header=T)
@@ -13,9 +13,9 @@ n <- dim(airline.df)[[1]]
 con <- rep(1,n)
 X=matrix(c(con,airline.df$x),nrow=n,ncol=2)
 y <- airline.df$y
-```
 
-```{r}
+#' 
+## -----------------------------------------------------------------------------------------------------------------------------------------------
 #Part 2
 
 #Setting conditions of when to end fisher scoring algorithm
@@ -58,9 +58,9 @@ beta.iteration <- beta.ini[complete.cases(beta.ini),]
 beta.fin=beta.ini[c,]
 
 beta.fin
-```
 
-```{r}
+#' 
+## -----------------------------------------------------------------------------------------------------------------------------------------------
 #Part 3
 #finding the variance-covariance matrix
 solve(t(X)%*%X)
@@ -81,11 +81,11 @@ qnorm(0.975)
 
 #p-value calculation
 1-pnorm(T.stat)
-```
 
-```{r}
+#' 
+## -----------------------------------------------------------------------------------------------------------------------------------------------
 #Checking with R's own function
 fit <- glm(y~X[,2], family  = Gamma(link="log"))
 summary(fit,dispersion=1)  
-```
 
+#' 

### 2022-11-24 code-34273638.R

@@ -1,0 +1,686 @@
+## R template from STAT6123 Assignment 2022-23 ##
+
+## Task 1 ##
+
+# Clear R workspace
+rm(list=ls())
+
+#' ---
+#' title: "General linear Models (STAT6123) Coursework"
+#' author: "Student ID: 34273638"
+#' geometry: margin = 1.5cm
+#' output:
+#'   pdf_document: default
+#' header-includes:
+#'   - \usepackage{wrapfig}
+#'   - \usepackage{lipsum}
+#'   - \usepackage{setspace}
+#'   - \usepackage{titlesec}
+#'   - \titlespacing{\title}{0pt}{\parskip}{-\parskip}
+#' ---
+#' # Task 1:
+#' \vspace{-5truemm}
+#' ## Part 1: Assessing distribution of expenditure and relationship between it and covariates
+#' \vspace{-5truemm}
+## ----echo=F,message=F-----------------------------------------------------------------------------------
+#Initial environment setup
+setwd("~/Desktop/GLM/Coursework/GLM-CW")
+expend.df <- read.table("expenditure.txt")
+library(tidyverse)
+library(kableExtra)
+library(gridExtra)
+library(broom)
+library(MASS)
+
+#' 
+## ----echo=F---------------------------------------------------------------------------------------------
+#Data cleaning
+
+#coding variables as factors
+expend.df$house.ten <- factor(expend.df$house.ten,levels=c(1:3),labels=c("Public Rented","Private rented","Owned"))
+expend.df$sex.hh <- factor(expend.df$sex.hh,levels=c(1,2),labels=c("Male","Female"))
+expend.df$lab.force <- factor(expend.df$lab.force,levels=c(1:4),labels=c("Full time", "Part time","Unemployed","Inactive"))
+expend.df$hh.size <- factor(expend.df$hh.size,levels=c(1:5),labels = c("1 person","2 people","3 people","4 people","5+ persons"))
+expend.df$hh.adults <- factor(expend.df$hh.adults,levels = c(1:4), labels = c("1 adult","2 adults","3 adults","4+ adults"))
+
+
+#' 
+## ----echo=F,fig.align = 'left',fig.width=10,fig.height=4------------------------------------------------
+#Histogram to examine the expenditure
+hist(expend.df$expenditure,breaks = 20,
+     xlab = "Total weekly household expenditure (£)",
+     main = "Histogram of expenditure")
+
+#' 
[…]
